### 2018-09-07 rebuttal.tex
Initial ShareLaTeX Import

\documentclass{article}

\usepackage{nips_2018_author_response}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\begin{document}
We give thanks to all for spending the effort to understand our submission and for the encouraging feedback!

\textbf{Reviewer 1:}
We could not achieve the $\sqrt{T}$ with a relaxation algorithm. While it is not even known how to do this in the simpler case of bandit feedback, $\sqrt{T}$ would be a good contribution (we have some ideas). We will expand the related works section (especially adding citation about other partial feedback games) and will work on the unclear sections of the proof. Thanks for the specific suggestions; we will make the corrections and have tried to answer your questions below.

\textit{L125:} The matrix L is fixed, but $\ell_i$ is the expected loss under the adversary's empirical distribution and can still be impossible to learn. We'll clarify.
\textit{L129:} to calculate $v_{i,j,k}$, it suffices to find a pseudo inverse of $S$.
Alg1: EXP4 used the estimator described on L123, but the recentering is novel (actually concurrently proposed by [11]).
\textit{L258:} yes, this is a good point.
\textit{L262:} ``all fixed actions'' means that, for every action, we include the policy that always selects that action. A lower bound with just an arbitrary policy class was too weak of a result, as it is easy to make all the policies bad.
%\textit{L271:} rank nullity theorem

\textbf{Reviewer 3:}
Models:
Our intention was not to be coy about the i.i.d. assumption, which is critical (it is first mentioned in the abstract, L13), but we agree that all model assumptions should be very clear, and we will emphasize that the EXP4.PM regret bounds hold for adaptive $x_t$, $j_t$, but the relaxation algorithm requires $x_t$ to be i.i.d. We will clarify in the intro and the section headers.

We completely agree with the reviewer: the interesting questions are how $\Pi$ and $R_T$ interact. We have only taken the coarsest view, either finding bounds in terms of $\log |\Pi|$ or the Rademacher complexity-like term in Corollary 1. Having said that, no paper in the contextual bandit setting has a more refined analysis either. However, the partial monitoring setting is more nuanced than the bandit setting; indeed, if we take $\Pi$ to be the set of constant actions, then $\sqrt{T}$ regret is possible with only local observability, so there is possibly a more subtle boundary for the rates.

Additionally, there are more refined, game-dependent notions of complexity where the uniform distribution over the hypercube in the complexity term is replaced with a uniform distribution over the columns of $H$, and hence the regret will depend on the structure of the feedback through more than a dependence on the number of actions. We omitted these arguments for brevity (and they also require $O(N^2)$ computation), but are working on using them for a more refined bound.

The complexity measure in Corollary 1 is what is needed to be able to prove admissibility of the relaxation, but other cleaner notions may be possible. In the full information setting, sequential Rademacher complexities are needed to handle adversarial contexts, but it is unknown how to extend the relaxation approach to adversarial contexts in both bandits or partial monitoring.

Finally, the $j_t = f(x_t)$ case was studied in [4], but only when $f$ is a linear or logistic function; indeed $\sqrt{T}$ is possible without a pairwise observability condition. We will add this comparison.

\textbf{Reviewer 4:}
We will be more clear about the small algorithmic innovations and the intuition behind them. Providing a clear picture of the lower bound shorter than a few pages was difficult and we opted to describe the construction in detail. The high-level intuition is as follows. In the contextual case, pick arbitrary non-neighboring actions $i$ and $j$; there is a policy class where $i$ and $j$ are essentially neighbors in that determining the optimal policy will require resolving between the loss of $i$ and $j$. Hence, if there exists such a pair, then the algorithm will be forced to play other actions to resolve $\ell_i-\ell_j$, and the $T^{2/3}$ lower bound reasoning applies. We'll add this intuition and streamline the rest.

We will also address the rest of your fixes and tone down the optimality claims. Extra thanks for the correct inequality!

\end{document}

### 2018-09-07 rebuttal.txt
Initial ShareLaTeX Import

Rebuttal:

We give thanks to all for spending the effort to understand our submission and for the encouraging feedback!

Reviewer 1:
We tried to but could not achieve the sqrt(T) with a relaxation algorithm. It is not even known how to do this in the simpler case of bandit feedback, but we agree that sqrt(T) would be a good contribution (we have some ideas).

We will beef up the related works section (especially adding citation about other partial feedback games such as graph feedback) and will work on the unclear sections of the proof. Thanks for the specifics; we will make the corrections and have tried to answer your questions below.

L125: The matrix L is fixed, but ell_i is the expected loss under the adversary's empirical distribution and can still be impossible to learn. We'll clarify.
129: to calculate v, it suffices to find a pseudo inverse of S
Alg1: EXP4 used the estimator described on L123, but the recentering is novel (actually concurrently proposed by [11])
L258: yes, good point
L262: "all fixed actions" means adding the identity policies to \Pi. A lower bound with just an arbitrary policy class was too weak of a result, as it is easy to make all the policies bad.
L271: rank nullity theorem

Reviewer 3:

Models:
Our intention was not to be coy about the iid assumption, which is critical (it is first mentioned in the abstract, L13), but we agree that all model assumptions should be very clear, an we will emphasize that the exp4.pm regret bounds hold for adaptive x_t, j_t, but the relaxation algorithm requires x_t to be i.i.d. and also needs access to x_t samples. We will clarity in the intro and the section headers.

We completely agree with the reviewer: the interesting questions are how Pi and R_T interact. We have only taken the most coarse view, either finding bounds in terms of log |Pi| or the Rademacher complexity-like term in Corollary 1. In our defense, no paper in the contextual bandit setting has a more refined analysis either.

The complexity measure for Pi is derived by what is needed to be able to prove admissibility of the relaxation, but other cleaner notions may be possible. In the full information setting, sequential Rademacher complexities to handle adversarial contexts. It is unknown how to extend the relaxation approach to adversarial contexts in both bandit or partial monitoring, so the Rademacher-like term is the most natural in this setting. However, the partial monitoring setting is more nuanced than the bandit setting; indeed, if we take Pi to be the set of constant actions, then T^1/2 regret is possible with only local observability, so there is possibly a more subtle boundary for the rates.

Additionally, there are more refined, game-dependent notions of complexity where the uniform distribution over the hypercube in the complexity term is replaced with a uniform distribution over the columns of H, and hence the regret will depend on the structure of the feedback through more than a dependence on the number of actions. We omitted these arguments for brevity, but are working on using them for a more refined bound.

Finally, the j_t = f(x_t) case was studied in [4], but only in the case where f is a linear or logistic function; indeed sqrt(T) is possible without a pairwise observability condition. We will add this comparison.

Reviewer 4:
We will be more clear about the small algorithmic innovations and the intuition behind them. Providing a clear picture of the lower bound shorter than a few pages was difficult and we opted to describe the construction in detail. The high level intuition is as follows. In the contextual case, pick arbitrary non-neighboring actions i and j; there is a policy class there i and j are essentially neighbors in that determining the optimal policy will require resolving between the loss of i and j. Hence, if there exists such a pair, then the algorithm will be forces to play other actions to resolve ell_i-ell_j, and the usual lower bound implies a T^2/3 lower bound. We'll add this intuition and streamline the rest.

We'll also address the rest of your fixes and tone down the optimality claims and add dimensions. Extra thanks for the correct inequality!

### 2018-09-07 pap.tex
Initial ShareLaTeX Import

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
\PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2018

% ready for submission
\usepackage{nips_2018}

% to compile a preprint version, e.g., for submission to arXiv, add
% add the [preprint] option:
% \usepackage[preprint]{nips_2018}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2018}

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{fullpage}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables
\usepackage{bm}
\usepackage{amsfonts,dsfont}
\usepackage{comment}
\usepackage{mathtools,amsthm}

\bibliographystyle{plainnat}

\usepackage{xcolor}
\usepackage{tikz}
%\usepackage[colorinlistoftodos, textwidth=26mm, shadow,color=blue!30!white]{todonotes}
\usepackage[disable]{todonotes}

\usepackage{algorithmic, algorithm}
\usepackage{hyperref}
\usepackage{wrapfig}

\newcommand{\KL}{\mathrm{KL}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\exop}{\mathop{\ex}}
\newcommand{\proj}{\mathrm{proj}}
\newcommand{\regret}{\mathcal{R}}
\newcommand{\spn}{\mathrm{span}}
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\rel}{\mathbf{Rel}}
\newcommand{\hist}{{\mathcal H}}
\newcommand{\im}{{\mathrm{Im}\;}}

\newtheorem{assumption}{Assumption}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{corollary}{Corollary}
\newtheorem{theorem}{Theorem}

\input{def}

\title{Minimax Rates in Contextual Partial Monitoring}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
    Alan Malek
  \\MIT
  \texttt{amalek@mit.edu}
  \and
  Ali Jadbabaie\\
  \\MIT
  \and
  Alexander Rakhlin
  MIT
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
  We generalize the finite partial monitoring problem to the contextual setting. Partial monitoring allows learning even when the loss of the chosen action is not observed. In the noncontextual problem, the minimax regret is known to be $O(T^{2/3})$ if a global observability condition is satisfied and improves to $O(\sqrt{T})$ under a stronger local observability condition. Perhaps surprisingly, we show that the same characterization does not hold in the contextual case and a stronger notion of \emph{pairwise observability} is necessary for $O(\sqrt{T})$ minimax regret. In particular, we provide a lower bound of $O(T^{2/3})$ for any non-pairwise observable game, including locally observable games, in the contextual setting. We propose two algorithms for adversarial environments. The first requires a finite policy class but allows for arbitrary contexts and can be tuned to obtain the optimal $O(\sqrt{T})$ rate in pairwise observable settings or the optimal $O(T^{2/3})$ rate otherwise. The second allows for arbitrary policy classes with an empirical risk minimization oracle but requires i.i.d. contexts; we also show an $O(T^{2/3})$ upper bound and an efficient implementation using only a constant number of oracle calls per round.
  % We formulate the contextual finite partial monitoring problem, which allows sequential learning when the loss is not observable, and characterize the possible regret rates. We introduce the pairwise observability condition, which is stronger that the previous locally observability condition, and show that $O(\sqrt{T})$ regret is only possible in this condition by providing an algorithm with this rate as well as a $O(T^{2/3})$ lower bound when the condition does not hold, even if the game is locally observable. This result is in marked contrast to the noncontextual finite partial monitoring, where local observability was necessary and sufficient for $O(\sqrt{T})$ regret. In the non pairwise observable setting, we provide two algorithm, one for finite policy classes and one that only needs an empirical risk minimization oracle, both matching the $O(T^{2/3})$ lower bound. 
\end{abstract}

\section{Introduction}
In online learning, we model the world as a sequential game of $T$ rounds between the learner and a possibly adversarial environment. In particular, this paper studies the finite partial monitoring setting proposed in \cite{piccolboni2001discrete}, where, for each round $t$, the learner chooses an action $I_t \in \underbar N := \{1,\ldots,N\}$, the environment chooses a response $j_t \in \underbar M: = \{1,\ldots,M\}$, and the learner incurs loss $L_{I_t,j_t}$, the $I_t,j_t$ entry of a fixed and known loss matrix $L\in[0,1]^{N\times M}$. However, the learner does not observe $j_t$ or $L_{I_t,j_t}$, but rather $H_{I_t,j_t}$, the corresponding entry from the fixed and known feedback matrix $H\in[0,1]^{N\times M}$. Intuitively, one should imagine that each row of $H$ has only a few distinct elements and that observing $H_{I_t,j_t}$ only allows the learner to determine $j_t$ up to some subset of $\underbar M$. In particular, the loss incurred by the algorithm is not observed, making partial monitoring more difficult that bandit feedback.

For example, the partial monitoring game where each row of $H$ has distinct elements (i.e. $1,\ldots,M$) is equivalent to full information, as we can infer $j_t$ and therefore the full loss vector. The \textit{Revealing action} game is more interesting. Define
  $L = \begin{bmatrix}
    0 & a\\
    a & 0\\
    c & c
  \end{bmatrix}$
  and
  $H = \begin{bmatrix}
    0 & 0\\
    0 & 0\\
    1 & 2
  \end{bmatrix}$,
  which encodes the game where the learner tries to match the play of the adversary and incurs loss $a$ if incorrect. However, the learner only obtains useful feedback if she plays $I=3$ at a fixed cost $c$, in which case $j_t$ can be deduced. Partial monitoring allows games where the learner must choose between a low cost but uninformative action and a high cost but informative action.

Partial monitoring is more than a technical challenge and can be used to model practical scenarios. For example, consider dynamic pricing: at each round, the learner sets a price $p_t$ and a buyer with valuation $v_t$ arrives. If $p_t \leq v_t$, the good is purchased and the cost to the learner is $v_t - p_t$, the lost potential revenue from not setting a higher price. However, the learner only observes whether $v_t < p_t$ and not the loss (see \cite{bartok2014partial} for details). More generally, partial monitoring can model scenarios where the learner only sees a quantized version of the true loss, such as:
(i) surveys where a numerical values are binned, which have a long history as ``partial identification'' in the econometrics literature,
(ii) recommendation engines, where only coarse feedback from the recommendation (e.g. a like/dislike) is provided, but the learner wishes to find the best recommendations and not just maximize the number of likes, and 
(iii) robust algorithms, where we only estimate parameters to some confidence region but still want good performance.

Perhaps the biggest obstacle to more widespread adoption of partial monitoring as a modeling tool has been its inability to use context. The learner often has access to a context vector and hopes to choose more informed actions because of it. For example, the noncontextual problem for recommendation engines is tantamount to learning the single best item across all users, but is devoid of personalization.

With this motivation in mind, we propose the contextual partial monitoring problem. In addition to $L$ and $H$, the learner is also provided with a policy class $\Pi$ and, at every round, a context $x_t\in\mathcal X$. A policy $\pi:\mathcal X \rightarrow \underbar N$ maps a context to an action. We only consider deterministic policies, but our results extend to randomized policies in expectation. When it is clear from context, we will also use $\pi(x)$ to represent $e_{\pi(x)}$, the unit vector corresponding to the choice of the policy. The goal of the learner is to minimize the contextual regret,
the excess cumulative loss when compared to the best policy in $\Pi$:
\begin{equation}
  \regret_T := \sum_{t=1}^T L_{I_t,j_t} - \min_{\pi\in\Pi}\sum_{t=1}^T L_{\pi(x_t),j_t}.
\end{equation}

We will present algorithms for two typical models of the policy class. The first assumes that  $\Pi$ is finite, which allows individual weights on every policy. The second assumes access to an empirical risk minimization (ERM) value oracle, which, given some list of contexts $x_1,\ldots, x_t$ and losses $\ell_1,\ldots, \ell_t$, returns $\min_{\pi\in\Pi}\sum_{s=1}^t \pi(x_s)^\top \ell_s$. Note that only the value of the optimal policy is required. We will present algorithms for both settings.

\subsection{Related Work}
To the best of our knowledge, there are no results on the contextual partial monitoring problem with arbitrary policy classes. The closest work to ours, by \citet{bartok2012partial}, considered actions chosen by $f(x_t)$, where $f$ is some some fixed but unknown function. The algorithms construct explicit confidence regions for $f$ and the play optimistically.

The noncontextual partial monitoring problem, proposed by \citet{piccolboni2001discrete}, has been well studied. The \emph{global observability} condition was established as a necessary and sufficient for sublinear regret \cite{cesa2006regret} by a $O(T^{2/3})$ upper bound and matching $\Omega(T^{2/3})$ lower bound. A faster rate of $O(\sqrt{T})$ is possible when a stronger \emph{local observability} holds \cite{bartok2011minimax} which matches a $\Omega(T^{2/3})$ lower bound for all games without local observability. These results were later extended to stochastic adversaries \cite{foster2012no} with the same classification, and the effect of degenerate actions was recently resolved \cite{lattimore2018cleaning}. See the work of \citet{bartok2014partial} for more discussion.

Partial monitoring is an extension of a large body of work on games with incomplete information, including bandit feedback \cite{bubeck2012regret}, semi-bandit feedback \cite{audibert2013regret}, graph feedback \cite{alon2015online}, and many others. Finally, we note that the problem of partial monitoring has a very long history in the econometrics literature under the name partial identification; see e.g. \cite{marschak1944random,manski2009identification} and references therein.

Relaxation based algorithms, first proposed by \citet{rakhlinrelax}, have recently been extended to the contextual bandit setting \cite{rakhlin2016bistro,syrgkanis2016improved}. More generally, algorithms for contextual settings that leveraging empirical risk minimization oracles has been applied to other online learning algorithms, such as follow the perturbed leader \cite{dudik2016oracle}.

\subsection{Our Contributions}
We propose the contextual finite partial monitoring problem and characterize the possible minimax rates. In the noncontextual case, the \emph{local observability} condition, which essentially requires that similar losses can be distinguished solely from the feedback from playing those losses, was shown to be necessary and sufficient to obtain $O(\sqrt{T})$ regret. To the contrary, we show that in the contextual case, a much stronger \emph{pairwise observability} condition is needed.

In Section~\ref{sec:ew}, we provide two exponential weights algorithms that dynamically change the point of reference of the loss estimates and are able to capture $O(\sqrt{T})$ regret in the pairwise observability setting and $O(T^{2/3})$ regret in the globally observable setting. These results hold for adversarial contexts and actions but require a finite policy class.

Section~\ref{sec:relaxation} provides the first relaxation based algorithm in the partial monitoring setting and shows a $O(T^{2/3})$ regret in the finite policy class case. This setting requires access to unlabeled samples of the contexts but allows any policy class with an ERM value oracle. We also provide an efficient implementation requiring only $O(N)$  oracle calls per round.

We turn to lower bounds in Section~\ref{sec:lower.bound} and show that, without pairwise observability, no algorithm can achieve better that $\Omega(T^{2/3})$ expected regret. The main implication is that the algorithms from Section~\ref{sec:ew} capture the correct structural dependence on the regret. Additionally, to the best of our knowledge, this is the first time relaxation based algorithms obtain the minimax regret in an adversarial partial information setting, since the $\Omega(T^{2/3})$ lower matches the upper bound of Section~\ref{sec:relaxation}. There are no known tight bounds for the bandit setting.

\paragraph{Notation}
Collections over time are denoted by $a_{1:t} := a_1,\ldots, a_t $. The $i$th standard basis vector is denoted $e_i$, the all ones vector denoted $\ones$, and the $i$th element of vector $v$ denoted $v(i):= v^\top e_i$. Functions are applied to vectors elementwise; in particular, $\sgn(v)$ is a vector with $i$th element equal to the sign function of $v(i)$. In general, adversary distribution are denoted by $p_t$ and player distributions by $q_t$. Finally,  $\mathds{1}\{\cdot\}$ is the indicator function.

\section{Partial Monitoring with Context}
Obtaining low regret in a partial monitoring game requires careful study of loss structure, the feedback structure, and how the relate. For clarity, we break up the discussion and the definitions in this manner.

\paragraph{The Loss Structure} Define the loss vector $\ell_i = \e_i^\top L^\top$ to be the transpose of the row corresponding to playing action $i$. The cell corresponding to action $i$ is $C_i:=\{ p \in \triangle_M: \ell_i^\top p \leq \ell_j^\top p\; \forall j\}$, the set of stochastic adversary strategies for which action $i$ is optimal. The cells $C_1,\ldots, C_N$ induce a partition of $\triangle_M$. Action $i$ is degenerate if $C_i \subset C_j$ for some $j$. Two non-degenerate actions $i$ and $j$ are neighbors if $C_i\cap C_j$ is an $M-2$ dimensional polytope (e.g. there is a set of $p\in\triangle_M$ where $i$ and $j$ are both optimal) and denote the neighbor set of $(i,j)$ as $N_{i,j} := \{k\in\underbar N : C_i\cap C_j\subseteq C_k\}$, which includes $i$, $j$, and any action $k$ with $C_k \subseteq C_i\cap C_j$. We may not simply ignore actions like $k$ because action $k$ could provide information that action $i$ or $j$ cannot. Finally, the set of all pairs of neighboring actions is denoted $\mathcal N$.

\paragraph{The Feedback Structure}
We enumerate the distinct values of the $i$th row of $H$ by $\sigma_1,\ldots,\sigma_{s_i}$; when we play $i$, we observe feedback $\sigma$ and can conclude that $j_t$ must have been such that $H_{I_t,j_t} = \sigma$. Define the signaling matrix $(S_k)_{(i,j)} = \mathds{1}\{H_{(k,j)} = \sigma_i\}$ such that the $j$th row of $S_{k}$ indicated all choices $j_t$ that could have produced $\sigma_i$. Since the exact values of the $\sigma$ do not matter, we may assume that the feedback received is $Y_t = S_{I_t}e_{j_t}$.

\paragraph{Their Interaction}
Estimating the loss vectors $\ell_i$ is impossible for many easy games; for example, in the revealing action game, it is impossible to learn $A$, but we may still determine which is the optimal action. In fact, estimating the pairwise loss differences is sufficient for low regret. The three following definitions, presented in decreasing generality, encapsulate this notion.
\begin{definition}
  A partial monitoring game is \emph{globally observable} if, for all pairs $i,j$, there exists a collection of actions $V_{i,j}\subseteq \underbar N$ and observer vectors
  $\{ v_{i,j,k}\in\Reals^{s_k} |k\in V_{i,j}\}$, such that 
  \begin{equation}
  \ell_i - \ell_j = \sum_{k\in V_{i,j}} S_k^\top v_{i,j,k}.
\end{equation}
Throughout, we use $V_\infty = \max_{i,j,k} \Vert v_{i,j,k}\Vert_\infty$. Having $V_{i,j} = \underbar N$ is sufficient for global observability.

If, in addition, we make take $V_{i,j} = N_{i,j}$ for all neighbor pairs $(i,j)\in\mathcal N$, then the game is said to be \emph{locally observable}. Finally, if we may take $V_{i,j} = \{i,j\}$ for all non-degenerate $i$ and $j$, then the game is \emph{pairwise observable}.
\end{definition}

Intuitively, global observability means that we can construct unbiased estimates of the loss differences from the feedback by exploiting the equality $(\ell_i-\ell_j)^\top e_{j_t} = \sum_{k\in V_{i,j}}  v_{i,j,k}^\top S_k e_{j_t}$; the left hand side is the actual loss difference between action $i$ and $j$ and the right hand side can be estimated from the feedback $Y_t = S_{I_t}e_{j_t}$. Each round, we will choose a base arm $b_t$ and estimate the column vector
\begin{equation}
  \Delta_t = (L - \ones \ell_{b_t}^\top) e_{j_t}
\end{equation}
so that $ e_{I_t}^\top \Delta_t = L_{I_t,j_t} - L_{b_t,j_t}$. Out general strategy will be to carefully select $b_t$ every round and use an unbiased estimate of $\Delta_t$ as a proxy for the true losses. 
% \begin{figure}[h]
% \centering
% \fbox{\begin{minipage}{.9\columnwidth}
% Given: $T$,  $L$, $H$, and $\Pi$,\\
% For $t=1,2, \ldots, T$:
% \begin{itemize}\setlength{\itemindent}{-.25cm}
% \item The learner observes a context $x_t\in\mathcal X$
% \item Using $\Pi$, the learner chooses $p_t$ and samples  $I_t\sim p_t$
% \item The adversary simultaneously chooses $j_t$
% \item The learner incurs loss $L_{I_t,j_t}$ and observes $H_{I_t,j_t}$
% \end{itemize}
% \end{minipage}}
% \caption{Protocol}\label{fig:protocol}
% \end{figure}

\section{Exponential Weights Algorithms}
\label{sec:ew}
This section extends the exponential weights algorithm to the contextual partial monitoring setting and provides upper bounds in the globally observable and pairwise observable settings.
The estimator $\hat\Delta_t(i) := \sum_{k\in V_{i,b_t}} \mathds{1}\{k = I_t\}\frac{ v_{i,b_t,k}^\top S_k e_{j_t}}{q_t(k)}$ uses importance sampling and is an unbiased estimator for $\Delta_t(i)$. 
Written in terms of the feedback $Y_t = S_{I_t}e_{j_t}$,
\begin{equation}\label{eqn:hat.Delta}
  \hat\Delta_t = \left(
    \mathds{1}\{I_t\in V_{1,b_t} \}\frac{ v_{1,b_t,I_t}^\top Y_t}{q_t(I_t)},
    \ldots,
    \mathds{1}\{I_t\in V_{N,b_t} \}\frac{ v_{N,b_t,I_t}^\top Y_t}{q_t(I_t)}
  \right)^\top,
\end{equation}
and we can show its unbiasedness by calculating
\[
  \hat\Delta_t(i)=
  \ex\left[   \mathds{1}\{I_t\in V_{i,b_t} \}\frac{ v_{i,b_t,I_t}^\top S_{I_t}e_{j_t}}{q_t(I_t)}\right]
  =
  \ex\left [\sum_{k \in V_{i,b_t}} q_t(k) \frac{ v_{i,b_t,k}^\top S_{k}e_{j_t}}{q_t(k)}\right]
  =(\ell_i - \ell_{b_t})^\top e_{j_t}
  =
  \Delta_t(i).
  \]

  \begin{wrapfigure}[14]{r}{0.5\textwidth}
    \vspace{-1.7em}
\begin{minipage}{.5\textwidth}
\begin{algorithm}[H]
  \caption{Recentered EXP4.PM}
  \label{alg:exp4.pm}
   \begin{algorithmic}
      \STATE \textbf{Input: } $\eta$, $\gamma$, $T$, $L$,$H$, and $\Pi$
      \STATE Calculate observer vectors $v_{i,j,k}$
    \STATE Initialize $w_1 = 1/K$
      \FORALL{$t=1,\ldots,T$}
      \STATE Receive context $x_t$
      \STATE $q_t \leftarrow (1-N \gamma)\sum_{k=1}^K \pi_k(x_t) w_t(k) + \gamma\ones$
      \STATE Play $I_t\sim q_t$, observe $Y_t = S_{I_t} e_{j_t}$ 
      \STATE $b_t \leftarrow \argmax_i q_t(i)$
      \STATE Calculate
         $\hat\Delta_{t}$ from \eqref{eqn:hat.Delta}
      \STATE $w_{t+1}(k) \leftarrow w_t(k)e^{-\eta\pi_k(x_t)^\top \hat\Delta_t}$
      \ENDFOR
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{wrapfigure}

Our algorithm, \textsc{EXP4.PM}, is presented in Algorithm~\ref{alg:exp4.pm}. At a high level, it is \textsc{EXP4} using  $\hat\Delta_t$ for loss estimates with $\gamma$ uniform exploration and a recentering step that moves the base action to the arm with the highest weight. A similar idea was concurrently proposed for the noncontextual setting by \citet{lattimore2018cleaning}.

As the following theorem shows, we can always tune $\eta$ and $\gamma$ to guarantee a $O(T^{2/3})$ regret. If pairwise observability holds, we can obtain a faster rate by using $\gamma = 0$ and  by playing the subgame with the degenerate actions removed. This subgame does not have higher regret (there is always a non-degenerate action with loss no higher) and pairwise observability ensures that we construct unbiased estimates of $\Delta_t$ from the plays of non-degenerate actions only.
\begin{theorem}\label{thm:exp4.pm4.regret}  
  For any globally observable game, arbitrary sequence of contexts $x_1,\ldots,x_T$ and adversary actions $j_1,\ldots, j_T$, Algorithm~\ref{alg:exp4.pm} with
$\eta = N^{-\frac13} \left(\frac{\log(K)}{V_\infty T}\right)^{\frac23}$
and
$\gamma = \left( \frac{ V_\infty^2 \log(K)}{N^2 T}\right)^{\frac13}$
yields an expected regret with the bound
\[
  \ex[\Regret_T] \leq 3 \left(N V_\infty^2 \log(K)\right)^{\frac13} T^{\frac23}.
\]

If pairwise observability holds, the same algorithm with degenerate actions removed and parameters $\gamma = 0$ and 
 $\eta = \sqrt{ \frac{\log(K)}{ T V_\infty^2 (N+3)}}$ observes
\begin{equation*}
  \ex[\regret_T] \leq 2 V_\infty \sqrt{ T (N+3) \log(K)}.
\end{equation*}
\end{theorem}
The observability and optimal choice of $\gamma$ and $\eta$ is determined a priori by $L$ and $H$, and one can use the standard doubling trick if $T$ is unknown to obtain the same regret rates with a worse constant.

The proof of Theorem~\ref{thm:exp4.pm4.regret} is mostly identical to the standard \textsc{EXP4} proof. Recall that the \textsc{EXP4} importance weighted estimate, $\hat\ell_t = e_{I_t}\ell_t(I_t)/q_t(I_t)$, only has support on the $I_t$ entry, which allows the variance term in the analysis to be easily bounded by
$
\ex\left[\sum_k w_t(x) (\hat\ell_t \pi_k(x_t))^2\right]
\leq
\ex\left[V_\infty q_t^\top \hat\ell_t q_t(I_t)^{-2}\right]
$
since $q_t^\top \hat\ell_t = \ell_t(I_t)\leq 1$. In contrast, the importance weighted estimates $\hat\Delta_t$ may be non-zero for any entry. Without pairwise observability,  $q_t^\top \hat\Delta_t$ could have magnitude $\max_i 1/q_t(i)$ which we control by setting $\gamma > 0$. With pairwise observability, we may choose $\hat\Delta_t$ to be supported on $e_{I_t}$ and $e_{b_t}$ only, allowing us to control $q_t^\top \hat\Delta_t$ by choosing $b_t$ such that $q_t(b_t)$ is not too small. The full proof is in Appendix~\ref{sec:ew.proof}.

\section{A Relaxation Algorithm}
\label{sec:relaxation}
We now turn our focus to policy classes where the only assumption made is access to an ERM value oracle, thereby extending the relaxation framework for contextual bandits \citep{rakhlin2016bistro} to the contextual partial monitoring setting. We first review the necessary details of the relaxation framework before describing an efficient (in terms of the number of oracle calls) algorithm with $O(T^{2/3})$ regret, which will match the lower bound of Section~\ref{sec:lower.bound}.

\subsection{A Sparser Offset Loss Estimate}
The regret analysis for relaxation algorithms requires careful control of the sparsity of the offset loss estimates. Throughout this section, fix a base arm $b$. Define $V(i) = [v_{1,b,i},\ldots,v_{N,b,i}]$, which implies that $\Delta_t = \sum_k V(k)^\top S_k e_{j_t}$. Instead of constructing an unbiased estimate for $\Delta_t$ from importance weighting, we will instead borrow a trick from \cite{syrgkanis2016improved} and correct for bias by multiplying $V(I_t)^\top Y_t$ by a Bernoulli random variable with expectation $\propto 1/q_t(I_t)$, as described by the following lemma. 
\begin{lemma}
  Assume that $q_t(i)\geq \gamma$ for all $i$ and define $\hat Z_t = V_\infty \gamma^{-1} \diag(\hat B_{1,t},\ldots,\hat B_{N,t})$ for
  $\hat B_{i,t} \sim \mathrm{Bernoulli}\left(\frac{\gamma}{V_\infty} \frac{|e_i^\top V(I_t)^\top Y_t|}{q_t(I_t)}\right)$.
Then, the following offset loss estimate is unbiased:
  \begin{equation}
    \hat\Delta_t := \hat Z_t \sgn( V(I_t)^\top Y_t).
  \end{equation}
\end{lemma}
\begin{proof}
  First, the probability that $\hat B_{i,t} = 1$ is well defined: $q_t(i)\geq \gamma$ and, since $Y_t = S_k e_{j_t}$ is a unit vector, $|V(i)^\top Y_t|\leq V_\infty$.
  %Using $\mathcal F_t$ for the filtration generated by $I_t$, $x_t$, $j_t$, and $\hat Z_t$,
  We can directly verify that
  \begin{align*}
    \ex[\hat\Delta_t(i)]
    &=
      \ex\left[    
      \sgn\left(e_i^\top V(I_t)^\top Y_t\right) V_\infty \gamma^{-1} \ex\left[\hat B_{i,t}\middle|I_t\right]\right]\\
    &=
      \ex\left[
      \sgn\left(e_i^\top V(I_t)^\top S_{I_t} e_{j_t}\right)
      \frac{\left|e_i^\top V(I_t)^\top S_{I_t} e_{j_t}\right|}{q_t(I_t)}\right]
      = \Delta_t(i).
  \end{align*}
\end{proof}

\subsection{Relaxations}
The relaxation framework allows one to simultaneous derive algorithms and upper bounds on regret for sequential learning problems. We will keep our description short and refer the reader to \cite{rakhlinrelax} and \cite{rakhlin2016bistro} for elaboration on the general sequential prediction and bandit feedback settings, respectively.

During round $t$, the algorithm collects history $(x_t, I_t, q_t, Y_t, \hat Z_t)$. Let $\hist^{t} = \hist^{t-1}\cap (x_t, I_t, q_t, Y_t, \hat Z_t)$ with $\hist^0=\emptyset$.
\begin{definition}
  A relaxation $\rel(\cdot)$ is a function from $\hist^{t}$ to $\Reals$ for all $t=1,\ldots, T$. It is admissible if,
  \begin{enumerate}
    \item for all $x_{1:T}$, $j_{1:T}$, and $q_{1:T}$,
  \begin{equation}\label{def:relax.base}
    \ex_{I_{1:T}\sim q_{1:T},\hat Z_{1:T}}[\rel(\hist^{T})] \geq -\inf_{\pi\in\Pi} \sum_{t=1}^T \pi(x_t)^\top \Delta_t, \text{ and }
      \end{equation}
  \item for all $t=1,\ldots,T$ and any history $\hist^{t-1}$,
  \begin{align}\label{def:relax.induction}
    &\ex_{x_t}\left[ \inf_{q_t} \sup_{j_t } \ex_{I_t\sim q_t, \hat Z_t}\left[ e_{I_t}^\top \Delta_t + \rel(\hist^{t-1}\cup \hist^t)\right]\right]
      \leq \rel(\hist^{t-1}).
  \end{align}
\end{enumerate}
  Furthermore, any strategy $q_t$ which satisfies \eqref{def:relax.induction} is called admissible. 
\end{definition}
The first condition ensures that $\rel(\hist^T)$ is an upper bound on the offset loss of the comparator, and the second condition ensures that, under the player strategy $q_t$, the relaxation remains an upper bound against all $j_t$. The main utility of $\rel$ is that it produces a bound on the regret, even though it is defined it terms of the offset losses.
\begin{lemma}\label{lem:relaxation.to.regret}
  If $\rel(\cdot)$ is an admissible relaxation, then for any $j_{1:T}$, we have
  \begin{equation*}
    \ex[\Regret_T] \leq \rel(\emptyset).
  \end{equation*}
\end{lemma}
\begin{proof}
  Applying Lemma~1 from \citet{rakhlin2016bistro} with $c_t = \Delta_t$,
\  \begin{align*}
    \rel(\emptyset)&\geq \ex[\sum_{t=1}^T q_t^\top \Delta_t - \inf_{\pi\in\Pi}\sum_{t=1}^T \pi(x_t)^\top\Delta_t]\\
     &=
       \ex[\sum_{t=1}^T q_t^\top(L -\ones\ell_b^\top)  e_{j_t}
       -
       \inf_{\pi\in\Pi}\sum_{t=1}^T \pi(x_t)^\top(L -\ones\ell_b^\top)  e_{j_t}]= \ex[\regret_T] .
  \end{align*}
\end{proof}

\subsection{An Admissible Relaxation}
The relaxation framework provides at automatic regret bound for any admissible relaxation; the challenge typically is in proving admissibility. The relaxation we use is is modeled on the relaxation for the adversarial contextual bandit settings \cite{rakhlin2016bistro, syrgkanis2016improved} and uses sequential Rademacher averages and a random rollout.

\paragraph{Notation} Let $\eepsilon_{t,i}$ denote the matrix with zero elements except for a Rademacher random variable (equal probability $\{1,-1\}$) in the $i,i$ entry, $\eepsilon_t$ denote their collection, and $\Z_t$ denote the random vector with i.i.d. coordinates equal to $V_\infty\gamma^{-1}$ with probability $\gamma$ and equal to $0$ otherwise. We will use $\mathcal D_1 = \{-V_\infty \gamma^{-1},0,V_\infty \gamma^{-1}\}$ and $\mathcal D = \{x\in\Reals^N: x^\top e_i \in \mathcal D_1\forall i\}$. We also define $\triangle_{\mathcal D}'$ to be distributions such that, if $X\sim p\in \triangle_{\mathcal D}'$, then $p(X^\top e_i = V_\infty\gamma^{-1})\leq N\gamma$ and $p(X^\top e_i = -V_\infty\gamma^{-1})\leq N\gamma$ for all $i$. Finally, we write $D_i = \diag(e_i)$ so that $u^\top v = \sum_{i=1}^N u^\top D_i v$.

The relaxation at round $t$ is a function of the past data, encapsulated in $\hist^{t-1}$, and some randomly drawn data representing the uncertainty in the future $\rho_t = \bigcup_{s \in t+1:T} \{x_{s},\eepsilon_{s},\Z_{s}\}$, where $x_s$ for $s>t$ are i.i.d. samples from the context distribution. Define
\begin{align}
    R_i\left(\hist^{t},\rho_t\right) =
  \sup_{\pi\in\Pi}
  -
  \underbrace{\sum_{s=1}^{t} \pi(x_s)^\top D_i \hat\Delta_s}_
  {\text{past data}}
  -
  \underbrace{\sum_{s=t+1}^{T} 2 \pi(x_s)^\top \eepsilon_{s,i} Z_s}_
  {\text{future uncertainty}}
  + (T-t)\gamma,
\end{align}
which is best fit of $\Pi$ to the $i$th coordinate of the past and future data.
\begin{theorem}\label{thm:admissibility}  
  The relaxation
  \begin{align}
    \rel(\hist^{t-1})
    &= \exop_{
      \rho_t}
      \left[
      \sum_{i=1}^N R_i\left(\hist^{t-1},\rho_t\right)\right]
 \label{eqn:relaxation.def}
  \end{align}
  and the strategy that samples $\rho_t = \bigcup_{s \in t+1:T} \{x_{s},\eepsilon_{s},\Z_{s}\}$ and plays  $q_t= (1-N\gamma)q_t^* + \gamma \ones$ for
\begin{align}
  q_t^*
  &=
  \argmin_{q \in \triangle_{N}}
  \sup_{p_t \in \triangle_{\mathcal D}'} 
    \exop_{\hat\Delta_t \sim p_t}\left[
    q^\top\hat\Delta_t
    +   \sum_{i=1}^N  R_i\left(\hist^{t-1},\rho_t\right)\right]
  \label{eqn:strategy.def}
\end{align}
are admissible.
\end{theorem}
The optimization objective in \eqref{eqn:strategy.def} has $\hat\Delta_t$ appearing in two places: the $q^\top\hat\Delta_t$ and $\hist^{t}$, and hence the $p_t$ optimization accounts for the worst case adversary action considering the loss introduced at round $t$ as well as the potential future losses (from the $R_i(\hist^t,\rho_t)$ terms). The algorithm picks $q_t$ to mitigate the regret caused by the worst case $p_t$.

% \begin{algorithm}[tb]
%    \caption{Relaxations for partial monitoring}
%    \label{alg:bistro.pm}
%    \begin{algorithmic}
%       \STATE\textbf{Input: } Game length $T$, $L$,$H$, and $\Pi$
%       \STATE Calculate observer matrices $V(1),\ldots,V(N)$
%     \STATE Initialize $p_1 = 1/K$
%       \FORALL{$t=1,\ldots,T$}
%       \STATE Receive context $x_t$
%       \STATE Sample future data $(x,Z,\eepsilon)_{t+1:T}$
%       \STATE Calculate $q_t^*$ from \eqref{eqn:strategy.def}
%       \STATE Draw $I_t\sim q_t = (1-N \gamma) q^*_t + \gamma\ones$
%       \STATE Observe $Y_t$
% \ENDFOR
% \end{algorithmic}
% \end{algorithm}

Since $\rel(\hist^t)$ is admissible, we may apply Lemma~\ref{lem:relaxation.to.regret} to $\rel(\emptyset)$ and obtain a regret bound. The random variable $\Z_t$ has maximum magnitude $\gamma^{-1}$, and hence we can optimize over $\gamma$ to obtain a sub-linear regret, as stated in the following corollary.
\begin{corollary}\label{cor:oracle.regret.bound}
  The algorithm that plays the $q_t$ defined in Theorem~\ref{thm:admissibility} has the regret bound
  \begin{equation*}
    \ex[\regret_t]
    \leq
    \exop_{\Z_{1:T},\eepsilon_{1:T}}\left[\sum_{i=1}^N\sup_{\pi\in\Pi} \sum_{t=1}^T \pi(x_t)^\top\eepsilon_{t,i} \Z_t \right]
    +
    T N \gamma.
  \end{equation*}
  If the policy class is finite, choosing $\gamma = \left(V_\infty\log(|\Pi|)/(2 TN) \right)^{\frac13}$ produces 
  \begin{equation}\label{eqn:rel.finite.regret}
    \ex[\Regret_T] \leq N \left(4 V_\infty \log(|\Pi|)\right)^{\frac13} T^{\frac23}.
  \end{equation} 
\end{corollary}
With details in Appendix~\ref{sec:rel.proofs}, 
the second claim follows from optimizing $\gamma$ over the bound
$\exop_{\Z_{1:T},\eepsilon_{1:T}}\left[\sum_{i=1}^N\sup_{\pi\in\Pi} \sum_{t=1}^T \pi(x_t)^\top\eepsilon_{t,i} \Z_t \right] \leq N\sqrt{2T\gamma^{-1}V_\infty \log(|\Pi|)}$.

\subsection{Computation}
At first glance, the relaxation algorithm, which samples a random rollout $\rho_t$ then computes 
$
  \argmin_{q \in \triangle_{N}}
  \sup_{p_t \in \triangle_{\mathcal D}'} 
    \exop_{\hat\Delta_t \sim p_t}\left[
    q^\top\hat\Delta_t
    +  \sum_{i=1}^N R_i \left(\hist^{t},\rho_t\right)
    \right]
    $,
does not see easy to compute. Fortunately, it is possible to exploit the structure of $\triangle_{\mathcal D}'$ and compute $q_t^*$ using only $3N$ oracle calls per round. We give pseudocode in Algorithm~\ref{alg:qstar} and specify every oracle query. The objective with the $\sup_{p_t}$ resolved is a convex function in $q_t$ which has a local minimum for any $q$ in the rectangle $[a,b]:=[a_1,b_1]\times\ldots\times[a_N,b_N]$ and slope with constant magnitude in all coordinates outside of $[a,b]$. Hence, any $q\in\triangle_N$ with minimum $\Vert\cdot\Vert_1$ distance to $[a,b]$ will be optimal. Consider the case when $\sum_i a_i \leq 1 \leq \sum_i b_i$. With $q_i(x) = a_i\mathds{1}\{x\leq a_i\} + x\mathds{1}\{a_i<x<b_i\} + b_i\mathds{1}\{x\geq b_i\}$, we return an $x$ such that $\sum_i q_i(x) = 1$; this is doable is $O(N)$ time because $q_i(x)$ is a increasing, piecewise linear function where the slope changes every time $x$ passes some $a_i$ or $b_i$; start $x_0 = a$ and increase $x$ until $\sum_i q_i(x)=1$. In the case where $\sum_i b_i \leq 1$, one can perform regular water-filling starting $b$, and if $\sum_i a_i \geq 1$, one can perform water-draining starting from $a$. See Appendix~\ref{sec:computation.appendix} for details and proofs.

\begin{lemma}\label{lem:computation}
  Algorithm~\ref{alg:qstar} correctly calculates $q_t^*(\rho)$, has complexity $O(N)$, and requires only $3N$ oracle calls.
\end{lemma}

\begin{algorithm}[tb]
   \caption{Computing $q_t^*$}
   \label{alg:qstar}
   \begin{algorithmic}
     \STATE\textbf{Input: } History $\hist^{t-1}$, random rollout $\rho_t$
     \STATE Calculate
     $ a_i := \frac{\gamma}{V_\infty} \min \left\{ \psi^0 - \psi_i^+, \psi_i^--\psi^0 \right\}$ and $b_i := \frac{\gamma}{V_\infty} \max \left\{ \psi^0 - \psi_i^+, \psi_i^--\psi^0 \right\}$, where     
     \begin{align*}
       \psi_i^+= \sup_{\pi\in\Pi}-\pi(x_t)^\top \frac{V_\infty e_i}{\gamma}
       +A_i(\pi),
       \psi_i^-= \sup_{\pi\in\Pi}\pi(x_t)^\top \frac{V_\infty e_i}{\gamma}
       +A_i(\pi), \text{ and }
       \psi_i^0= \sup_{\pi\in\Pi} A_i(\pi)
     \end{align*}
     for $A_i(\pi) = -\sum_{s=1}^{t-1} \pi(x_s)^\top D_i \hat \Delta_s - \sum_{s=t+1}^T 2\pi(x_s)^\top  \eepsilon_{s,i} \Z_{s}$.     
     \STATE Return the closest $q\in\triangle_N$ to $[a_1,b_1]\times\ldots\times[a_N,b_N]$ in $\Vert\cdot\Vert_1$ norm.
   \end{algorithmic}
\end{algorithm}

\section{Lower Bound}
\label{sec:lower.bound}
The previous algorithms only delivered fast $O(\sqrt{T})$ rates if pairwise observability holds. This section shows our upper bounds are tight: pairwise observability is necessary for the fast rate.
\begin{theorem} \label{thm:lower.bound}
  Consider a contextual partial monitoring game that is not pairwise observable. Then there exists a policy class and stochastic adversary such that any algorithm will incur expected regret w.r.t. the policy class and all fixed actions of at least
  \[
    \ex[ \regret_T] \geq C T^{2/3},
  \]
  where $C$ is some constant depending on $L$ and $H$ only.  
\end{theorem}
Such a lower bound cannot hold for arbitrary policy classes; if this were true, then picking the policy class of constant actions would contradict the lower bound for local observability.

The complete proof is presented in Appendix~\ref{sec:lower.bound.proof}, but the high level ideas are given here. The proof explicitly constructs a hard example. Let $x_t\sim\mathrm{Uniform}([0,1])$ (fortunately, we do not need to turn the distribution of $x_t$). Assume action 1 and 2 are not pairwise observable, and hence $\ell_1 -\ell_2\notin\im(S_1^\top)\oplus\im(S_2^\top)$. This implies that there some $v\in\ker(S_1)\cap\ker(S_2)$ with $(\ell_1-\ell_2)^\top v = 1$ (since we can scale $v$) because the kernel of a matrix $X$ is the orthogonal complement of $\im(X^\top)$. Also, $\ones\in\S_1^\top$, so $v^\top \ones = 0$.

Fix some $q_1\in C_1$ and $q_2\in\C_2$ and define $P_1(\epsilon) = (q_1 - \epsilon v)\mathds{1}\{x\leq \frac12\} + (q_2 - \epsilon v)\mathds{1}\{x >\frac12\}$ and $P_2(\epsilon)=P_1(-\epsilon)$. Under $P_1$, action 1 incurs slightly less loss and action 2 incurs slightly more. The key is that, for any $i, j\in \{1,2\}$, $S_i (q_j - \epsilon v) = S_i(q_j + \epsilon v)$, so the distribution of feedback symbols observed by the algorithm is exactly the same if actions 1 or 2 is played.

We define $\pi_1(x) = e_1 \mathds{1}\{x\leq \nicefrac12 + \beta_1\} + e_2 \mathds{1}\{x > \nicefrac12 + \beta_1\}$ and
$\pi_2(x) = e_1 \mathds{1}\{x\leq \nicefrac12 - \beta_2\} - e_2 \mathds{1}\{x > \nicefrac12 - \beta_2\}$. Policy $\pi_i$ has a bias towards action $i$ (playing it $2\beta$ more), and hence $\pi_i$ does better on $P_i$ by $O(\epsilon(\beta+1+\beta_2))$. We show that $\beta_1$ and $\beta_2$ can be tuned to a problem dependent constant such that either policy still outperforms all actions by a constant. The feedback structure ensures that any algorithm receives the same feedback distribution when following $\pi_1$ or $\pi_2$, and hence the algorithm must play other actions to determine which policy is better. By our construction, doing so will add constant regret.

The strategies $P_i$ are $O(\epsilon^2)$ apart in KL-divergence, which allows us to show that the strategies are  hard to separate given the feedback from any action. Hence, the learner must balance playing suboptimal actions with learning which $\pi_i$ is better. Setting $\epsilon=T^{-1/3}$ produces the lower bound.

%\section{Collaborative Filtering}
%\label{sec:CF}

\section{Conclusion}
This paper characterized the minimax regret for contextual partial monitoring for finite policy classes. We showed that pairwise observability is necessary and sufficient for the fast $O(\sqrt{T})$ rate. This result is surprising, since the noncontextual setting needs the significantly less strong notion of local observability for the $O(\sqrt{T})$ rate. Our lower bound implies that the relaxation algorithm is optimal for the local and global observability settings; this is the first known adversarial partial information setting where a relaxation algorithm obtains the minimax regret.

A few open problems remain. First, how does the complexity of the context affect the rate? Consider a game that is locally but not pairwise observable. If $x_t$ is a constant (or if the policy class ignores it), then the contextual case reduces to the noncontextual case and a fast rate is achievable. However, if $x_t\sim\mathrm{Uniform}([0,1])$, then we showed that the fast rate is impossible. Can one obtain lower and upper bounds in terms of the complexity of the context and interpolate between these two regimes? Second, is it possible to obtain $O(\sqrt{T})$ rates with a relaxation algorithm? The particular form of the optimal $q_t$ in Algorithm~\ref{alg:qstar} suggests a way forward by using properties of $\Pi$ to control $a$ from below and thereby controlling the importance weights without needing uniform exploration.

% \section{Experimental results}

% \begin{itemize}
% \item 10 items, 80 users. T=300 rounds. Used a scaler for the contexts and thresholded the ratings at 3.0 both to construct L and to provide feedback to the algorithm.
%   \begin{tabular}{|r|| l | l | l | l |}
%     Preprocessing proportion& Loss of partial monitoring& loss of uniform& loss of svd& loss of user/user\\
%     \hline\\
%     0.05000& PM=0.2334& unif=0.2100& SVD=0.2401& UU=0.2120\\
%     0.11250& PM=0.2247& unif=0.2515& SVD=0.2368& UU=0.2120\\
%     0.17500& PM=0.2308& unif=0.2301& SVD=0.1920& UU=0.1946\\
%     0.23750& PM=0.2401& unif=0.2548& SVD=0.1880& UU=0.2268
%   \end{tabular}
% \item 
% \end{itemize}

\newpage
\bibliography{../../big_bib}

\newpage
\appendix
\section{Proof of Theorem~\ref{thm:exp4.pm4.regret}}
\label{sec:ew.proof}
For this section, $\mathcal F_t$ denotes the filtration generated by $I_t$ and $x_t$, and $\ex_t[\cdot]:=\ex_{I_t}[\cdot|\mathcal F]$ is the conditional expectation over the player's actions. Our analysis will borrow the following theorem:
\begin{theorem}[\cite{abernethy2009beating}Theorem 2.1]
  \label{thm:ew.regret}
  The exponential weights algorithm using loss $\ell_t$, which plays $q_t(i) \propto \exp(-\eta\sum_{s=1}^{t-1} \ell_t(i))$ has regret
  \begin{equation}
\label{eqn:ew.regret}
  \sum_{t=1}^T (\ell_t)^\top (q_t - u)\leq \eta \sum_{t=1}^T (\Vert \ell_t\Vert_{q_t}^*)^2 + \frac{\log (K)}{\eta},
\end{equation}
where $\Vert \ell_t\Vert_{q_t}^*$ is the local norm defined by
$
  \Vert \ell_t\Vert_{q_t}^*
  =
  \sqrt{(\ell_t)^\top \diag(q_t) \ell_t}$
  and $u\in\triangle_N$ is any fixed distribution of actions.
\end{theorem}

\begin{proof}[Proof of Theorem~\ref{thm:exp4.pm4.regret}]
  For this proof, define $\Pi(x_t)$ to be the matrix with columns $\pi_1(x_t),\ldots, \pi_K(x_t)$, which allows us to write the expected loss from a distribution of policies $w_t \in \triangle_K$ as $\Delta_t^\top \Pi(x_t) w_t$ and the strategy of Algorithm~\ref{alg:exp4.pm} as $q_t = (1-N\gamma) \Pi(x_t) w_t + \gamma\ones$.

  Consider the exponential weights strategy that plays $w_t\propto \exp(- \eta\sum_{s=1}^{t-1}\Delta_s \Pi(x_s)^\top)$. The expectation of the guarantee from applying Theorem~\ref{thm:ew.regret} with $\ell_t = \Pi(x_t)^\top\hat\Delta_t$ is \begin{align*}
  \eta \sum_{t=1}^T \ex\left[(\Vert \Pi(x_t)^\top \hat\Delta_t  \Vert_{w_t}^*)^2\right] + \frac{\log (K)}{\eta}
  &\geq
  \ex\left[\sum_{t=1}^T(\Pi(x_t)^\top\hat\Delta_t)^\top (w_t - u) \right]\\
  &=
    \ex\left[\sum_{t=1}^T\ex_t\left[(\Pi(x_t)^\top\hat\Delta_t)^\top (w_t - u)
     \right]\right]\\
  &=
    \sum_{t=1}^T  \Delta_t^\top \Pi(x_t)w_t
    -\sum_{t=1}^T \Delta_t^\top \Pi(x_t)u\\
  &=
    \sum_{t=1}^T \e_{j_t}^\top (L - \ones\ell_{b_t}^\top)^\top\Pi(x_t) w_t
    -\sum_{t=1}^T \e_{j_t}^\top(L - \ones\ell_{b_t}^\top)^\top\Pi(x_t) u\\
  &=
    \sum_{t=1}^T \e_{j_t}^\top L^\top\Pi(x_t) w_t
    -\sum_{t=1}^T \e_{j_t}^\top L^\top\Pi(x_t) u.
\end{align*}
We can relate the last line to the loss of Algorithm~\ref{alg:exp4.pm} by the simple inequality
\begin{align*}  
  \sum_{t=1}^T e_{j_t}^\top L^\top \Pi(x_t)w_t
  -
  \sum_{t=1}^T e_{j_t}^\top L^\top  q_t
  &=
    \sum_{t=1}^T e_{j_t}^\top L^\top    
    (  N \gamma \Pi(x_t)^\top w_t - \gamma\ones )
    \leq
    \gamma T N.
\end{align*}
Combining the two inequalities above, we can show that
\begin{align}
  \ex[\Regret_T]
  &=
    \sum_{t=1}^T e_{j_t}^\top L^\top  q_t -  \min_u \sum_{t=1}^T \e_{j_t}^\top L^\top\Pi(x_t) u\notag\\
  &\leq
    \sum_{t=1}^T e_{j_t}^\top L^\top  \Pi(x_t) w_t
    - \min_u \sum_{t=1}^T \e_{j_t}^\top L^\top\Pi(x_t) u + \gamma T N\notag\\
    &\leq
      \eta \sum_{t=1}^T \ex\left[(\Vert \Pi(x_t)^\top \hat\Delta_t  \Vert_{w_t}^*)^2\right]
      + \frac{\log (K)}{\eta} + \gamma T N.
      \label{eqn:general.ew.regret}
\end{align}

The analysis diverges depending on whether pairwise observability holds. First, assume that it does not hold. To easy notation, define the matrix
\[
  V_t(k) := \left( v_{1,b_t,k}, \ldots,v_{N,b_t,k}\right)
\]
so that $\Delta_t = \sum_{k=1}^N V_t(k)^\top S_k e_{j_t}$ and $\hat\Delta_t = V_t(I_t)Y_t/q_t(I_t)$. The bound on the observability vectors implies that $V_\infty \geq \max_i\Vert V(i)\Vert_\infty$. We can bound the first term in \eqref{eqn:general.ew.regret} by
\begin{align*}
    \ex_t\left[(\Vert \Pi(x_t)^\top\hat\Delta_t\Vert_{w_t}^*)^2\right]
  &=
    \ex_t\left[\hat\Delta_t^\top\Pi(x_t) \diag(w_t)\Pi(x_t)^\top \hat\Delta_t \right]\\
  &=
    \ex_t\left[\sum_{k=1}^K w_t(k)(\hat\Delta_t^\top\pi_k(x_t))^2 \right]\\
  &=
    \ex_t\left[\sum_{k=1}^K w_t(k)\left(\frac{V(I_t)^\top\pi_k(x_t)}{q_t(I_t)}\right)^2\right]\\
    &\leq
      \ex_t\left[\frac{V_\infty V(I_t)^\top \Pi(x_t)w_t}{q_t(I_t)^2}\right]\\
  &\leq
    V_\infty^2 \gamma^{-1}.
\end{align*}
where, in the last inequality, we used $\Pi(x_t)w_t \in \triangle_N$ and $q_t \geq \gamma$. Combining with \eqref{eqn:general.ew.regret}, we have 
\begin{equation*}
  \ex[\regret_t] \leq
  \frac{\eta}{\gamma} T V_\infty^2  + \frac{\log (K)}{\eta} + \gamma T N.
\end{equation*}
Optimizing over the parameters and setting
$\eta = N^{-\frac13} \left(\frac{\log(K)}{V_\infty T}\right)^{\frac23}$ and
$\gamma = \left( \frac{ V_\infty^2 \log(K)}{N^2 T}\right)^{\frac13}$ yields
\[
  \ex[\Regret_T] \leq 3 \left(N V_\infty^2 \log(K)\right)^{\frac13} T^{\frac23}.
\]

Now, consider the pairwise observability case where $\gamma=0$. We may always choose $V_{i,j} = \{i,j\}$, and so the offset loss estimate will have support on $I_t$ and $b_t$ only. This implies that 
\[
  \hat\Delta_t(i) = \mathds{1}\{I_t = i\} \frac{v_{i,b,I_t}^\top Y_t}{q_t(I_t)}
  +
  \mathds{1}\{I_t = b_t\} \frac{v_{i,b_t,I_t}^\top Y_t}{q_t(I_t)}.
\]
Since $\pi_k(x_t)$ is a unit vector, we have $\mathds{1}\{ \pi_k(x_t) = I_t\} = \pi_k(x_t)^\top e_{I_t}$, so we can write
\[
  \pi_k(x_t)^\top \hat\Delta_t
  =
  \pi_k(x_t)^\top e_{I_t}
   \frac{v_{I_t,b_t,I_t}^\top Y_t}{q_t(I_t)}
  +
  \mathds{1}\{I_t = b_t\} \frac{v_{b_t,\pi_k(x_t),I_t}^\top Y_t}{q_t(I_t)}.
\]
The variance term can therefore be bounded by
\begin{align*}
  \ex_t\left[(\Vert \Pi(x_t)^\top\hat\Delta_t\Vert_{w_t}^*)^2\right]
  &=
  \ex \left[ \sum_{k=1}^K w_t(k)
    \left(
    \pi_k(x_t)^\top e_{I_t} \frac{v_{\pi_k(x_t),b_t,I_t}^\top Y_t}{q_t(I_t)}
    +
    \mathds{1}\{I_t = b_t\} \frac{v_{b_t,\pi_k(x_t),I_t}^\top Y_t}{q_t(b_t)}
    \right)^2\right]\\
  &\leq
    V_\infty^2
    \ex \left[ \sum_{k=1}^K w_t(k)
    \left(
    \frac{\pi_k(x_t)^\top e_{I_t} }{q_t(I_t)}
    +
    \frac{\mathds{1}\{I_t = b_t\} }{q_t(b_t)}
    \right)^2\right]\\
    &=
    V_\infty^2
    \ex \left[ \sum_{k=1}^K w_t(k)
    \left(
      \frac{\pi_k(x_t)^\top e_{I_t} }{q_t(I_t)^2}
      + 2
      \frac{\pi_k(x_t)^\top e_{I_t}\mathds{1}\{I_t = b_t\} }{q_t(I_t)q_t(b_t)}
    +
    \frac{\mathds{1}\{I_t = b_t\} }{q_t(b_t)^2}
      \right)\right]\\
  &\leq
    V_\infty^2
    \ex \left[
      \frac{q_t^\top e_{I_t} }{q_t(I_t)^2}
    +
    2 \frac{q_t^\top e_{I_t}\mathds{1}\{I_t = b_t\} }{q_t(I_t)q_t(b_t)}
    +  
    \frac{\mathds{1}\{I_t = b_t\} }{q_t(b_t)^2}
    \right]\\
  &\leq
    V_\infty^2
    \ex \left[
    \frac{q_t(I_t) }{q_t(I_t)^2}
    +
    2 \frac{q_t(I_t)\mathds{1}\{I_t = b_t\} }{q_t(I_t)q_t(b_t)}
    +  
    \frac{\mathds{1}\{I_t = b_t\} }{q_t(b_t)^2}
    \right]\\
  &=
    V_\infty^2\left(    
    3
    +  
    \frac{1 }{q_t(b_t)}
    \right).
\end{align*}
Since we choose $b_t = \argmax_i q_t(i)$, we must have $q_t(b_t)\geq 1/N$ and the previous term is bounded by $V_\infty^2(N+3)$. Combining this inequality with \eqref{eqn:general.ew.regret} and setting $\gamma=0$ produces
\begin{equation*}
  \ex[\regret_t] \leq
  \eta T V_\infty^2(N+3)
  + \frac{\log (K)}{\eta}.
\end{equation*}
The theorem statement follows from setting
\[
  \eta = \sqrt{ \frac{\log(K)}{ T V_\infty^2 (N+3)}}.
\]
\end{proof}

\section{Proofs for Relaxations}
\label{sec:rel.proofs}

\begin{lemma}
  \label{lem:complexity.bound}
  For any random vector $Z_t$ with $\ex[(Z_t^\top e_i)^2] \leq W$, $\epsilon_{t,i}$ i.i.d.\ Rademacher random variables, and $\eepsilon_t$ denoting the collection across $i$ of $\eepsilon_{t,i}=\diag(e_i \epsilon_{t,i})$,
  \begin{equation}
    \exop_{\Z_{1:T},\eepsilon_{1:T}}\left[\sum_{i=1}^N \sup_{\pi\in\Pi} \sum_{t=1}^T \pi(x_t)^\top \eepsilon_{t,i} \Z_t \right] \leq N\sqrt{2 T W \log( \vert \Pi\vert)}.
  \end{equation}
\end{lemma}
\begin{proof}%[Proof of Lemma~\ref{lem:complexity.bound}]
  This reasoning is a small refinement of the proof of Lemma 2 in \cite{syrgkanis2016improved}. We evaluate
%   \begin{align*}
%     \ex_{\Z_{1:T},\eepsilon_{1:T}}\left[ \sum_{i=1}^N\sup_{\pi\in\Pi} \sum_{t=1}^T \pi(x_t)^\top \eepsilon_{t,i} \Z_t \right]
%     &\leq
%       \ex_{\Z_{1:T},\eepsilon_{1:T}}\left[ \sup_{\pi\in\Pi} \sup_i\sum_{t=1}^T N\pi(x_t)^\top \eepsilon_{t,i} \Z_t \right]\\
%     &=
%       \ex_{\Z_{1:T}}\frac{1}{\lambda}\ex_{\eepsilon_{1:T}}\left[
%       \log  \left( 
%       \sup_{\pi\in\Pi}\sup_i
%       e^{\lambda \sum_{t=1}^T N \pi(x_t)^\top \eepsilon_{t,i} \Z_t} \right)\right]\\
%     &\leq
%       \ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
%       \log\left(
%       \ex_{\eepsilon_{1:T}}\left[
%       \sum_{\pi\in\Pi}\sum_{i=1}^N e^{\lambda \sum_{t=1}^T N\pi(x_t)^\top \eepsilon_{t,i} \Z_t}\right]\right)
%       \right]\\
%     &\leq
%       \ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
%       \log\left(\sum_{i=1}^N \sum_{\pi\in\Pi}\prod_{t=1}^T
%       \ex_{\eepsilon_{1:T}}\left[
%       e^{\lambda N \pi(x_t)^\top \eepsilon_{t,i} \Z_t}\right]\right)
%       \right].
%   \end{align*}
%   We have the inequality
%   \[
%     \ex_{\eepsilon_{t}}\left[ e^{\lambda N  \pi(x_t)^\top \eepsilon_{t,i} \Z_t}\right]
%     =
%     \ex_{\epsilon_t}\left[
%       e^{\lambda N (\pi(x_t)^\top e_i) \epsilon_t Z_{t,i}}
%     \right]
%     =
%     \frac{
%       e^{\lambda N (\pi(x_t)^\top e_i)  Z_{t,i}}
%       +
%       e^{-\lambda N (\pi(x_t)^\top e_i) Z_{t,i}}
%     }{2}
%     \leq
%     e^{\frac{N^2\lambda^2  Z_{t,i}^2}{2}}.
%   \]
% Combining with the above expression produces
%   \begin{align*}
%     \sum_{i=1}^N\ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
%     \log\left( \sum_{i=1}^N \sum_{\pi\in\Pi}\prod_{t=1}^T
%     \ex_{\eepsilon_{1:T}}\left[
%     e^{\lambda N\pi(x_t)^\top \eepsilon_{t,i} \Z_t}\right]\right)
%     \right]
%     &\leq
%       \ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
%       \log\left(\sum_{i=1}^N \sum_{\pi\in\Pi}\prod_{t=1}^T
%       e^{\frac{\lambda^2 N^2 Z_{t,i}^2 }{2}}
%       \right)
%       \right]\\
%     &\leq
%       \frac{1}{\lambda}\log( N \vert \Pi\vert )
%       +       
%       T\ex_{\Z}\left[
%       \frac{\lambda N^2 Z_{t_i}^2 }{2}
%       \right]\\    
%     &\leq
%       \frac{1}{\lambda}\log(\vert \Pi\vert)
%       +
%       \frac{N^2}{2}W T\lambda 
%   \end{align*}
%   Setting $\lambda = \sqrt{2\log(\vert \Pi\vert)/W T N^2}$ finishes the proof.

  \begin{align*}
    \sum_{i=1}^N\ex_{\Z_{1:T},\eepsilon_{1:T}}\left[ \sup_{\pi\in\Pi} \sum_{t=1}^T \pi(x_t)^\top \eepsilon_{t,i} \Z_t \right]
    &=
      \sum_{i=1}^N\ex_{\Z_{1:T}}\frac{1}{\lambda}\ex_{\eepsilon_{1:T}}\left[
      \log \left(
      \sup_{\pi\in\Pi}
      e^{\lambda \sum_{t=1}^T \pi(x_t)^\top \eepsilon_{t,i} \Z_t} \right)\right]\\
    &=
      \sum_{i=1}^N
      \ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
      \ex_{\eepsilon_{1:T}}\left[
      \log\left(
      \sup_{\pi\in\Pi}e^{\lambda \sum_{t=1}^T \pi(x_t)^\top \eepsilon_{t,i} \Z_t}\right)\right]
      \right]\\
    &\leq
      \sum_{i=1}^N 
      \ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
      \log\left(
      \ex_{\eepsilon_{1:T}}\left[
      \sum_{\pi\in\Pi}e^{\lambda \sum_{t=1}^T \pi(x_t)^\top \eepsilon_{t,i} \Z_t}\right]\right)
      \right]\\
    &\leq
       \sum_{i=1}^N 
      \ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
      \log\left(\sum_{\pi\in\Pi}\prod_{t=1}^T
      \ex_{\eepsilon_{1:T}}\left[
      e^{\lambda \pi(x_t)^\top \eepsilon_{t,i} \Z_t}\right]\right)
      \right].
  \end{align*}  
  We have the inequality
  \[
    \ex_{\eepsilon_{t}}\left[ e^{\lambda \pi(x_t)^\top \eepsilon_{t,i} \Z_t}\right]
    =
    \ex_{\epsilon_t}\left[
      e^{\lambda (\pi(x_t)^\top e_i) \epsilon_t Z_{t,i}}
    \right]
    =
    \frac{
      e^{\lambda (\pi(x_t)^\top e_i)  Z_{t,i}}
      +
      e^{-\lambda (\pi(x_t)^\top e_i) Z_{t,i}}
    }{2}
    \leq
    e^{\frac{1}{2}\lambda^2  Z_{t,i}^2}.
  \]
Combining with the above expression produces
  \begin{align*}
    \sum_{i=1}^N\ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
    \log\left( \sum_{\pi\in\Pi}\prod_{t=1}^T
    \ex_{\eepsilon_{1:T}}\left[
    e^{\lambda \pi(x_t)^\top \eepsilon_{t,i} \Z_t}\right]\right)
    \right]
    &\leq
      \sum_{i=1}^N\ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
      \log\left(\sum_{\pi\in\Pi}\prod_{t=1}^T
      e^{\frac{\lambda^2 Z_{t,i}^2 }{2}}
      \right)
      \right]\\
    &\leq
      N\frac{1}{\lambda}\log( \vert \Pi\vert )
      +       
      N T\lambda \ex_{\Z}\left[
      \frac{\lambda Z_{t_i}^2 }{2}
      \right]\\    
    &\leq
      \frac{N}{\lambda}\log(\vert \Pi\vert)
      +
      \frac{N}{2}W T\lambda 
  \end{align*}
  Setting $\lambda = \sqrt{2\log(\vert \Pi\vert)/W T}$ finishes the proof.
\end{proof}

\begin{proof}[Proof of Corollary~\ref{cor:oracle.regret.bound}]
  Lemma~\ref{lem:complexity.bound} with $W = V_\infty \gamma^{-1}$ yields a bound of
  \[
    \exop_{\Z_{1:T},\eepsilon_{1:T}}\left[\sum_{i=1}^N\sup_{\pi\in\Pi} \sum_{t=1}^T \pi(x_t)^\top \eepsilon_{t,i} \Z_t \right] \leq N \sqrt{2 T V_\infty \gamma^{-1} \log(\vert \Pi\vert)},
  \]
  which produces the theorem with the given value of $\gamma$.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:admissibility}]
  The base case is easy. Using the convexity of supremum and the unbiasedness of $\hat\Delta_t$,
  \begin{align*}
  \ex_{I_{1:T},\hat Z_{1:T}}\left[
  \rel(\hist^T)
  \right]
  &=
    \ex_{I_{1:T},\hat Z_{1:T}}\left[
    \sum_{i=1}^N\sup_{\pi\in\Pi}
    -\sum_{s=1}^T \pi(x_t)^\top D_i \hat \Delta_t
    \right]\\
  &\geq
    \ex_{I_{1:T},\hat Z_{1:T}}\left[
    \sup_{\pi\in\Pi}
    -\sum_{i=1}^N \sum_{s=1}^T \pi(x_t)^\top D_i \hat \Delta_t
    \right]\\
  &=
    \ex_{I_{1:T},\hat Z_{1:T}}\left[    
    \sup_{\pi\in\Pi}
    -\sum_{s=1}^T \pi(x_t)^\top \hat \Delta_t
    \right]\\
  &\geq
    \sup_{\pi\in\Pi}
    \sum_{s=1}^T
    \ex_{I_{1:T},\hat Z_{1:T}}\left[
    - \pi(x_t)^\top \hat \Delta_t
    \right]\\
  &=
    \sup_{\pi\in\Pi}
    -\sum_{s=1}^T \pi(x_t)^\top \Delta_t.
  \end{align*}

  We now check the inductive step. We define $\rho = (x_t,\eepsilon_t,\Z_t)_{t+1:T}$ to the collection of random variables in the relaxation. Recall that our aim is to prove admissibility of the strategy $q_t(\rho) = (1-N\gamma)q_t^*(\rho)+ \gamma \ones$ where $q_t^*(\rho)$ was defined by
  \begin{align*}
  q_t^*(\rho)
  &=
  \argmin_{q \in \triangle_{N}}
  \sup_{p_t \in \triangle_{\mathcal D}'} 
    \exop_{\hat\Delta_t \sim p_t}\left[
    q^\top\hat\Delta_t
    +  \sum_{i=1}^N R_i\left(\hist^{t},\rho\right)
    \right].
\end{align*}
By construction, no entry of $q_t(\rho)$ is below $\gamma$. It is then easy to check that, for $q_t^* = \ex_{\rho}[q_t^*(\rho)]$ and $q_t = \ex_{\rho}[q_t(\rho)]$,
\[
  \ex_{I_t \sim q_t}\left[ \Delta_t ^\top e_{I_t} \right] = \Delta_t^\top q_t  
  \leq \Delta_t^\top q_t^* + N\gamma
  =
  \ex_{I_t\sim q_t} \left[ \hat\Delta_t(I_t,j_t,q_t)^\top q_t^* \right] + N\gamma,
\]
where we have been explicit that $\hat\Delta_t(I_t,j_t,q_t)$ is a random variable that depends on the player actions and the $q_t$ (since $\Z_t$'s distribution is a function of $q_t$).

For every fixed $x_t$, we can apply the above inequality and unpack the definition of the relaxation to find
\begin{align*}
  \sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[e_{I_t}^\top\hat\Delta_t(I_t,j_t,q_t)
  +  \rel(\hist^t) \right]
  &=
    \sup_{j_t } \mathop{\ex}_{I_t\sim q_t}\left[e_{I_t}^\top\hat\Delta_t(I_t,j_t,q_t)
    +  \sum_{i=1}^N \exop_\rho\left[R_i\left(\hist^{t},\rho\right)\right]\right]\\
  &\quad + (T-t)N\gamma\\
  &\leq
    \sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[(q_t^*)^\top\hat\Delta_t(I_t,j_t,q_t)
    +
  \sum_{i=1}^N\exop_\rho\left[R_i\left(\hist^{t},\rho\right)\right]\right]\\
  &\quad+(T-t+1)N\gamma.
\end{align*}
Defining $A_i(\pi) = -\sum_{s=1}^{t-1} \pi(x_s)^\top D_i \hat \Delta_s - \sum_{s=t+1}^T 2\pi(x_s) \eepsilon_{s,i} \Z_{s}$, the previous line is equal to
\begin{align*}
\lefteqn{\sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[(q_t^*)^\top\hat\Delta_t(I_t,j_t,q_t)
    +
  \mathop{\ex}_{\rho}\left[\sum_{i=1}^N \sup_{\pi} - \pi(x_t)^\top \hat\Delta_t(I_t,j_t,q_t) + A_i(\pi)
  \right]\right] + (T-t+1)N\gamma}\\
  &\quad =
    \sup_{j_t} \sum_{i=1}^N\mathop{\ex}_{I_t\sim q_t}\left[\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top \hat \Delta_t (I_t,j_t,q_t)
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t(I_t,j_t,q_t) + A_i(\pi)\right]\right]\\
  &\quad+ (T-t+1)N\gamma.
\end{align*}

This optimization is intractable without strong assumptions on $\Pi$; the $j_t$ optimization needs to account for how the supremum over the policy class will be affected. Therefore, we reduce the constraints on the adversary to relax the problem by allowing play of distributions over $\hat\Delta_t$ instead of constraining the $\hat\Delta_t$ to correspond to 
a specific choice of $I_t,j_t$, and $q_t$. However, we carefully defined $\hat\Delta_t(I_t,j_t,q_t)$ so that it would have coordinate-wise sparseness and only expand the plays of the adversary to distributions with the same sparseness.

Specifically, every coordinate of $\hat\Delta_t(I_t,j_t,q_t)$  has large probability of being zero: if we fix $q_t$ by conditioning on $\rho$, then
\begin{align*}
  P\left(\hat\Delta_t(I_t,j_t,q_t)^\top e_i = 0 \right)
  &=
    \ex_{\rho}\left[ \ex_{I_t \sim q_t(\rho)}\left[ P\left(\hat\Delta_t(I_t,j_t,q_t)^\top e_i = 0 \middle |\rho, I_t\right)\right]\right]\\
    &=
    \ex_{\rho}\left[ \ex_{I_t \sim q_t(\rho)}\left[
    1-\frac{\gamma \vert e_i^\top V(I_t)^\top Y_t\vert }{V_\infty q_t(\rho)(I_t)}
    \right]\right]\\
  &\geq
    \ex_{\rho}\left[ \ex_{I_t \sim q_t(\rho)}\left[
    1-\frac{\gamma}{q_t(\rho)(I_t)}
    \right]\right]\\
  &= \ex_{\rho}\left[
    \sum_{i=1}^N q_t(\rho)(i) \left(1- \frac{\gamma}{q_t(\rho)(i)}\right)\right]\\
  &= 1 - N \gamma.
\end{align*}
Therefore, the distribution of $\hat\Delta_t(I_t,j_t,q_t)$ is always in $\triangle_{\mathcal D}'$, and we obtain an upper bound by allowing the adversary to play distributions over a random variable named $\hat\Delta_t$ with distribution in $\triangle_{\mathcal D}'$. With this substitution, $I_t$ and $j_t$ no longer appear in the expression. Suppressing the $(T-t+1)N\gamma$ term, we have
\begin{align*}
  &\lefteqn{
        \sup_{j_t } \sum_{i=1}^N\mathop{\ex}_{I_t\sim q_t}\left[\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top \hat \Delta_t(I_t,j_t,q_t)
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t(I_t,j_t,q_t) + A_i(\pi)\right]\right]
    }\\
  &\leq
    \sup_{p_t\in\triangle_{\mathcal D}'} \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[
    \sum_{i=1}^N\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top \hat \Delta_t
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]\right]\\
    &\leq
\mathop{\ex}_{\rho}\left[ \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^*(\rho)^\top \hat \Delta_t
    +
  \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]\right].
\end{align*}

Our ability to move the expectation over $\rho$ to the outside allows us to obtain the same bound in expectation by sampling a single $\rho$ and playing $q_t(\rho)$ instead of calculating the infimum over $q$.

For the remainder, fix some $\rho$. We defined
\[
  q_t^*(\rho)
  =
  \argmin_{q \in \triangle_{N}}
  \sup_{p_t \in \triangle_{\mathcal D}'}
  \ex_{\hat\Delta_t\sim p_t}\left[
  q^\top \hat \Delta_t
    +
    \sum_{i=1}^N \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right],
  \]
  and so
  \begin{align*}
    &\lefteqn{\sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^*(\rho)^\top \hat \Delta_t
    +
  \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]}\\
    &=
      \inf_{q_t}
      \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top \hat \Delta_t
    +
  \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right].
\end{align*}
% Recall that this expression is an upper bound on 
% $  \sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[ e_{I_t}^\top \hat \Delta_t(I_t,j_t,q_t)
%    + \rel(\hist^t)\right] -(t+1)N\gamma$.

We continue bounding this saddle point problem from above. Noting that the objective is linear in both $q_t$ and $p_t$, we may perform a min-max swap:
\begin{align*}
  &\lefteqn{
          \inf_{q_t}
      \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top \hat \Delta_t
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]
    }\\
  &=
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \inf_{q_t}    
    \sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top \hat \Delta_t
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]\\
  &\leq
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \inf_{j}  
    \sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j^\top \hat \Delta_t
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]\\
  &=
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N
    \inf_{j}  \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j^\top \hat \Delta_t\right]
    +
\mathop{\ex}_{\hat\Delta_t\sim p_t}\left[  \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]. 
\end{align*}
Because $\pi$ is deterministic, we must have
\[
  \min_i \ex_{\hat\Delta_t \sim p_t}
  [e_i^\top\hat\Delta_t]
  \leq
  \ex_{\hat\Delta_t \sim p_t }[\pi(x_t)^\top\hat\Delta_t]
  \leq
  \ex_{\hat\Delta_t \sim p_t }[\pi(x_t)^\top D_i\hat\Delta_t],
\]
 which allows us to upper bound the previous expression. Performing the usual symmeterization (using $\epsilon$ as a single Rademacher random variable) yields
\begin{align*}
  &\lefteqn{
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t}\left[
     \sup_{\pi\in\Pi} A_i(\pi)
    +\min_j \ex_{\hat\Delta_t'\sim p_t}\left[e_j^\top \hat\Delta_t'\right]
    -\pi(x_t)^\top D_i \hat\Delta_t
  \right]
  }\\
  &\leq
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t}\left[
     \sup_{\pi\in\Pi} A_i(\pi)
    +\ex_{\hat\Delta_t'\sim p_t}\left[\pi(x_t)^\top D_i \hat\Delta_t'\right]
    -\pi(x_t)^\top D_i \hat\Delta_t
  \right]\\
  &\leq
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t, \hat\Delta_t'\sim p_t}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        \pi(x_t)^\top D_i\left(\hat\Delta_t'-\hat\Delta_t\right)
    \right]\\
  &=
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t, \hat\Delta_t'\sim p_t,\epsilon_i}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        \epsilon_i\pi(x_t)^\top D_i\left(\hat\Delta_t'-\hat\Delta_t\right)
    \right]\\
    &\leq
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t,\epsilon_i}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\epsilon_i\pi(x_t)^\top D_i\hat\Delta_t
        \right].
\end{align*}
Since $D_i\hat\Delta_t = e_i \hat\Delta_t(i)$, each term in the sum only involves one coordinate if $\hat\Delta_t$. Therefore, it is without loss of generality to assume that $p_t$ is a product distribution. Using $\triangle_{\{0,V_\infty \gamma^{-1}\}}$ to denote distributions over the singletons $0$ and $V_\infty\gamma^{-1}$, define 
\[
  \triangle_1' := \left\{ p \in \triangle_{\{0,V_\infty \gamma^{-1}\}}: p(0)\geq 1-N\gamma\right\}
\]
and observe that if $\hat\Delta_t\sim p_t\in\triangle_{\mathcal D'}$, then $\epsilon_i X_i e_t \stackrel{\mathcal L}{=} \hat\Delta_t(i)$ for some $X_i\sim p_i\in\triangle_1'$. We then have 
\begin{align*}
  \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t,\epsilon}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\epsilon\pi(x_t)^\top D_i \hat\Delta_t
  \right]
  &=
      \sup_{p_1,\ldots,p_N \in\triangle_1'}
    \sum_{i=1}^N \ex_{X_i\sim p_i,\epsilon_{1:N}}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\pi(x_t)^\top e_i \epsilon_i X_i
    \right]\\
      &=
        \sum_{i=1}^N
        \sup_{p_i \in\triangle_1'}
        \ex_{X_i\sim p_i,\epsilon_i}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\pi(x_t)^\top e_i \epsilon_i X_i
    \right].
\end{align*}

We will now argue that a witness to the supremum of 
\begin{align*}
    \sup_{p_i\in\triangle_1'}
  \ex_{X_i\sim p_i,\epsilon_i}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\pi(x_t)^\top e_i \epsilon_i X_i
        \right]
\end{align*}
is the distribution that puts mass $N\gamma$ on $V_\infty\gamma^{-1}$ and the rest on $0$. Define the convex function 
$g(x) :=  \sup_{\pi\in\Pi} A_i(\pi)+ 2\pi(x_t)^\top e_i x$. The expectation $\ex_{\epsilon}\left[ g(\epsilon x)\right]$ is increasing on $x\geq 0$. To see this, consider some $0\leq a<b$. We can write $a = \theta b + (1-\theta)(-b)$ for some $\theta$ and use the definition of convexity to conclude 
 \begin{align*}
   \ex_{\epsilon}\left[ g(\epsilon a) \right]
   = 
   \frac{g(a)+g(-a)}{2}
   \leq
   \frac{\theta g(b)+(1-\theta)g(-b)
   +(1-\theta) g(b)+\theta g(-b)
   }{2}
   =
   \ex_{\epsilon}\left[ g(\epsilon b) \right].
 \end{align*}
 Since $\ex_{\epsilon}\left[ g(\epsilon x)\right]$ is increasing, the supremum of $p_i$ puts maximum mass on $V_\infty\gamma^{-1}$. Hence, defining the random vector $Z_{t}$ with elements $P(Z_{t,i} = 0) = 1-N\gamma$ and $P( Z_{t,i} = V_\infty\gamma^{-1}) = N\gamma$, we have
 \begin{align*}
    \sup_{p\in\triangle_1'}
  \ex_{X_i\sim p,\epsilon_i}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\pi(x_t)^\top e_i \epsilon_i X_i
        \right]
   &=
  \ex_{\epsilon_i}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\epsilon_i\pi(x_t)^\top D_i \Z_t
   \right]\\
      &=
  \ex_{\eepsilon_t}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\pi(x_t)^\top \eepsilon_{t,i} \Z_t
        \right].
\end{align*}

In total, we have shown that, for every fixed $x_t$ and $\rho$, playing $q_t(\rho)$ allows the bound
\begin{align*}
    \sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[ e_{I_t}^\top \Delta_t
  + \rel(\hist^t)\right]
  &\leq
    \ex_\rho \left[
  \ex_{\epsilon_t, \Z_t}\left[
  \sum_{i=1}^N
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\pi(x_t)^\top \eepsilon_{t,i} \Z_t
  \right]
  \right]
  + (T-t+1)N\gamma\\
  &=
  \rel(\hist^{t-1}),
\end{align*}
as required.
\end{proof}

\section{Proof of Lemma~\ref{lem:computation}}
\label{sec:computation.appendix}
\begin{proof}
  Recall that the relaxation algorithm sampled $\rho_t$ then calculated the $q_t(\rho)$ that minimized
 \begin{align*}
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N
    \exop_{\hat\Delta_t(i) \sim p_i}\left[
    q^\top D_i \hat\Delta_t
    +
    \sup_{\pi\in\Pi}-\pi(x_t)^\top D_i \hat\Delta_t
    +A_i(\pi)\right],
 \end{align*}
where $A_i(\pi) = -\sum_{s=1}^{t-1} \pi(x_s)^\top D_i \hat \Delta_s - \sum_{s=t+1}^T 2\pi(x_s) \eepsilon_{s,i} \Z_{s}$ and the $(T-t)N\gamma$ term is suppressed. This optimization decomposes over coordinates of $\hat\Delta_t$ and therefore over marginals of $p_t$. Since $\triangle_{\mathcal D}'$ only puts support on vectors with coordinates in $\{-V_\infty \gamma^{-1},0,V_\infty \gamma^{-1}\}$,  we can fully parameterize the problem with  $p_i^+ := p_i(\cdot = V_\infty \gamma^{-1})$ and $p_i^- = p_i(\cdot =- V_\infty \gamma^{-1})$ for $i=1,\ldots,N$. The definition of $\triangle_{\mathcal D}'$ implies that $p_i^+\leq\gamma$ and $\pi_i^-\leq\gamma$. The objective becomes
 \begin{align*}
\lefteqn{    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N
    \exop_{\hat\Delta_t(i) \sim p_i}\left[
    q^\top D_i \hat\Delta_t
    +
    \sup_{\pi\in\Pi}-\pi(x_t)^\top D_i \hat\Delta_t
    +A_i(\pi)\right]
}\\
&=
  \sum_{i=1}^N
    \sup_{p_i^+\leq\gamma, p_i^- \leq\gamma}
    \frac{V_\infty q_i}{\gamma} (p_i^+-p_i^-)
    +
    \exop_{\hat\Delta_t(i) \sim p_i}\left[
    \sup_{\pi\in\Pi}-\pi(x_t)^\top e_i\hat\Delta_t(i)
    +A_i(\pi_i)
  \right]\\
&=
  \sum_{i=1}^N
    \sup_{p_i^+\leq\gamma, p_i^- \leq\gamma}
    \frac{V_\infty q_i}{\gamma} (p_i^+-p_i^-)
  +
  p_i^+ \left( \sup_{\pi\in\Pi} -\pi(x_t)^\top e_i\frac{V_\infty}{\gamma}
  +A_i(\pi_i)\right)\\
&\quad +
  p_i^- \left( \sup_{\pi\in\Pi}  \pi(x_t)^\top e_i\frac{V_\infty}{\gamma}
  +A_i(\pi_i)\right)
  +
  (1-p_i^+-p_i^-) \left( \sup_{\pi\in\Pi} A_i(\pi_i)\right),
 \end{align*}
and we can immediately see that we only require invoking the oracle $3N$ times to evaluate
\begin{align*}
  \psi_i^+:= \sup_{\pi\in\Pi}-\pi(x_t)^\top \frac{V_\infty e_i}{\gamma}
  +A_i(\pi),
  \psi_i^-:= \sup_{\pi\in\Pi}\pi(x_t)^\top \frac{V_\infty e_i}{\gamma}
  +A_i(\pi), \text{ and }
  \psi_i^0:= \sup_{\pi\in\Pi} A_i(\pi)
\end{align*}
for $i=1,\ldots, N$. In terms of these quantities, the objective becomes
\begin{align}
  \sum_{i=1}^N
      \sup_{p_i^+\leq\gamma, p_i^- \leq\gamma}
  \left(
  p_i^+\left(\psi_i^+ + \frac{V_\infty q_i}{\gamma} - \psi_i^0\right)
  +
    p_i^-\left(\psi_i^- - \frac{V_\infty q_i}{\gamma} - \psi_i^0\right)
  + \psi_i^0
     \right).
       \label{eqn:sup.with.p}
\end{align}

Since $p_i^+$ and $p_i^-$ are bounded by $\gamma$, the supremum will be at 
$
  p_i^+
  =
  \gamma \mathds{1}\left\{\left(\psi_i^+ + \frac{V_\infty q_i}{\gamma}\right) > \psi_i^0\right\}
  $
  and
  $
  p_i^-
  =
  \gamma \mathds{1}\left\{\left(\psi_i^- - \frac{V_\infty q_i}{\gamma}\right) > \psi_i^0\right\}
$.
Hence, \eqref{eqn:sup.with.p} evaluates to 
\begin{align*}  
  N\psi_0
  + \sum_{i=1}^N
  \max \left\{-\gamma(\psi_i^0-\psi_i^+) + V_\infty q_i, 0\right\}
  +
  \max \left\{\gamma(\psi_i^--\psi_i^0) - V_\infty q_i, 0\right\}.
\end{align*}

The positivist of $\pi(x_t)$ ensures that $\psi_i^+ \leq \psi_i^0 \leq \psi_i^-$ which implies that $(\psi^0-\psi_i^+)\geq 0$ and $(\psi_i^--\psi^0)\geq 0$. Since the $\max$ switches at $\frac{\gamma}{V_\infty}(\psi^0 - \psi_i^+)$ and the second at $\frac{\gamma}{V_\infty}(\psi_i^--\psi^0)$, the minimizer of $q_i$ is between these two values where the slope vanishes. If
$\psi_i^--\psi^0 \leq \psi^0 - \psi_i^+$,
then the minimum is $0$; otherwise, it is $\gamma (\psi_i^++\psi_i^- -2\psi_0)$.

By defining $a_i := \frac{\gamma}{V_\infty} \min \left\{ \psi^0 - \psi_i^+, \psi_i^--\psi^0 \right\}$ and $ b_i := \frac{\gamma}{V_\infty} \max \left\{ \psi^0 - \psi_i^+, \psi_i^--\psi^0 \right\}$, we can compactly write the objective as 
\[
  \sum_{i=1}^N
  V_\infty\max\{q_i-a_i,0\}
  +
  V_\infty\max\{b_i-q_i,0\}
  +
  \max\{ \gamma (\psi_i^++\psi_i^- -2\psi_0), 0\}.
\]
Let $A_t$ be the rectangle of $\Reals^N$ with $i$th coordinate in $[a_i, b_i]$. If $A_t$ has  has nonempty intersection with $\triangle_N$, then any point in the intersection is optimal. Otherwise, we can exploit the fact that the objective value at $a_i - \epsilon$ is exactly the same as the objective value at $b_i + \epsilon$. Hence, the value of any point $x$ is the $L_1$ distance to $A_t$.

  There are three cases to consider. First, if $\sum_i a_i \leq 1 \leq \sum_i b_i$, then any $q$ with $\sum_i q_i = 1$ and $q_i \in [a_i,b_i]$ is optimal. In particular, we can select the $q$ with by a constrained water-filling algorithm as follows. Define $q(x)$ to have coordinates
\[
    q_i(x) =
    \begin{cases}
      a_i & \text{ if } x \leq a_i, \\
      x & \text{ if } a_i \leq x \leq b_i, \text{ and }\\
      b_i & \text{ if } x \leq b_i.
    \end{cases}
\]
Select $q^* = q(x_{fill})$ where $x_{fill} := \max\{x : \sum_i q_i(x) \leq 1\}$. Because $\sum_i a_i \leq 1 \leq \sum_i b_i$, such an $x_{fill}$ must exist. We can find $x_{fill}$ easily since $\sum_i q_i(x)$ is a piecewise linear increasing function in $x$ with at most $2N$ points where the slope changes.

The second case is $\sum_i b_i < 1$, which implies that $A_t$ does not intersect $\triangle_N$. In this case, the suboptimality is exactly $V_\infty \Vert q - b\Vert_1$, which can be minimized a water-filling algorithm as described above with no upper limit to the coordinates of $q_i$. The final case is $\sum_i a_i > 1$, which results in an inverse water-filling algorithm.

% Finally, in the case when $\sum_i a_i > 1$, we need to preform a reverse water-filling. Analogously define $g_v(y) = \sum_i \min\{ \max_j v_j - y, v_i\}$. We can write this as
% \begin{align*}
%   g_v(y)
%   &=
%     \sum_i \min\{ \max_j v_j - y, v_i\}\\
%   &=
%     -\sum_i \max \{ y- \max_j v_j , - v_i\}\\
%   &=
%     N(\max_j v_j)-\sum_i \max \{ y, \max_j v_j  -v_i\}\\
%   &=
%     N(\max_j v_j) - f_{(\max_j v_j  -v)}( y - \min_j v_j).
% \end{align*}
% We wish to solve for the $y$ such that $g_a(y) = 1$, which implies that $y^* = f_{(\max_j a_j  -a)}^{-1}( N \max_j a_j-1) + \min_j a_j$, so
% $q_i^* = \min\{ \max_j a_j - y^*, a_i\}$. 
\end{proof}

\section{Proof of Lower Bound}
\label{sec:lower.bound.proof}
For readability, we break up the proof into four sections.

\subsection{Defining the alternatives}  
By assumption, there exist two non-degenerate actions that are not pairwise observable. Without loss of generality, assume that these are actions 1 and 2. Define $S_{1,2} = \begin{bmatrix} S_1 \\ S_2
\end{bmatrix}$. Since actions 1 and 2 are not pairwise observable, we must have $\ell_1 - \ell_2 \notin \im( S_{1,2}^\top)$. Let $w$ be the orthogonal projection of $\ell_1 - \ell_2$ onto $\im( S_{1,2}^\top)$. Since $\Reals^M = \im(S_{1,2}^\top) \oplus \ker (S_{1,2})$, we must have $w\in\ker (S_{1,2})$. Let $v$ be a scaling of $w$ such that $(\ell_2 - \ell_1)^\top v = 1$, and note that  $\ones\in\im(S_{1,2})$ implies that $v^\top \ones = 0$ (where $\ones$ is the all ones vector). To summarize, there exists a vector $v$ with:
\begin{enumerate}
\item $v^\top \ones = 0$,
\item $(\ell_1 - \ell_2)^\top v = 1$, and
\item $S_1v = S_2 v = 0$.
\end{enumerate}

We now exploit the existence of such a $v$ to design two adversary strategies, $P_1$ and $P_2$, and a policy class $\Pi$, such that $P_1$ and $P_2$ are difficult to distinguish when playing in $\Pi$ but the excess loss of playing outside of $\Pi$ is large.

Recalling that $x_t\sim\mathrm{Uniform}([0,1])$, we will choose an adversary strategy $P_1$ that plays $q_1\in C_1$ when  $x\leq \beta$ and $q_2 \in C_2$ when $x>\beta$ for some constant $\beta\in [0,1]$ we will optimize later. Strategy $P_2$ is analogously defined but for distributions $q_3 \in C_1$ and $q_4 \in C_2$. That is,
\begin{align*}
    P_1 &= q_1 \mathds{1}\{ x \leq \beta\} + q_2 \mathds{1}\{ x > \beta\}\\
    P_2 &= q_3 \mathds{1}\{ x \leq \beta\} + q_4 \mathds{1}\{ x > \beta\}.
\end{align*}
In particular, for some $\epsilon>0$, $q_a \in C_1$ and $q_b \in C_2$ (which will be described shortly), the four distribution are 
\begin{equation}
  q_1 = q_a - \epsilon v,
  q_2 = q_b - \epsilon v,
  q_3 = q_a + \epsilon v,\text{ and }
  q_4 = q_b + \epsilon v.
\end{equation}
This construction ensures that the feedback distribution of $q_1$ and $q_2$ is the same when playing action $1$ or $2$ (and is the same when following policies $\pi_1$ or $\pi_2$), but the two policies will get different losses when they disagree because $(\ell_1 - \ell_2)^\top v \neq 0$. We will choose $q_a$ and $q_b$ to maximize this difference.

To be precise, define
$Q_1(q) := \max \{\epsilon'\geq 0: q+\epsilon' v \in C_1, q - \epsilon'v \in C_1\}$
and
$Q_2(q) := \max \{\epsilon'\geq 0: q+\epsilon' v \in C_2, q-\epsilon' v \in C_2 \}$
and choose $q_a \in\argmax_q Q_1(q)$ and $q_b \in \argmax_q Q_2(q)$. The maximum $\epsilon$ translation with $q_1,q_3\in C_1$ and $q_2,q_4\in C_2$ is $\epsilon_0 = \max\{ Q_1(q_a),Q_2(q_b)\}$. Because $C_1$ and $C_2$ are convex, $q_1,q_3\in C_1$ and $q_2,q_4\in C_2$ for all $\epsilon \leq \epsilon_0$.

Next, we create a policy class $\Pi$ that is able to use the context to do better than any fixed action. We will consider the threshold policies 
\begin{align*}
  \pi_1(x) &= e_1 \mathds{1}\{ x \leq \beta + \beta_1\} + e_2 \mathds{1}\{ x > \beta+\beta_1\} \text{ and }\\
  \pi_2(x) &= e_1 \mathds{1}\{ x \leq \beta - \beta_2\} + e_2 \mathds{1}\{ x > \beta-\beta_2\}
\end{align*}
for constants $0 < \beta_1 \leq 1- \beta$ and $0 < \beta_2 \leq \beta$, which we will set later. As we can see, both policies use the context to track the optimal action given $x_t$, but do so with some error. Policy $\pi_1$ slightly favors action $1$ and strategy $P_1$ plays distributions which gives lower loss to 1 over action 2. The situation is reversed for $\pi_2$ and $P_2$. Our construction ensures that $\pi_1$ holds a slight edge over $\pi_2$ under $P_1$. We will quantify the difference in expected losses in the next section.

\subsection{Calculating the loss differences}
In the sequel, we will use $\ex_1$ to denote the expectation over context $X$ and adversary action $J\sim P_1|X$. We analogously use $\ex_2$ for expectations over $X$ and $J\sim P_2|X$.

We now show that is possible to set $\beta$, $\beta_1$, and $\beta_2$ so that expected loss difference between $\pi_1$ and $\pi_2$ is $O(\epsilon)$, but the loss of either policy is a constant better than any fixed action. We denote the expected loss of playing action $i$ and following policy $\pi_i$ under strategy $P_j$ by $\ell_i^j:=\ex_{J\sim P_j}[e_i^\top L e_J]$ and $\ell_{\pi_i}^j:=\ex_{J\sim P_j}[\pi_i(x)^\top Le_J]$, respectively. 
\begin{lemma}\label{lemma:expected.loss.bounds}
  There exists $\beta$, $\beta_1$, $\beta_2$ and some $\epsilon_0'$ such that, for all $\epsilon\leq\epsilon_0'$, the following inequalities are simultaneously satisfied:
  \begin{align*}
    \ell_i^1 - \ell_{\pi_1}^1 &\geq c_1, \\
    \ell_i^2 - \ell_{\pi_2}^2 &\geq c_1, \\
    \ell_{\pi_2}^1 - \ell_{\pi_1}^1 &\geq c_2 \epsilon,\\
    \ell_{\pi_1}^2 - \ell_{\pi_2}^2 &\geq c_2 \epsilon,
  \end{align*}
  for some constants $c_1>0$ and $c_2>0$ that depend only on the structure of the game. 
\end{lemma}
\begin{proof}
An easy calculation yields
\begin{align*}
  \ell_{\pi_1}^1 &= \ell_1^\top q_1 (\beta - \beta_2) + \ell_1^\top q_1 \beta_2
          + \ell_1^\top q_2 \beta_1 + \ell_2^\top q_2 (1-\beta-\beta_1),\text{ and }\\
  \ell_{\pi_2}^1 &= \ell_1^\top q_1 (\beta - \beta_2) + \ell_2^\top q_1 \beta_2
          + \ell_2^\top q_2 \beta_1 + \ell_2^\top q_2 (1-\beta-\beta_1).
\end{align*}
Thus,
\begin{align*}
  \ell_{\pi_2}^1 - \ell_{\pi_1}^1
  &=
    \ell_2^\top q_1 \beta_2
    + \ell_2^\top q_2 \beta_1
    -\ell_1^\top q_1 \beta_2
    - \ell_1^\top q_2 \beta_1\\
  &=
    \beta_2(\ell_2 - \ell_1)^\top q_1 
    - \beta_1(\ell_1-\ell_2)^\top q_2 \\
  &=
    \beta_2(\ell_2 - \ell_1)^\top q_a
    -\beta_2 \epsilon (\ell_2 - \ell_1)^\top v
    - \beta_1(\ell_1-\ell_2)^\top q_b
    + \beta_1 \epsilon  (\ell_1-\ell_2)^\top v\\
  &=
    \beta_2(\ell_2 - \ell_1)^\top q_a
    - \beta_1(\ell_1-\ell_2)^\top q_b
    + (\beta_1+\beta_2) \epsilon,
\end{align*}
where $(\ell_2 - \ell_1)^\top q_a$ and $(\ell_1-\ell_2)^\top q_b$ are both positive constants. For readability, we will define the constants $\delta_{1\rightarrow i} := (\ell_i - \ell_1)^\top q_a>0$ and $\delta_{2 \rightarrow i} := (\ell_i - \ell_2)^\top q_b>0$, where the first quantity is  the excess loss of playing action $i$ instead of action $1$ under $q_a$ (where action $1$ is optimal) and the second is the analogous quantity for $q_b$. Hence, under $P_1$, the excess loss of following policy $\pi_2$ instead of the optimal $\pi_1$ is 
\begin{align*}
  \ell_{\pi_2}^1 - \ell_{\pi_1}^1
  &=
    \beta_2\delta_{1\rightarrow 2} 
    -
    \beta_1 \delta_{2\rightarrow 1}
    + (\beta_1 + \beta_2)\epsilon .
\end{align*}
The calculation for $\ell_{\pi_2}^2 - \ell_{\pi_1}^2$ is quite similar:
\begin{align*}
  \ell_{\pi_1}^2 &= \ell_1^\top q_3 (\beta - \beta_2) + \ell_1^\top q_3 \beta_2
          + \ell_1^\top q_4 \beta_1 + \ell_2^\top q_4 (1-\beta-\beta_1) \text{ and }\\
  \ell_{\pi_2}^2 &= \ell_1^\top q_3 (\beta - \beta_2) + \ell_2^\top q_3 \beta_2
          + \ell_2^\top q_4 \beta_1 + \ell_2^\top q_4 (1-\beta-\beta_1),
\end{align*}
and so we conclude that
\begin{align*}
\ell_{\pi_1}^2 - \ell_{\pi_2}^2
  &= 
    \beta_1(\ell_1-\ell_2)^\top q_4 
    -
    \beta_2(\ell_2 - \ell_1)^\top q_3\\
  &=
    \beta_1(\ell_1-\ell_2)^\top q_b
    +
    \epsilon \beta_1 (\ell_1-\ell_2)^\top v
    -
    \beta_2(\ell_2 - \ell_1)^\top q_a 
    - \epsilon  \beta_2(\ell_2 - \ell_1)^\top v\\
  &=
    \beta_1\delta_{2\rightarrow 1}
    -
    \beta_2\delta_{1\rightarrow 2}
    + (\beta_1+\beta_2)\epsilon .
\end{align*}

We also need to evaluate $\ell_i^j - \ell_{\pi_1}^j$ for $j=1,2$. It is easy to calculate $\ell_i^1 = \beta \ell_i^\top q_1 + (1-\beta)\ell_i^\top q_2$, and therefore,
\begin{align*}
    \ell_i^1 - \ell_{\pi_1}^1
  &=
    \beta \ell_i^\top q_1 + (1-\beta)\ell_i^\top q_2 - \beta\ell_1^\top q_1
    - (1-\beta - \beta_1)\ell_2^\top q_2- \beta_1 \ell_1^\top q_2\\
  &=
    \beta (\ell_i - \ell_1)^\top q_1 + (1-\beta)(\ell_i - \ell_2)^\top q_2
    + \beta_1( \ell_2-\ell_1)^\top q_2\\
  &=
    \beta \delta_{1\rightarrow i}
    + (1-\beta)\delta_{2 \rightarrow i}
    - \beta_1 \delta_{2\rightarrow 1}
    -\epsilon \left(
    \beta (\ell_i - \ell_1)^\top v  
    +(1-\beta)(\ell_i - \ell_2)^\top v
    -\beta_1 \right)\\
  &=
    \beta \delta_{1\rightarrow i}+ (1-\beta)\delta_{2 \rightarrow i} - \beta_1 \delta_{2\rightarrow 1}
    +\epsilon\left(\beta+\beta_1
    -(\ell_i - \ell_2)^\top v\right)\\
    &=
    \beta \delta_{1\rightarrow i}+ (1-\beta)\delta_{2 \rightarrow i} - \beta_1 \delta_{2\rightarrow 1}
    +\epsilon\left(\beta+\beta_1
    -\frac{1+\left(2\ell_i - \ell_2-\ell_1\right)^\top v}{2}\right),
\end{align*}
where the last line used $(\ell_i - \ell_2)^\top v = (\ell_i - \ell_2)^\top v - \frac12(\ell_1 - \ell_2)^\top v+ \frac12 = (\ell_i - (\ell_1 - \ell_2)/2)^\top v + \frac12$.
Similarly, the $j=2$ case is
\begin{align*}
    \ell_i^2 - \ell_{\pi_2}^2
  &=
    \beta \ell_i^\top q_3 + (1-\beta)\ell_i^\top q_4 - \beta_2\ell_2^\top q_3
    - (1-\beta)\ell_2^\top q_4- (1-\beta -\beta_2) \ell_1^\top q_3\\
  &=
    \beta (\ell_i - \ell_1)^\top q_3 + (1-\beta)(\ell_i - \ell_2)^\top q_4
    + \beta_2( \ell_1 - \ell_2)^\top q_3\\
  &=
    \beta \delta_{1\rightarrow i}
    + (1-\beta)\delta_{2\rightarrow i}
    - \beta_2 \delta_{1\rightarrow 2}
    + \epsilon\left(\beta (\ell_i - \ell_1)^\top v + (1-\beta)(\ell_i - \ell_2)^\top v + \beta_2\right)\\
    &=
    \beta \delta_{1\rightarrow i}
    + (1-\beta)\delta_{2\rightarrow i}
      - \beta_2 \delta_{1\rightarrow 2}
      +\epsilon\left(      
       (\ell_i - \ell_2)^\top v  - (\beta-\beta_2)
      \right)\\
  &=
    \beta \delta_{1\rightarrow i}
    + (1-\beta)\delta_{2\rightarrow i}
      - \beta_2 \delta_{1\rightarrow 2}
      +\epsilon\left(         
    \frac{1+\left(2\ell_i - \ell_2-\ell_1\right)^\top v}{2}
    - (\beta-\beta_2)
      \right).
\end{align*}

% The last two cases can be deduced by combining the equalities above:
% \begin{align*}
%   \ell_1^1 - \ell_{\pi_2}^1
%   &=
%     \ell_1^1 - \ell_{\pi_1}^1 + (\ell_{\pi_1}^1 - \ell_{\pi_2}^1) \\
%   &= (1-\beta)\delta_{2\rightarrow 1} - \beta_1\delta_{1\rightarrow 2} -\epsilon( 1 - (\beta-\beta_2)),\text{ and }\\
%   \ell_2^2 - \ell_{\pi_1}^2
%   &=
%     \ell_2^2 - \ell_{\pi_2}^2 + (\ell_{\pi_2}^2 - \ell_{\pi_1}^2)\\
%   &=
%     \beta\delta_{1\rightarrow 2} - \beta_2\delta_{2\rightarrow 1} - \epsilon(\beta + \beta_1).
% \end{align*}

We now optimize for $\beta$, $\beta_1$, and $\beta_2$. First, it suffices to take $\beta = \frac12$, which lets us bound
\begin{align*}
    \ell_i^1 - \ell_{\pi_1}^1
  &=
    \frac{\delta_{1\rightarrow i}+\delta_{2 \rightarrow i}}{2} - \beta_1 \delta_{2\rightarrow 1}
    +\epsilon\left(\beta_1
    -\frac{\left(2\ell_i - \ell_2-\ell_1\right)^\top v}{2}\right) \\
&\geq
    \frac{\delta_{1\rightarrow i}+\delta_{2 \rightarrow i}}{2}
    -\beta_1\delta_{2\rightarrow 1}
    -\frac{\epsilon}{2}
    \left\vert \left(2\ell_i - \ell_2-\ell_1\right)^\top v\right\vert \text{ and }\\
  \ell_i^2 - \ell_{\pi_2}^2 
  &=
    \frac{\delta_{1\rightarrow i}+\delta_{2 \rightarrow i}}{2}
      - \beta_2 \delta_{1\rightarrow 2}
      +\epsilon\left( \beta_2 
    +\frac{\left(2\ell_i - \ell_2-\ell_1\right)^\top v}{2}
    \right)\\
          &\geq
          \frac{\delta_{1\rightarrow i}+\delta_{2 \rightarrow i}}{2}
      - \beta_2 \delta_{1\rightarrow 2}
      -\frac{\epsilon}{2}
        \left\vert \left(2\ell_i - \ell_2-\ell_1\right)^\top v\right\vert.
\end{align*}

We define the two constants 
\[
c_3 = \max_i \left\vert \left(2\ell_i - \ell_2-\ell_1\right)^\top v\right\vert \text{ and } c_1 = \min_i\frac{ \delta_{1\rightarrow i}+\delta_{2\rightarrow i}}{4},
\]
where the latter is strictly positive; the non-degeneracy of action 1 implies that  $\delta_{1\rightarrow i}>0$ for all $i\neq 1$, and the non-degeneracy of action 2 implies that $\delta_{2\rightarrow i}>0$ for all $i\neq 2$. Combining $\epsilon \leq \epsilon_0$ with these constants give the bounds
\begin{align*}
  \ell_i^1 - \ell_{\pi_1}^1
  &\geq
  2c_1  - \beta_1\delta_{2\rightarrow 1} - \epsilon_0 c_3
\text{ and}\\
  \ell_i^2 - \ell_{\pi_2}^2
  &\geq
  2c_1 - \beta_2\delta_{1\rightarrow 2} - \epsilon_0 c_3.
\end{align*}
Therefore, under the assumption that $\epsilon_0 < c_1/c_3$, choosing $\beta_1 \leq \frac{ c_1 - \epsilon_0 c_3}{\delta_{2\rightarrow 1}}$ and $\beta_2 \leq \frac{ c_1 - \epsilon_0 c_3}{\delta_{1\rightarrow 2}}$ yields the first two inequalities.

% We can lower bound the cross terms by 
% \begin{align*}
%   \ell_1^1 - \ell_{\pi_2}^1
%   &\geq
%     \frac{\delta_{2\rightarrow 1}}{2} - \beta_1\delta_{1\rightarrow 2} -\frac{\epsilon}{2}
%     \geq
%     c_4 - \beta_1\delta_{1\rightarrow 2} -\frac{\epsilon_0}{2}
%     \text{ and }\\
%   \ell_2^2 - \ell_{\pi_1}^2
%   &\geq
%     \frac{\delta_{1\rightarrow 2}}{2} - \beta_2\delta_{2\rightarrow 1} -\frac{\epsilon}{2}
%   \geq
%     c_4 - \beta_2\delta_{2\rightarrow 1} -\frac{\epsilon_0}{2},
% \end{align*}
% and so it suffices to take 
% $\beta_1 \leq \frac{ c_4 - \epsilon_0}{2\delta_{1\rightarrow 2}}$ and $\beta_2 \leq \frac{ c_4 - \epsilon_0}{2\delta_{2\rightarrow 1}}$
% to guarantee that $  \ell_1^1 - \ell_{\pi_2}^1 \geq c_4/4$ and $\ell_2^2 - \ell_{\pi_1}^2\geq c_4/4$.

We now turn to the latter two inequalities and notice that setting $\beta_2 =  \beta_1 \frac{\delta_{2\rightarrow 1}}{\delta_{1\rightarrow 2}}$ implies that 
\[
  \ell_{\pi_2}^1 - \ell_{\pi_1}^1 = \ell_{\pi_1}^2-\ell_{\pi_2}^2 = \epsilon \beta_1 \left( 1 + \frac{\delta_{2\rightarrow 1}}{\delta_{1\rightarrow 2}}\right).
\]
Therefore, we choose  $\beta_1 =\frac{ c_1 - \epsilon_0 c_3}{2\delta_{2\rightarrow 1}}$ and
$\beta_2 = \frac{ c_1 - \epsilon_0 c_3}{2\delta_{1\rightarrow 2}}$, which satisfies the first two inequalities, has $\beta_2 =  \beta_1 \frac{\delta_{2\rightarrow 1}}{\delta_{1\rightarrow 2}}$, and therefore implies that
\[
  \ell_{\pi_2}^1 - \ell_{\pi_1}^1 = \ell_{\pi_1}^2-\ell_{\pi_2}^2 \geq \frac{\epsilon}{2} ( c_1 - \epsilon_0 c_3) \left( \frac{1}{\delta_{2\rightarrow 1}} + 
    \frac{1}{\delta_{1\rightarrow 2}}\right).
\]
That is, we may take $c_2 = \frac12( c_1 - \epsilon_0 c_3) \left( \frac{1}{\delta_{2\rightarrow 1}} + 
  \frac{1}{\delta_{1\rightarrow 2}}\right)$
and long as $\epsilon_0 \leq c_4/(2c_3)$, which we may assume (since $\epsilon_0$ is a parameter we control as well).
\end{proof}

\subsection{Bounding the KL-Divergence}
At a high level, any randomized algorithm must make similar decisions when given similar data. In our setting, the algorithm observes the contexts, which do not depend on the stategy of the adversary, as well as the feedback symbols. An important step in the argument is to lower bound the difference in feedback distributions as a function of the expected number of plays of different actions. As is standard, we will use the KL-divergence between distribution $p$ and $q$, denoted by $\KL(p\Vert q)$, to quantify differences in distributions.

Denote the symbol received at round $t$ by $f_t\in\Sigma$. Let $P_j^*(\cdot| f_{1:t-1},x_{1:t})$ be the mass function over $\Sigma$ at round $t$ generated by the algorithm's choices if the adversary uses strategy $P_j$, and let $N_i$ be the number of times the algorithm plays action $i$.
\begin{lemma}\label{lemma:feedback.kl}
  For all sufficiently small $\epsilon$, the relative entropy between the feedback symbols of strategy $P_1$ and $P_2$ has the upper bound
  \begin{equation}
    \KL(P_1^*||P_2^*)
  \leq
      \sum_{i>2} \ex_1[N_i] c_5\epsilon^2,
\end{equation}
where $c_5$ is some game-dependent constant. 
\end{lemma}
\begin{proof}
  Fix some algorithm $\mathcal A$ which maps the information available for selecting $I_t$, denoted $H^t := f_{1:t-1},x_{1:t}$.  We can apply the KL-divergence chain rule $T$ times to conclude
  \begin{align*}
   \KL(P_2^*||P_1^*)
    &=
      \sum_{t=1}^{T-1}\sum_{f_{1:t-1}}P_2^*(f_{1:t-1}) \sum_{f_t} P_2^*(f_t|H^t)
      \log \frac{ P_2^*(f_t|H^t)}{ P_1^*(f_t|H^t)}\\
    &=
      \sum_{t=1}^{T-1}\sum_{f_{1:t-1}}P_2^*(f_{1:t-1})
      \sum_{i=1}^N\mathds{1}\left\{ \mathcal A(H^t) = i \right\}
      \sum_{f_t} P_2^*(f_t|H^t)
      \log \frac{ P_2^*(f_t|H^t)}{ P_1^*(f_t|H^t)}.
  \end{align*}

By a slight abuse of notation, define $S_i P_j$ to be the distribution of feedback symbols if action $i$ is played under strategy $P_j$ with $x_t$ marginalized out, i.e. $S_i P_1 = \beta S_i q_1 + (1-\beta) S_i q_2$ and $S_i P_2 = \beta S_i q_3 + (1-\beta) S_i q_4$. However, because $v\in\ker(S_{1,2})$, we have that $S_i P_1 = S_i P_2$ for $i=1,2$. Hence, the $\mathcal A(H^t) \in\{ 1, 2\}$ terms in the summation vanish. We can then evaluate
\begin{align*}
    \KL(P_1^*||P_2^*)
    &=
      \sum_{t=1}^{T-1}\sum_{f_{1:t-1}}P_1^*(f_{1:t-1})
      \sum_{i>2}^N\mathds{1}\left\{ \mathcal A(H^t) = i \right\}
      \sum_{f_t} P_1^*(f_t|H^t)
      \log \frac{ P_1^*(f_t|H^t)}{ P_2^*(f_t|H^t)}\\
    &\leq
      \sum_{i>2} \ex_1[N_i] \KL( S_i P_1\Vert S_i P_2)\\
    &\leq
      \sum_{i>2} \ex_1[N_i] \KL( P_1\Vert P_2).
\end{align*}

Thus, we may turn a bound on $\KL(P_1\Vert P_2)$ into a bound on $\KL(P_1^*\Vert P_2^*)$, so let us consider the first quantity. We will briefly be explicit about the $X$ dependence. Let $P_j(J,X)$, for $j=1,2$, denote the joint distribution of the context $X\sim \mathrm{Uniform}([0,1])$ and adversary choice $J\sim P_j|X$. The chain rule yields
\begin{align*}
    \KL(P_1(J,X)\Vert P_2(J,X))
  &=
    \KL(P_1(X)\Vert P_2(X)) +
    \KL(P_1(J|X)\Vert P_2(J|X))\\
  &=
    0 +
    \beta \KL(q_1\Vert q_3) + (1-\beta) \KL(q_2\Vert q_4).
\end{align*}
To bound each term, recall that $q_1 = q_a - \epsilon v$ and $q_3 = q_a + \epsilon v$ and apply Lemma~12 from \cite{bartok2011minimax}, which implies that, for all $\epsilon$ small enough and some positive constants $c_1'$ and $c_2'$,
\[
  \KL(q_a-\epsilon v || q_a + \epsilon v) \leq c_1' \epsilon^2\Vert v\Vert_\infty^2
  \text{ and }
  \KL(q_b-\epsilon v || q_b + \epsilon v) \leq c_2' \epsilon^2\Vert v\Vert_\infty^2.
\]

Thus, our total relative entropy bound is
\begin{equation}
    \KL(P_1^*||P_2^*)
  \leq
      \sum_{i>2} \ex_1[N_i] c_5\epsilon^2
\end{equation}
with $c_5 = \max\{ c_1', c_2'\}\Vert v \Vert_\infty^2$ and for all $\epsilon$ small enough.. 
\end{proof}
The next tool we need is a way to translate relative entropy bounds into high probability bounds: a high-probability version of Pinsker's inequality.
\begin{lemma}\cite[Lemma 2.6]{Tsybakov-2008-book}
For probability distribution $P$ and $Q$ with $P\ll Q$,
  \[
    \int \min \{ dP, dQ \} \geq \frac12 \exp(- \KL(P \Vert Q)).
    \]
\end{lemma}
In particular, for some event $A$ with $P(A) \leq Q(A)$, integrating the indicator function of $A$ and $A^c$ yields $P(A) \geq \frac12 \exp(- \KL(P \Vert Q))$ and $Q(A^c)\geq \frac12 \exp(- \KL(P \Vert Q))$. In the case of $P(A) > Q(A)$, the same argument applied to $A^c$ yields the same conclusion, which implies that
\begin{equation}\label{eqn:pinsker.high.prob}
  P(A) + Q(A^c)  \geq \exp(- \KL(P \Vert Q)).
\end{equation}

\subsection{Finalizing the Bound}
We are now ready to prove a lower bound on regret, and we begin by exploiting the construction of the game which allows  the expected regret to be easily calculated from the action counts. Let $N_{\pi_i}(T)$ be the number of times algorithm $\mathcal A$ chooses policy $\pi_i$ over a game of length $T$, which is a random variable dependent on the context and the random choices of the algorithm and adversary. We will also define $N_0(T)$ to be the number of times an $\mathcal A$ chooses any action $i>2$.

We assume that the strategy uses some $\beta,\beta_1$, and $\beta_2$ that satisfy Lemma~\ref{lemma:expected.loss.bounds} and that $\epsilon$ is small enough such that Lemmas~\ref{lemma:expected.loss.bounds} and \ref{lemma:feedback.kl} hold. Under $P_1$, the optimal policy is $\pi_1$, playing policy $\pi_2$ incurs at least $c_2 \epsilon$ more expected loss, and playing any other action incurs at least $c_1$ more expected loss. The following construction is modeled after the slick proof of \citet{lattimore2018cleaning}. We can lower bound the expected regret under $P_1$ by
\begin{align*}
  \ex_1[\regret_T]
  &\geq
    \ex_1\left[ \sum_{t=1}^T\sum_{i=1}^N
    \mathds{1}\left\{ \mathcal A(H^t) = i \right\}
    (\ell_i - \ell_{\pi_1(x_t)})^\top J_t
    \right]\\
  &\geq
    \ex_1\left[ \sum_{t=1}^T\sum_{i=3}^N
    \mathds{1}\left\{ \mathcal A(H^t) = i \right\}
    (\ell_i - \ell_{\pi_1(x_t)})^\top J_t\right]\\
    &\quad +\ex_1\left[\sum_{t=1}^T
    \mathds{1}\left\{ \mathcal A(H^t) = \pi_2(x_t) \right\}    
    (\ell_{\pi_2(x_t)} - \ell_{\pi_1(x_t)})^\top J_t
    \right]\\
  &\geq
    c_1\ex_1\left[ \sum_{t=1}^T\mathds{1}\left\{ \mathcal A(H^t) > 2 \right\}
    \right]
    +
    \epsilon c_2 \ex_1\left[\sum_{t=1}^T\mathds{1}\left\{ \mathcal A(H^t) = \pi_2(x_t) \right\}\right]\\
  &=
    c_1\ex_1\left[N_0(T)\right]
    +
    \epsilon c_2 \ex_1\left[N_2(T)\right]\\
  &\geq
    c_1\ex_1\left[N_0(T)\right]
    + \frac{\epsilon c_2 T}{2}
    P_1\left(N_2(T)\geq \frac{T}{2}\right).
\end{align*}
Similarly, we can bound the expected regret under $P_2$ by 
\begin{align*}
  \ex_2[\regret_T]
  &\geq
    \ex_2\left[ \sum_{t=1}^T
    \sum_{i=1}^N
    \mathds{1}\left\{ \mathcal A(H^t) = i \right\}    
    (\ell_i - \ell_{\pi_2(x_t)})^\top J_t
    \right]\\
  &\geq
    c_1\ex_2\left[ \sum_{t=1}^T\mathds{1}\left\{ \mathcal A(H^t) > 2 \right\}
    \right]
    +
    \epsilon c_2 \ex_2\left[\sum_{t=1}^T\mathds{1}\left\{ \mathcal A(H^t) = \pi_1(x_t) \right\}\right]\\
  &\geq
    \min\{ c_1,\epsilon c_2\}
    \ex_2\left[N_0(T) + N_1(T)\right]\\
  &\geq
    \min\{ c_1,\epsilon c_2\}\frac{T}{2}
    P_2\left(N_0(T) + N_1(T)>\frac{T}{2}\right)\\
  &\geq
    \min\{ c_1,\epsilon c_2\}\frac{T}{2}
    P_2\left(N_2(T)\leq\frac{T}{2}\right),
\end{align*}
with the last line following from $N_2(T) + N_1(T) + N_0(T) \leq T$.
Note that  we simply dropped the terms where the algorithms plays action $1$ or $2$ but in disagreement with both policies.

The final adversary strategy will be a uniform mixture between $P_1$ and $P_2$, and, under the assumption that $\epsilon\leq c_1/c_2$,
\begin{equation}
  \label{eqn:regret.as.indicators}
    \ex[\regret_T]
  =
    \frac{\ex_1[\regret_T] + \ex_2[\regret_T]}{2}
  \geq
    \ex_1\left[N_0(T)\right]c_1
    + \frac{\epsilon c_2 T}{2}
    \left(
    P_1\left(N_2(T)\geq \frac{T}{2}\right)
    +    
    P_2\left(N_2(T)\leq\frac{T}{2}\right)
    \right).
  \end{equation}

We can finally assemble all the ingredients into a lower bound proof.
\begin{proof}[Proof of Theorem~\ref{thm:lower.bound}]
  First, by definition, $N_0(T) = \sum_{i>2}^N N_i(T)$. Now, combining the regret lower bound from \eqref{eqn:regret.as.indicators} with \eqref{eqn:pinsker.high.prob} and Lemma~\ref{lemma:feedback.kl} yields
\begin{align*}   
  \ex[\regret_T]
&\geq
    \ex_1\left[N_0(T)\right]c_1
    + \frac{\epsilon c_2 T}{2}
  \exp(-\KL(P_1\Vert P_2) )\\
&\geq
      \ex_1\left[N_0(T)\right]c_1
    + \frac{\epsilon c_2 T}{2}
    \exp(- \ex_1[N_0(T)] c_5 \epsilon^2 ).
\end{align*}
If $\ex_1\left[N_0(T)\right] \geq T^{2/3}$, then the desired bound follows immediately. Otherwise, setting $\epsilon = T^{-\frac13}$ yields 
\begin{align*}   
  \ex[\regret_T]
&\geq
    \ex_1\left[N_0(T)\right]c_1
    + \frac{\epsilon c_2 T}{2}
  \exp(-\KL(P_1\Vert P_2) )\\
&\geq
      \ex_1\left[N_0(T)\right]c_1
    +  T^{\frac23}\frac{c_2}{2}
    \exp(- \ex_1[N_0(T)] c_5 T^{-\frac23} )
\end{align*}
with the term in the exponential approaching 0. Since $\epsilon$ is a decreasing sequence, all the assumptions that $\epsilon$ is smaller than certain quantities will eventually be satisfied.
\end{proof}

\iffalse
\paragraph{A detail}\todo{remove this paragraph}
While the regret bound, proven w.r.t. the two policies $\pi_1$ and $\pi_2$ and all fixed actions, 
In particular, there exist a policy that is better that $\pi_1$ and $\pi_2$: the one that switches exactly correctly at $\beta$. To fix this, we can instead define
\[
  P_1 = q_1 \mathds{1}\{x \leq \beta-\beta_2\} + q_2 \mathds{1}\{x \geq \beta+\beta_1\}
  + \frac{q_1+q_2}{2}\mathds{1}\{\beta-\beta_2\leq x \leq \beta+\beta_1\}
\]
and analogously for $P_2$. Then the ``perfect policy'' from before does not do any better than $\pi_1$ or $\pi_2$ in expectation.

Unfortunately, this set of alternatives does not have the nice property that the feedback distribution of playing $\pi_1$ under $P_1$ and $P_2$ is identical; there is some small change from $S_1 q_a$ to $S_1 q_b$. It might be possible to balance terms and control the regret.
\fi

\end{document}

### 2018-09-07 pap_old.tex
Initial ShareLaTeX Import

\documentclass[11pt]{article}

\usepackage{fullpage}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables
\usepackage{bm}
\usepackage{amsfonts,dsfont}
\usepackage{comment}
\usepackage{mathtools,amsthm}

\bibliographystyle{plainnat}

\usepackage{xcolor}
\usepackage{tikz}
\usepackage[colorinlistoftodos, textwidth=26mm, shadow,color=blue!30!white]{todonotes}
%\usepackage[disable]{todonotes}

\usepackage{algorithmic, algorithm}
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

\newcommand{\KL}{\mathrm{KL}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\exop}{\mathop{\ex}}
\newcommand{\proj}{\mathrm{proj}}
\newcommand{\regret}{\mathcal{R}}
\newcommand{\spn}{\mathrm{span}}
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\rel}{\mathbf{Rel}}
\newcommand{\hist}{{\mathcal H}}
\newcommand{\im}{{\mathrm{Im}\;}}

\newtheorem{assumption}{Assumption}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{corollary}{Corollary}
\newtheorem{theorem}{Theorem}

\input{../def}

%\renewcommand{\inf}{\vphantom{\sup}\inf}

\begin{document}

%---Uncomment until abstract 
\title{Online Collaborative Filtering with Partial Monitoring}

\author{
  Alan Malek
  \\MIT
  \texttt{amalek@mit.edu}
  \and
  Alexander Rakhlin
  \\MIT
  \and
  Ali Jadbabaie\\
  MIT
}

\begin{abstract}
We formulate and solve the partial monitoring with context problem. To demonstrate its effectiveness, we apply the method to online collaborative filtering.
\end{abstract}

\section{Upper bound for very easy games}

The condition for fast rates in contextual partial monitoring is pair-wise observability:
\begin{definition}
  A contextual partial monitoring game is pairwise obvervable if every pair of actions is locally observable. That is, for all $i$ and $j$, there exists a vector $v_{i,j}$ such that 
  \[
    \ell_i - \ell_j =
    \begin{bmatrix}
    S_i & S_j 
    \end{bmatrix}^\top 
    v_{i,j}.
  \]
\end{definition}

In this section, we prove that in games where every pair of actions is locally observable, there is a $\sqrt{T}$ algorithm. To make the ideas clearer, we describe the feedback in a slightly different way, but the overall idea is that we will use importance weighted samples for the loss differences. However, instead of controlling how small the importance weights can get with uniform exploration (which precluded the fast $O(\sqrt{T})$ rate), we will use the pair-wise observable structure instead.

\subsection{Specific Example}
It is a good example to work out a specific example then see how to generalize. We will begin with a game with three actions with locally observable pairs $(1,3)$ and $(2,3)$.

We could check that the label efficient prediction satisfies this, but here is a concrete example:
\[
  S_1 = S_2 =
  \begin{bmatrix}
    1&0&0\\
    0&1&1
  \end{bmatrix},
  S_3 =
  \begin{bmatrix}
    1&1&0\\
    0&0&1
  \end{bmatrix},
  \ell_i = e_1, \ell_2 = e_2,
  \text{ and } \ell_3 = (\frac34, \frac34, \frac12)^\top.
\]
The cells are defined by
\[
  C_1 = \{ p: p_1 \geq 3 p_2 + 4 p_3\},
  C_2 = \{ p: p_2 \geq 3 p_1 + 4 p_3\},
  \text{ and } C_3 = \triangle_3 \setminus (C_1 \cup C_2).
\]
Since $C_3$ is symmetric, it is easy to check that $C_3 = \{(p,1-p,q)^\top:q\in[0,1], p\in[1/4-q,3/4+q]\}$. Hence, $C_1$ and $C_2$ are not neighbors.

We also have the necessary observability conditions since $\mathcal R( S_1, S_2, S_3) = \Reals^3$ but $\ell_1 - \ell_2 = (1,-1,0)^\top \notin \mathcal R(S_1)$. Does the confoundability condition hold? Well, $\ker S_1 = \{(0,q,-q):q\in\Reals\}$, so it seems that $p_A - p_B \notin \ker S_1$ for any $p_A\in C_1$ and $p_B \in C_2$.

Let's try 
\[
  S_3 = 
  \begin{bmatrix}
    1&0&0\\
    0&1&1
  \end{bmatrix},
  S_1 = S_2 =
  \begin{bmatrix}
    1&1&0\\
    0&0&1
  \end{bmatrix}.
\]
Then $\ker S_1 = \{(q,-q,0):q\in\Reals\}$, which means we need the first two coordinates of $p_A$ and $p_B$ to be mirrored, which also fixes the third coordinate; hence, $p_A = (a,b,c)$ and $p_B = (b,a,c)$. It seems it would suffice to take $a=.8,b=.1$.

Finally, checking that $\ell_1 - \ell_2 \notin \im S_1^\top$, we see that the confoundable condition is not trivial. However, the first definition of $S_i$ also had the necessary neighborhood structure but was not confoundable, so this definition is a weakening.

\section{Introduction}

Second, we demonstrate the applicability of our formalism on the very practical problem of online collaborative filtering. Each round, a user with context $x_t$ arrives and the learner needs to select one item from a catalog of $N$ items (e.g. movies) to show the user. The user has some natural affinity towards the item, but the learner only sees some coarse reward, such as a ``like/not-like'' rating. Our goal, a departure from previous works, is to maximize the quality of recommendations and not just the number of positive ratings. This problem fits naturally into partial monitoring, which allows us to handle the unobservability of the true loss. We note that the use of context is essential for solving this problem;  without it, every incoming user will be treated identically and our algorithm will merely find the single best item.

We assume that items are associated with a feature space $\mathcal F\subseteq \Reals^d$ that could be composed of any item features and aggregated rating of the items. We will allow for these features to be updated as more information is received, and define $L$ and $H$ as follows. Using $w_i$ for the feature vector of item  $i$, define $L_{i,j} = w_i^\top w_j$ and $H_{i,j} = \mathds{1}\{L_{i,j} \leq \gamma\}$. The policy class $\pi$ naturally maps contexts $x\in\mathcal X$ to item recommendations, and the loss matrix measures how close our item recommendation is to the best recommendation we could have made. What the partial monitoring formulation adds is the ability to infer the best item matches from the coarse feedback from $H_{i,j}$. Intuitively, the role of the adversary is to establish where in $\mathcal F$ the user's preferences lie.

The contextual partial monitoring algorithms we present can be applied to finite policy classes or classes with a value oracle. The regret bounds guarantee that if $\Pi$ contains a good mapping from contexts to items, then our predictions will match its performance. For concreteness, we will evaluate, both theoretically and empirically, the policy class
that reccomends an item proportional to $M x_t$ for some $M \in \left\{ \Reals^{N \times d}
    \Vert M\Vert_2 \leq R, \rank(M) \leq r
  \right\}$,
where $d$ is the dimension of the context space.

\subsection{Related work }
Partial monitoring, without context, was first formalized in \cite{piccolboni2001discrete} and non-trivial regret bounds were derived. Optimal bounds for the stochastic case were established in \cite{bartok2011minimax}, and it was shown that the regret falls into four categories: trivial games with 0 regret, ``easy'' games with a local-observability condition with $\tilde\Theta(\sqrt{T})$ regret, ``hard'' games with a weaker global-observability condition with $\Theta(T^{2/3})$ regret, and finally impossible games with no observability condition and linear regret. Algorithms with matching lower and upper bounds were exhibited for all categories. Surprisingly, the exactly same boundaries also exist under oblivious adversarial data \cite{foster2012no}. We are the first to extend partial monitoring to the contextual case.

Our algorithms build on several well established online learning algorithms. Our algorithm for finite policy class and adversarial contexts is very similar to the \textsc{EXP4} \cite{auer2002exp4} algorithm, and our algorithm for i.i.d. contexts with a policy oracle builds on the relaxation algorithms, first proposed in \cite{rakhlinrelax}, then extended to the contextual bandit case with a $O(T^{2/3})$ regret (for a finite policy class) in \cite{rakhlin2016bistro,syrgkanis2016improved}. It therefore seems natural that these relaxation based methods would be able to yield an optimal algorithm for the contextual partial monitoring case.

Collaborative filtering is a well established problem in machine learning and used in many practical systems like Amazon's item recommendation \cite{linden2003amazon} and GroupLens news recommendation system \cite{resnick1994grouplens} and, despite that its natural setting is online, theoretical guarantees of online collaborative filtering are much more recent \cite{bresler2014latent,bresler2016collaborative}. \todo{expand}

\subsection{Contributions}
We formulate the contextual partial monitoring problem and provide two optimal algorithms for solving it. The first, a variant of \textsc{EXP4}\cite{auer2002exp4}, requires keeping weights over the policies and hence is not suitable for large $\Pi$. However, the algorithm makes no assumptions about the context sequence or $j_t$; i.e. they can both be adversarial.

The second method is a relaxation-based method \cite{rakhlinrelax} using a random roll-out of future data and is similar to the corresponding methods for the contextual bandit problem \cite{rakhlin2016bistro,syrgkanis2016improved}. These methods require the ability to sample  contexts, but interface with the policy class only through a value-oracle. That is, we assume that for any sequence of contexts $x_1',\ldots, x_s'$ and losses $\ell_1,\ldots,\ell_s$, we can evaluate
$\min_{\pi\in\Pi} \sum_{t=1}^s \pi(x_t)^\top \ell_t$.
In particular, the policy class need not be bounded and the algorithm only makes a small number of oracle calls.

For both algorithms, we prove a $O(T^{2/3})$ upper bound on regret, matching the lower bound in the non-contextual partial monitoring setting \cite{cesa2006regret}.

Our main application will be to recommendation systems. Using partial monitoring, we are able to describe an algorithm that learns fine preferences for user contexts despite receiving coarse feedback. While our framework can provide regret bounds for arbitrary item-recommendation classes, we explicitly evaluate a natural linear class.

% \subsection{Notation}

% $p$ for adversary strategies, $q$ for player strategies. 
% We will use $\mathcal N_i$ to denote both the set actions neighboring action $i$ (not inclusive of $i$) as well as the set of edges between these actions and action $i$; which definition will be clear from context.

% Player actions, rows, and graph nodes will typically be indexed by $i$, $i'$, $I_t$ (for random quantities), etc. Caligraphic capital letters will denote subsets of player actions, e.g. $\mathcal N_i$ for the neighborhood around action $i$. Adversary actions and columns will be indexed by $j$.

% The closest element of a subset $\mathcal S\subseteq\underbar N$ to some action $i$ is denoted $\proj(i,\mathcal S) = \argmin_{i'\in\mathcal S} d(i,i')$. This element is unique if $|\mathcal S \cap \mathcal N_i| \leq 1$.

\section{Partial Monitoring with Context}
In this section, we formulate the partial monitoring with context problem, propose two algorithms with regret bounds, and provide a matching lower bound. We begin by reviewing the  observability structure, which is necessary and sufficient for learning in partial monitoring; we closely follow \cite{bartok2014partial}.

\subsection{Finite policy class, arbitrary contexts}
We momentarily concentrate on the setting where  $\Pi = \{\pi_1,\ldots,\pi_K\}$ is finite and $x_t$ and $j_t$ are arbitrary. A finite $\Pi$ allows us to use unbiased estimates of the losses of every policy an input to a full information algorithm analogous to \textsc{EXP4}  \cite{auer2002exp4}. Let $\Pi(x_t)$ be the matrix with columns $\pi_1(x_t),\ldots, \pi_K(x_t)$. It is easy to see that $\Pi(x_t)^\top \hat\Delta_t$ is an unbiased estimate of the offset losses for every policy: $\Pi(x_t)$ is independent of $I_t$ and hence $\Pi(x_t)^\top\hat\Delta_t$ is a linear function of an unbiased estimate, hence unbiased.

\subsection{Offline oracle, i.i.d. contexts}

\subsection{Matching Lower Bound}
Lower bounds for the non-contextual problem were investigated in prior works. The strongest is from \cite{bartok2011minimax}, which we paraphrase below.
\begin{theorem}
  Let $L,H$ be a globally observable game that is not locally observable. Then there exists a game-dependent constant $c$, policy class $\Pi$, distribution of contexts and stochastic adversary strategy such that, for any deterministic algorithm, $\ex[\Regret_T] \geq c T^{\frac23}$.
\end{theorem}
This theorem is a trivial application of \cite{bartok2011minimax}[Theorem~9] with $\Pi = \{e_1,\ldots,e_N\}$.

However, there is a stronger version that holds with the following condition.
\begin{definition} \todo{Maybe this happens for all locally-observable games?}
  In a locally observable game, we say that actions $i$ and $i'$ are confoundable if $\ell_i - \ell_{i'}\notin \im S_{i,i'}^\top$ but there exists $p\in C_i$ and $p' \in C_{i'}$ such that $p - p' \in \ker S_{i,i'}$.
\end{definition}
This condition obviously holds if the game is not locally observable, but holds for many locally observable games as well.
\begin{theorem}
  Suppose a locally observable game $(L,H)$ has two confoundable actions. Then there exist a problem dependent constant $c$, a policy class $\Pi$, and a distribution over contexts such that, for any algorithm $\mathcal A$ and $T$, the expected regret has the lower bound
  \begin{equation*}
\ex[ \Regret_T(\mathcal A)] \geq c T^{2/3}.
  \end{equation*}
\end{theorem}
\begin{proof}[Proof outline]
  Without loss of generality, assume that action 1 and 2 are confoundable and let $S = S_{1,2}$ be their signaling matrix. It will suffice to use a stochastic counterexample. For some $\delta$, let the context $x_t$ be an i.i.d. sample from the distribution $P(X = 1) = 1-\delta$ and $P(X=2) = \delta$ and consider two policies: $\pi_1(x) = x$ and $\pi_2(x) = 1$.

  We design the adversary strategy as follows.  Since $1,2$ are confoundable, there exists a vector $v$ such that $(\ell_1 - \ell_2)^\top v \neq 0$ and distributions $p_A\in C_1$ and $p_B \in C_2$ with $p_A - p_B = v$. Define the adversary strategies $p_1:= p_A\mathds{1}\{X=1\} + p_B\mathds{1}\{X=2\}$ and $p_2 = p_B$.  Policy $\pi_1$ is optimal under $p_1$, since $p_1 \in C_{\pi_1(x)}$ holds for $X=1$ and $X=2$. Similarly, for $p_2$, the optimal action is always action $1$, and so on the event $X=2$, $\pi_1$ incurs extra loss while $\pi_2$ does not. Note that any other action incurs more loss as the distributions played by the environment always fall into $C_1$ or $C_2$.

The lower bound then argues that under $p_i$, the regret is lower bounded by the number of times an action not corresponding to $\pi_i$'s action is played. Then, the confoundable actions condition implies that the distribution of the feedback for policies playing in $C_1$ and $C_2$ is very similar, only differing when the algorithm chooses actions $i>2$, and which is further bounded by the closeness of $p_1$ and $p_2$. The full proof is in Section~\ref{appendix:lower.bound.proof}.
\end{proof}

\subsection{Changing Losses}
It is a straightforward extension to changing losses. Suppose the following game protocol: for each round $t=1,\ldots,T$,
\begin{itemize}
  \item The algorithm receives $x_t$, $L_t$, and $H_t$ 
  \item The algorithm calculates $q_t$, plays $I_t \sim q_t$, receives loss $e_{I_t}^\top L_t e_{j_t}$ and feedback $Y_t = e_{I_t}^\top H_t e_{j_t}$.
  \item Policy weights are updated with $\hat\Delta_t$
\end{itemize}
We assume that, for each $L_t$ and $H_t$, there are matrices $V_t(1),\ldots, V_t(N)$ such that $\tilde L_t = \sum_{i=1}^N V_t(i)^\top S_{t,i}$.

\begin{lemma}
  Suppose that the learner plays the changing loss game with $\hat\Delta_t = Z_t \sgn(V_t(I_t)^\top Y_t$ for a random variable $Z_t$ with $Z_t^\top e_i = V_\infty \gamma^{-1}$ with probability $\frac{\gamma}{V_\infty} \frac{|e_i^\top V(I_t)^\top Y_t|}{q_t(I_t)}$ and $0$ otherwise.
  Then both Algorithms~\ref{alg:exp4.pm} and \ref{alg:bistro.pm}, under their respective settings, control the regret
\begin{equation}
  \regret_T = \sum_{t=1}^T e_{I_t}^\top L_{t}e_{j_t} - \min_{\pi\in\Pi}\sum_{t=1}^T \pi(x_t)^\top L_{t}e_{j_t}
\end{equation}
to the same order.  
\end{lemma}

\section{Online Collaborative Filtering}
Recall our modeling assumptions: given a set of item feature vectors $w_1,\ldots,w_N$ and a threshold $\gamma$, we define the loss matrix $L_{i,j} = w_i^\top w_j$ and feedback matrix $H_{i,j} = \mathds{1}\{L_{i,j} \leq \gamma\}$. Hence, since there are only two possible feedback symbols, we wish to find $v$ coefficients such that $\ell_i - \ell_b = \sum_k S_k^\top v_{i,b,k}$, where the signaling matrices have the simple form
$
  S_i =
  \begin{bmatrix}
    e_i^\top H\\
    \ones- e_i^\top H
  \end{bmatrix}
  $.
  We wish to find matrices $V(i)$ with $\tilde L V(i)^\top S_k$; if we define the total signaling matrix $S$ as the stacked $S_k$ matrices and $V^\top = (V(1)^\top,\ldots,V(N)^\top)$, then we need only find a matrix such that $\tilde L = V^\top S$. We now observe that appylng the same permutation to the columns in $V$ and $S$ does not change their product, as so we may take $S=
  \begin{bmatrix}
    H\\
    \ones \ones^\top - H
  \end{bmatrix}$
  and solve for $V$ such that $\tilde L = V S$. In this case, $\hat\Delta_t$ is the random variable with expectation $\frac{1}{q(I_t)}V \begin{bmatrix}
 e_{I_t}\\e_{I_t} 
\end{bmatrix}
Y_t$.

We remark that, in this setting, local observability does hold: we would need $(r_i - r_{i'})^\top \in \im H^\top\left(e_i, e_{i'}\right) \oplus \im \ones$, which in general does not hold since $e_i^\top H$ is binary but $r_i$ is not. Without local observability, the relaxation algorithm gets the optimal $O(T^{2/3})$ rate.

\subsection{Policy Class}
The contextual partial monitoring algorithms above can be applied to arbitrary policy classes. Here, we conduct a detailed analysis of the policy class $\Pi(r,R)$.

The policy class we propose must map to $\triangle_N$, and hence cannot simply be linear, even if we assume that $x_t$ is normalized. Instead, we will consider $\Pi = \{ x\mapsto \sigma(M x) : \Vert m_k \Vert_2 \leq R\}$ where $\sigma$ maps positive vectors to the probability simplex (e.g. it is an exponential weights function) and $m_k$ is the $k$th row. Importantly, we want something that is fast to compute.

Here, we exploit the fact that we need only evaluate the oracle for data $(x_s, e_i (e_i^\top \tilde\Delta_t))$, e.g. that is sparse on the same cooordinate. Here are two proposals for $\Pi$. First, we let $\sigma$ be a simple normalization, i.e. $\sigma(x)_i = x_i / \Vert x \Vert_1$, and we constrain $M$ to be non-negative. In this case, the optimum $\pi$ only has support on the row vector corresponding to $D_i$. The size constraint $B$ turns out not to matter as $W_i$ will be constrained to have unit 1 norm anyway. The optimum $\pi$ is the solution of
\begin{align*}
    \max_{m_i\geq 0:\Vert m_i\Vert_1 = 1} \sum_s (\tilde\Delta_s^\top e_i) m_i^\top x_s
  &=
    \max_{m_i\geq 0:\Vert m_i\Vert_1 = 1} m_i^\top \sum_s (\tilde\Delta_s^\top e_i)x_s,
\end{align*}
which we can evaluate by inpection to be $\frac{[\sum_s (\tilde\Delta_s^\top e_i)x_s]_+}{\Vert [\sum_s (\tilde\Delta_s^\top e_i)x_s]_+\Vert_1}$.

Second, we let $\sigma(x)_i = \frac{\exp(-\eta x_i)}{ \sum_i \exp(-\eta x_t)}$, which does not require non-negativity of $x$. Unfortunately, this is harder to optimize, but there are connectios with multi-class logistic regression.

We describe how to evaluate the oracle. A straightfroward implementation would require calculating, for arbitrary contexts $x_1,\ldots,x_t$ and $\ell_1,\ldots,\ell_t$,
\[
  \min_{M: \Vert M \Vert \leq R, \rank(M) \leq r} \sum_{s=1}^t \ell_t (\argmin_j e_j^\top M x_t),
\]
which is intractable. Instead, we will allow the policy class to predict a distribution over items and minimize
\[
  \min_{M: \Vert M \Vert \leq R, \rank(M) \leq r} \sum_{s=1}^t \ell_t^\top M x_t
\]
which is equivalent to learning the PCA of $M$ in the following way.

The singular decomposition of $M$ is $M = \sum_{k=1}^r \sigma_k u_k v_k^\top$, for scalars $\sigma_k$ and unit vectors $u_k$ and $v_k$. Hence, the optimization becomes
\begin{align*}
  \sum_{k=1}^r\sum_{s=1}^t \sigma_k \ell_t^\top u_k v_k^\top x_t
  =
  \sum_{k=1}^r \sigma_k u^\top_k \left( \sum_{s=1}^t \ell_t x_t^\top\right) v_k
\end{align*}
subject to $u_k$, $v_k$ unit vectors and some bound on $\sum_k \sigma_k$.

The singular vectors of a matrix $M$ can be derived from the varitional formula
\[
  u_k, v_k = \argmin_{u, v: ||u|| = ||v|| = 1, u\perp u_1,\ldots,u_{k-1}, v\perp v_1,\ldots,v_{k-1}} v^\top M v
  \]
  and $\sigma_k = v_k^\top M u_k$. Hence, it seems that we should choose $u_k$ and $v_k$ equal to the singular vectors of $\sum_t \ell_t x_t^\top$ and $\sigma_k$ equal to the singular values. However, if we are just maximizing the above quantity for $u_k,v_k$, and $\sigma_k$, then it seems that we should allocate all the $R$ budget to $u_1$ and $v_1$.

  How should we threshould $\sigma$? If we are looking to maximize $M$ subject to $||M||_2\leq R$, then thin equivalent to all $\sigma_k \leq R$ and we should hard threshold everything. If we want a total budget on all the $\sigma$, then it will probably make the most sense to put the whole budget on the first singular value. In fact, we can evaluate the best $M$ explicitly; using $\Delta_i$ to represent the $i$th row of the matrix with columns $\hat\Delta_t$ and $X^t$ to represent the matrix with columns $x_1,\ldots, x_T$, we can show $M = e_i X^t\Delta_i$.

% \subsection{The policy class}
% A policy $\pi$ maps from the context space $\mathcal X$ into a recommended action $\pi(x_t)$. Our framework applies to any policy class that is either finite or has an ERM oracle. For concreteness, we will propose the following policy class.

% Each policy $\pi$ will be parameterized by a type matrix $T_\pi\in \Reals^{d\times M}$ and will recommend action $\argmax_i r_i^\top T_\pi x_t$. We can interpret $T_\pi x_t$ as a vector of weights specifying how likely user $t$ belongs to each type according to policy $\pi$ and $r_i^\top T_\pi x_t$ as the weighted reward of showing action $i$; it is then natural for the policy to recommend the best weighted action. In the case when $T_\pi x_t$ exactly captures the relationship, policy $\pi$ will recommend the best fitting item for each user.

% \subsection{User item matrix}
% A central object in the study of many collaborative filtering algorithms is the user-item matrix. The latent user-type model is expressible in terms of the SVD of this matrix, $U\Sigma V^\top$, where $U$ is the item-type matrix and $V$ is the type-user matrix. We can define the policy class as the set of all type-user matrices, and hence our regret bound will imply that we predict almost as well as if we knew the user's type.

% Hence, we can view our formalism as precomputing the item-type matrix and learning the type-user matrix online. The question is: can we overparameterize the item-type (by increasing the number of types) to allow greater flexibility in the model?

\subsection{Regret Bound}
We can apply Corollary~\label{cor:oracle.regret.bound} to obtain a regret bound for our algorithm once we have a bound on the complexity term:
\begin{lemma} Let $\eepsilon_t$ be a diagonal matrix of $N$ Rademacher random variables and $\Z_t$ the vector of Bernoulli $\gamma$ random variables. Then the complexity term for $\Pi(r,R)$ has the following bound:
  \begin{equation}
    \exop_{\Z_{1:T},\eepsilon_{1:T}}\left[\sup_{\pi\in\Pi(r,R)} \sum_{t=1}^T \pi(x_t)^\top\eepsilon_t Z_t \right]
    \leq
    C(r,R).
  \end{equation}
\end{lemma}
\begin{proof}
  Some math...
\end{proof}
\begin{corollary}
  The collaborative filtering algorithm has an expected regret bound of 
\begin{equation*}
    \ex[\regret_t]
    \leq
\exop_{\Z_{1:T},\eepsilon_{1:T}}\left[\sup_{\pi\in\Pi} \sum_{t=1}^T \pi(x_t)^\top\eepsilon_t Z_t \right]    
    +
    T N \gamma.
  \end{equation*}
\end{corollary}

\section{Experiments}
We ran the algorithm and this is what we found.
\section{Conclusion}

\bibliography{../big_bib}

\appendix
\onecolumn
\section{Missing Proofs}\label{sec:appendix.proofs}

\subsection{Missing proofs for relaxations}
\begin{proof}[Proof of Lemma~\ref{lem:complexity.bound}]
  This reasoning is a small refinement of the proof of Lemma 2 in \cite{syrgkanis2016improved}. We evaluate
  \begin{align*}
        \ex_{\Z_{1:T},\eepsilon_{1:T}}\left[\sup_{\pi\in\Pi} \sum_{t=1}^T \pi(x_t)^\top \diag( \eepsilon_t) \Z_t \right]
    &=
      \ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
      \ex_{\eepsilon_{1:T}}\left[
      \log\left(
      \sup_{\pi\in\Pi}e^{\lambda \sum_{t=1}^T \pi(x_t)^\top \diag(\eepsilon_t) \Z_t}\right)\right]
      \right]\\
    &\leq
      \ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
      \log\left(
      \ex_{\eepsilon_{1:T}}\left[
      \sum_{\pi\in\Pi}e^{\lambda \sum_{t=1}^T \pi(x_t)^\top \diag(\eepsilon_t) \Z_t}\right]\right)
      \right]\\
    &\leq
      \ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
      \log\left(\sum_{\pi\in\Pi}\prod_{t=1}^T
      \ex_{\eepsilon_{1:T}}\left[
      e^{\lambda \pi(x_t)^\top \diag(\eepsilon_t) \Z_t}\right]\right)
      \right].
  \end{align*}
Using the inequality
  \[
    \ex_{\eepsilon_{1:T}}\left[ e^{\lambda \pi(x_t)^\top \diag(\eepsilon_t) \Z_t}\right]
    =
    \prod_{i=1}^N
    \ex_{\epsilon_i}\left[
      e^{\lambda \epsilon_{t,t} Z_{t,i}}
    \right]
    =
    \prod_{i=1}^N
    \frac{
      e^{\lambda\epsilon_{t,t} Z_{t,i}}
      +
      e^{-\lambda \epsilon_{t,t} Z_{t,i}}
    }{2}
    \leq
    e^{\frac{1}{2}\lambda^2 \sum_{i=1}^N Z_{t,i}^2}
  \]
with the above expression produces
  \begin{align*}
    \ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
    \log\left(\sum_{\pi\in\Pi}\prod_{t=1}^T
    \ex_{\eepsilon_{1:T}}\left[
    e^{\lambda \pi(x_t)^\top \diag(\eepsilon_t) \Z_t}\right]\right)
    \right]
    &\leq
      \ex_{\Z_{1:T}}\left[\frac{1}{\lambda}
      \log\left(\sum_{\pi\in\Pi}\prod_{t=1}^T
      e^{\frac{\lambda^2 \sum_{i=1}^N Z_{t,i}^2 }{2}}
      \right)
      \right]\\
    &\leq
      \frac{1}{\lambda}\log(\vert \Pi\vert)
      + 
      \sum_{t=1}^T \sum_{i=1}^N \ex_{\Z}\left[
      \frac{\lambda Z_{t_i}^2 }{2}
      \right]\\    
    &\leq
      \frac{1}{\lambda}\log(\vert \Pi\vert)
      +
      \frac{1}{2}W T\lambda N
  \end{align*}
  Setting $\lambda = \sqrt{2\log(\vert \Pi\vert)/W N T}$ finishes the proof.
\end{proof}
\begin{proof}[Proof of Theorem~\ref{thm:admissibility}]
  The base case is easy. Using convexity of $\sup$ and the unbiasedness of $\hat\Delta_t$,
  \todo{write out in more explicit detail}
  \begin{align*}
  \ex_{I_{1:T},\hat Z_{1:T}}\left[
  \rel(\hist_{1:T})
  \right]
  &=
    \ex_{I_{1:T},\hat Z_{1:T}}\left[
    \sum_{i=1}^N\sup_{\pi\in\Pi}
    -\sum_{s=1}^T \pi(x_t)^\top D_i \hat \Delta_t
    \right]\\
  &\geq
    \ex_{I_{1:T},\hat Z_{1:T}}\left[    
    \sup_{\pi\in\Pi}
    -\sum_{s=1}^T \pi(x_t)^\top \hat \Delta_t
    \right]\\
  &\geq
    \sup_{\pi\in\Pi}
    \ex_{I_{1:T},\hat Z_{1:T}}\left[
    -\sum_{s=1}^T \pi(x_t)^\top \hat \Delta_t
    \right]\\
      &\geq
    \sup_{\pi\in\Pi}
    -\sum_{s=1}^T \pi(x_t)^\top \tilde L e_{j_t}.
  \end{align*}

  We now check the inductive step. We define $\rho = (x,\eepsilon,Z)_{t+1:T}$ to be a sample of the random variables in the relaxation. Recall that our aim is to prove admissibility of the strategy $q_t(\rho) = (1-N\gamma)q_t^*(\rho)- \gamma \ones$ where $q_t^*$ was defined in \eqref{eqn:strategy.def}.  By construction, no entry of $q_t(\rho)$ is below $\gamma$. It is then easy to check that, for $q_t^* = \ex_{\rho}[q_t^*(\rho)]$ and $q_t = \ex_{\rho}[q_t(\rho)]$,
\[
  \ex_{I_t \sim q_t}\left[ e_{I_t}^\top \tilde L e_{j_t} \right] = q_t^\top \tilde L e_{j_t} \leq (q_t^*)^\top\tilde L e_{j_t} + N\gamma
  =
  \ex_{I_t\sim q_t} \left[ (q_t^*)^\top\hat \Delta_t(I_t,j_t,q_t) \right] + N\gamma.
\]
% We will use the shorthand $\hat L_t(I_t,q_t) = \frac{M(I_t)S_{I_t}}{q_t(I_t)}$, noting that $\hat\Delta_t(i,j) = L_t(i)e_j$. Since we always have an $S_{I_t}$ left multiplying any $e_{j_t}$, we can calculate $\hat L_t(I_t, q_t) e_{j_t}$ from $Y_t$. Hence, we can replace $\hat\Delta_t$ by $\hat L(I_t, q_t)e_{j_t}$.
To ease notation, we write $\hat\Delta_t$ instead of $\hat \Delta_t(I_t,j_t,q_t)$. For every fixed $x_t$,
\begin{align*}
  \sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[ e_{I_t}^\top \hat \Delta_t
  + \mathop{\ex}_{\rho}[\rel(1:t)]\right]
  &\leq
  % \inf_{q_t\in\triangle_N'}
    \sup_{j_t } \mathop{\ex}_{I_t\sim q_t}\left[ (q_t^*)^\top \hat \Delta_t
    + \mathop{\ex}_{\rho}[\rel(1:t)]\right] + N\gamma\\
  &\leq
    %\inf_{q_t\in\triangle_N'}
    \sup_{j_t } \mathop{\ex}_{I_t\sim q_t}\left[ (q_t^*)^\top \hat \Delta_t
    +
    \mathop{\ex}_{\rho}\left[\sum_{i=1}^N\sup_{\pi\in\Pi} - \pi(x_t)^\top D_i\Delta_t
    + A_i(\pi)\right]\right]\\
  &\quad+(N+1)\gamma\\
  &\leq
  % \inf_{q_t\in\triangle_N'}
    \sup_{j_t } \sum_{i=1}^N\mathop{\ex}_{I_t\sim q_t}\left[\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top \hat \Delta_t
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\Delta_t + A_i(\pi)\right]\right]\\
  &\quad+(N+1)\gamma.
\end{align*}
where $A_i(\pi) = -\sum_{s=1}^{t-1} \pi(x_s)^\top D_i \hat \Delta_s - \sum_{s=t+1}^T 2\pi(x_s) D_i \eepsilon_{s} Z_{s}$.

We will relax the problem by allowing the adversary to optimize over distributions of $\hat\Delta_t$ instead of a $\hat\Delta_t(I_t,j_t,q_t,\hat Z_t)$ that corresponds to a specific choice of $I_t,j_t$, and $q_t$. An easy calculation show that $\hat\Delta_t$ has large probability of being zero. Conditioning on $\rho$,
\begin{align*}
  P\left(\hat\Delta_t^\top e_i = 2\gamma^{-1}\right)
  =
  \ex_\rho\left[P\left(\hat\Delta_t^\top e_i = 2\gamma^{-1}\middle | \rho \right) \right],
\end{align*}
and so
\begin{align*}
  P\left(\hat\Delta_t = 0 \right)
  &=
  \ex_{\rho}\left[ \ex_{I_t \sim q_t(\rho)}\left[ P\left(\hat\Delta_t =0 \middle |\rho, I_t\right))\right]\right]\\
  &=
    \ex_{\rho}\left[ \ex_{I_t \sim q_t(\rho)}\left[
    1-\frac{\gamma}{q_t(\rho)(I_t)}
    \right]\right]\\
  &= \ex_{\rho}\left[
    \sum_{i=1}^N q_t(\rho)(i) \left(1- \frac{\gamma}{q_t(\rho)(i)}\right)\right]\\
  &= 1 - N \gamma.
\end{align*}

Hence, the distribution of $\hat\Delta_t \in \triangle_{\mathcal D}'$, and we can obtain an upper bound of the previous term (suppressing the $(N+1)\gamma$ for now) of \todo{Make the step where we replace $\hat\Delta(I_t,q_t,\hat Z_t,j_t)$ by a generic $\hat\Delta_t$ clear}
\begin{align*}
  &\lefteqn{
        \sup_{j_t } \sum_{i=1}^N\mathop{\ex}_{I_t\sim q_t}\left[\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top \hat \Delta_t
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\Delta_t + A(\pi)[j]\right]\right]
    }\\
  &\leq
    \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top \hat \Delta_t
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\Delta_t + A_i(\pi)\right]\right]\\
    &\leq
\mathop{\ex}_{\rho}\left[ \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^*(\rho)^\top \hat \Delta_t
    +
  \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]\right].
\end{align*}

For now, fix $\rho$. Our definition of $q_t^*$ was 
\[
  q_t^*(\rho)
  =
  \argmin_{q \in \triangle_{N}}
  \sup_{p_t \in \triangle_{\mathcal D}'}
  \ex_{\hat\Delta_t\sim p_t}\left[
  q^\top \hat \Delta_t
    +
    \sum_{i=1}^N \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right],
  \]
  and so
  \begin{align*}
    &\lefteqn{\sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^*(\rho)^\top \hat \Delta_t
    +
  \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]}\\
    &=
      \inf_{q_t}
      \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top \hat \Delta_t
    +
  \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]
\end{align*}
is an upper bound on $  \sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[ e_{I_t}^\top \hat \Delta_t
  + \mathop{\ex}_{\rho}[\rel(1:t)]\right] -(N+1)\gamma$.

We continue bounding this saddle point problem from above. Noting that the objective is linear in both $q_t$ and $p_t$, we may perform a min-max swap:
\begin{align*}
  &\lefteqn{
          \inf_{q_t}
      \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top \hat \Delta_t
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]
    }\\
  &\leq
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \inf_{q_t}    
    \sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top \hat \Delta_t
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]\\
  &\leq
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \inf_{j}  
    \sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j \hat \Delta_t
    +
    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]\\
  &=
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N
    \inf_{j}  \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j \hat \Delta_t\right]
    +
\mathop{\ex}_{\hat\Delta_t\sim p_t}\left[  \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]. 
\end{align*}
Because $\pi$ is deterministic, we must have $\min_i \ex_{\hat\Delta_t \sim p_t}
[e_i^\top\hat\Delta_t] \leq \ex_{\hat\Delta_t \sim p_t }[\pi(x_t)^\top\hat\Delta_t]$, which allows us to upper bound the previous expression and perform the usual symmeterization (using $\epsilon$ as a single Rademacher random variable)
\begin{align*}
  &\lefteqn{
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t}\left[
     \sup_{\pi\in\Pi} A_i(\pi)
    +\min_j \ex_{\hat\Delta_t'\sim p_t}\left[e_j^\top \hat\Delta_t'\right]
    -\pi(x_t)^\top D_i \hat\Delta_t
  \right]
  }\\
  &\leq
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t}\left[
     \sup_{\pi\in\Pi} A_i(\pi)
    +\ex_{\hat\Delta_t'\sim p_t}\left[\pi(x_t)^\top \hat\Delta_t'\right]
    -\pi(x_t)^\top D_i \hat\Delta_t
  \right]\\
  &\leq
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t, \hat\Delta_t'\sim p_t}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        \pi(x_t)^\top D_i\left(\hat\Delta_t'-\hat\Delta_t\right)
    \right]\\
  &=
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t, \hat\Delta_t'\sim p_t,\epsilon_i}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        \epsilon_i\pi(x_t)^\top D_i\left(\hat\Delta_t'-\hat\Delta_t\right)
    \right]\\
    &\leq
    \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t,\epsilon}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\epsilon\pi(x_t)^\top D_i\hat\Delta_t
        \right].
\end{align*}

Define $\triangle_1'$ to be the one-dimensional restriction of $\triangle_{\mathcal D}'$, that is
\[
  \triangle_1' = \{ p \in \triangle_{\{V_\infty \gamma^{-1} e_i,0,-V_\infty \gamma^{-1} e_i\}}: p(0)\geq 1-N\gamma\}.
\]
 We can think of $\triangle_{\mathcal D}'$ as distributions such that, if $X\sim p\in\triangle_{\mathcal D'}$, then $X^\top e_i$ has a distribution in $\triangle_1'$. The supremum can then be simplified as follows:
\begin{align*}
  \sup_{p_t\in\triangle_{\mathcal D}'}
    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t,\epsilon}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\epsilon\pi(x_t)^\top D_i \hat\Delta_t
  \right]
  &\leq
      \sup_{p_1,\ldots,p_N \in\triangle_1'}
    \sum_{i=1}^N \ex_{X_i\sim p_i,\epsilon}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\epsilon\pi(x_t)^\top e_i X_i
    \right].
\end{align*}

We will now argue that the witness to the supremum of 
\begin{align*}
    \sup_{p\in\triangle_i'}
  \ex_{\delta\sim p,\epsilon}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\epsilon\pi(x_t)^\top e_i X_i
        \right]
\end{align*}
is the distribution that puts mass $N\gamma$ on $V_\infty\gamma^{-1}$ and the rest on $0$.
Define the convex function 
$g(x) :=  \sup_{\pi\in\Pi} A_i(\pi)+ 2\pi(x_t)^\top e_i x$. The expectation $\ex_{\epsilon}\left[ g(\epsilon x)\right]$ is increasing on $x\geq 0$. For an easy proof, let $0\leq a<b$, which allows us to write $a = \theta b + (1-\theta)(-b)$ for some $\theta$. Then, by the definition of convexity,
 \begin{align*}
   \ex_{\epsilon}\left[ g(\epsilon a) \right]
   = 
   \frac{g(a)+g(-a)}{2}
   \leq
   \frac{\theta g(b)+(1-\theta)g(-b)
   +(1-\theta) g(b)+\theta g(-b)
   }{2}
   =
   \ex_{\epsilon}\left[ g(\epsilon b) \right].
 \end{align*}
 Since $\ex_{\epsilon}\left[ g(\epsilon x)\right]$ in increasing, the supremum of $p_i$ puts as much mass as possible on the largest element, $V_\infty\gamma^{-1}$. Hence, defining the random variable $Z_{t,1} \in \{0,V_\infty\gamma^{-1}\}$ with $P(Z = 0) = 1-N\gamma$, we have
 \begin{align*}
    \sup_{X_i\in\triangle_i'}
  \ex_{X_i\sim p,\epsilon}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\epsilon\pi(x_t)^\top X_i
   \right]
   =
  \ex_{\epsilon}\left[
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\epsilon\pi(x_t)^\top D_i Z_t
        \right].
\end{align*}

In total, we have shown that, for every fixed $x_t$,
\begin{align*}
    \sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[ e_{I_t}^\top \Delta_t
  + \mathop{\ex}_{\rho}[\rel(1:t)]\right]
  &\leq
    \ex_\rho \left[
  \ex_{\epsilon, Z}\left[
  \sum_{i=1}^N
        \sup_{\pi\in\Pi} A_i(\pi)+
        2\pi(x_t)^\top D_i \eepsilon_t Z_t
  \right]
  \right]
  + (t-1)N\gamma\\
  &=
  \ex_\rho[\rel(1:t-1)],
\end{align*}
as required.
% \paragraph{Possible improvement:} define $\hat\Delta_t$ to set each coordinate to 0 separately may yield $P(\hat\Delta_t[i] = 0) = 1-2\gamma$, shaving of a factor of $N$. We will pick up extra $N$ in the finite-class complexity bound lemma, though. 
\end{proof}

\section{Proof of the lower bound}\label{appendix:lower.bound.proof}
This section proved the $T^{2/3}$ lower bound for the contextual partial monitoring.
\begin{proof}[Proof of Theorem~\ref{thm:lower.bound}]
  The proof reuses many arguments from \cite{bartok2011minimax}.

  Without loss of generality, assume that action 1 and 2 are confoundable and let $S = S_{1,2}$ be their signaling matrix. It will suffice to use a stochastic counterexample. For some $\delta$, let the context $x_t$ be an i.i.d. sample from the distribution $P(X = 1) = 1-\delta$ and $P(X=2) = \delta$ and consider three policies: $\pi_1(x) = x$, $\pi_2(x) = 1$, and any $\pi_3$ that can arbitrarily choose from actions $i > 2$.

Since $1,2$ are confoundable, there exists a vector $v$ such that $(\ell_1 - \ell_2)^\top v \neq 0$ and distributions $p_A\in C_1$ and $p_B \in C_2$ with $p_A - p_B = v$. Define the adversary strategies $p_1:= p_A\mathds{1}\{X=1\} + p_B\mathds{1}\{X=2\}$ and $p_2 = p_B$.  Policy $\pi_1$ is optimal under $p_1$, since $p_1 \in C_{\pi_1(x)}$ holds for $X=1$ and $X=2$. Similarly, for $p_2$, the optimal action is always action $1$, and so on the event $X=2$, $\pi_1$ incurs extra loss while $\pi_2$ does not. Note that any other action incurs more loss as the distributions played by the environment always fall into $C_1$ or $C_2$.

We now define a few variables. Let $N_k^{p_i}$, for $i = {1,2}$ and $k = {1,2}$, be the expected number of times algorithm $\mathcal A$ follows policy $\pi_k$ when the environment plays distribution $p_i$. Also define $N_3^{p_i}$ for $i=1,2$ to be the expected number of times the algorithm plays any other action that does not agree with $\pi_1$ or $\pi_2$. These expectations are over the context and the randomness of $J_t$.

Define $\beta = \min_{i>2} \min\left\{\ell_i^\top p_A - \ell_1^\top p_A, \ell_i^\top p_B - \ell_2^\top p_B\right\}$, a lower bound on the excess loss on choosing any action $i>2$. Also let $\epsilon = \left |\ell_1^\top p_B - \ell_2^\top p_A\right|$ be the extra loss of choosing action $3-i$ when action $i$ would be optimal.

As shown earlier, under $p_i$, policy $\pi_i$ will be the minimizer in the comparator term in the regret. Hence, using $\Regret^{p_i}_T$ for the expected regret under $p_i$, we can bound, for $i={1,2}$,
\[
  \Regret^{p_i}_T \geq N_{3-i}^{p_i}\epsilon\delta + N_3^{p_i}\beta.
\]

We now need to show that $N^{p_i}_{3-i}$ and $N^{p_i}_3$ occur often enough since $p_1$ and $p_2$ are close and the feedback provided from picking action 1 and 2 is the same.

\begin{lemma}\label{lem:N.lower.bound}
  There exists a problem dependent constant $c$ such that, for $i=1,2$, we have
  \begin{equation*}
  N^{p_{3-i}}_{i} \geq N^{p_{i}}_{i}  - T \delta^2 c_1 \sqrt{N^{p_{i}}_{3}}.
  \end{equation*}  
\end{lemma}
\begin{proof}
  Any algorithm receives two forms of feedback: the contexts $x_t$ and the feedback symbols, and we denote the symbol received at round $t$ by $f_t\in\Sigma$. Let $p_i^*(f_t| f_{1:t-1},x_{1:t})$ be the mass function over $\Sigma$ at round $t$ generated by the algorithm's choices if the adversary uses strategy $p_i$. For $i\in {1,2}$ and $k\in\{1,2,3\}$, we can bound
  \begin{align*}
    N^{p_i}_k-N^{p_{3-i}}_k
    &=
      \sum_{f^{T-1}}( p_i^* (f_{1:T-1}|x_{1:T}) - p_i^* (f_{1:T-1}|x_{1:T}))
      \sum_{t=0}^{T-1} \mathds{1}\left\{ \mathcal A(f_{1:t-1},x_{1:t}) = k
      \right\}\\
    &\leq
      \sum_{f^{T-1}:p_i^* (f_{1:T-1}|x_{1:T}) \geq p_i^* (f_{1:T-1}|x_{1:T})}
      ( p_i^* (f_{1:T-1}|x_{1:T}) - p_i^* (f_{1:T-1}|x_{1:T}))
      \sum_{t=0}^{T-1} \mathds{1}\left\{ \mathcal A(f_{1:t-1},x_{1:t}) = k
      \right\}\\
    &\leq
      T \sum_{f^{T-1}:p_i^* (f_{1:T-1}|x_{1:T}) \geq p_i^* (f_{1:T-1}|x_{1:T})}
      p_i^* (f_{1:T-1}|x_{1:T}) - p_i^* (f_{1:T-1}|x_{1:T})\\
    &=
      \frac{T}{2}\left\Vert p_i^* (f_{1:T-1}|x_{1:T}) - p_i^* (f_{1:T-1}|x_{1:T})\right\Vert_1 \\
    &\leq
      T \sqrt{ \frac{D_{KL}\left( p_2^* || p_1^*\right)}{2}},
  \end{align*}
where the last line followed by Pinsker's inequality.

  We define $p_1^*(f_t|f_{1:t-1},x_{1:t})$ to be the conditional probability of feedback symbol $f_t$ given all the knowledge accrued by the algorithm until that point, which allows us to apply KL-divergences chain rule $T$ times:
  \begin{align*}
    D_{KL}(p_2^*||p_1^*)
    &=
      \sum_{t=1}^{T-1}\sum_{f_{1:t-1}}p_2^*(f_{1:t-1}) \sum_{f_t} p_2^*(f_t|f_{1:t-1},x_{1:t})
      \log \frac{ p_2^*(f_t|f_{1:t-1},x_{1:t})}{ p_1^*(f_t|f_{1:t-1},x_{1:t})}\\
    &=
      \sum_{t=1}^{T-1}\sum_{f_{1:t-1}}p_2^*(f_{1:t-1})
      \sum_{k=1}^3\mathds{1}\left\{ \mathcal A(f_{1:t-1},x_{1:t}) = k \right\}
      \sum_{f_t} p_2^*(f_t|f_{1:t-1},x_{1:t})
      \log \frac{ p_2^*(f_t|f_{1:t-1},x_{1:t})}{ p_1^*(f_t|f_{1:t-1},x_{1:t})}.
  \end{align*}
  Let us examine the event where $A(f_{1:t-1},x_{1:t})\in\{ 1,2\}$ and feedback $f_t$ is received, where $f_t$ must be in the feedback set of action 1 or 2 by definition of our policy class. If the feedback set of $\pi_1(x_t)$ does not cover $f_t$, then $p_i^*(f_t| f_{1:t-1},x_{1:t}) = 0$ for $i=1,2$ and we the contribution to the $\sum_{f_t}$ is zero. Otherwise, let $e_{f_t}$ be the basis vector identifying the row of feedback symbol $f_t$, i.e. $e_{f_t}^\top S e_{j_t} = Y_t$. Then $p_i^*(f_t| f_{1:t-1},x_{1:t}) = e_{f_t}^\top p_i(x_t)$, where $p_i(x_t) = p_a$ or $p_b$ depending on $x_t$ and $i$. For every case, $e_{f_t}^\top p_a = e_{f_t}^\top p_b$ by construction of $p_a$, $p_b$, and so $p_i^*(f_t| f_{1:t-1},x_{1:t}) = p_{3-i}^*(f_t| f_{1:t-1},x_{1:t})$. Thus, only the $\{A(f_{1:t-1},x_{1:t})=3\}$ terms contribute to the $\sum_{f_t}$.

  It still remains to bound the $D_{KL}( p_2^*(f_t|f_{1:t-1},x_{1:t}) || p_1^*(f_t|f_{1:t-1},x_{1:t}))$. Let us define $p = (1-\frac{\delta}{2})p_a + \frac{\delta}{2} p_b$ and $p_\delta = \frac{\delta}{2} (p_b - p_a)$. Thus, $p_1 = p + p_\delta$ and $p_2 = p - p_\delta$. Hence, we can apply Lemma~12 from \cite{bartok2011minimax}, which implies that
\[
  D_{KL}(p_2 || p_1) = O(\Vert p_\delta\Vert_2^2)
\]
for $\delta$ small enough. Hence, $D_{KL}(p_2 || p_1) \leq c_1 \delta^2$ for some constants $c_1$ and $\delta_{\max}$ and any $\delta \leq \delta_{\max}$. The constants depend on the problem but not on $\delta$.

Finally, we can apply this bound and calculate
\begin{align*}
    D_{KL}(p_2^*||p_1^*)
    &\leq
      \sum_{t=1}^{T-1}\sum_{f_{1:t-1}}p_2^*(f_{1:t-1})
      \mathds{1}\left\{ \mathcal A(f_{1:t-1},x_{1:t}) = 3 \right\}
      \sum_{f_t} p_2^*(f_t|f_{1:t-1},x_{1:t})
      \log \frac{ p_2^*(f_t|f_{1:t-1},x_{1:t})}{ p_1^*(f_t|f_{1:t-1},x_{1:t})}\\
    &\leq
      \sum_{t=1}^{T-1}\sum_{f_{1:t-1}}p_2^*(f_{1:t-1})
      \mathds{1}\left\{ \mathcal A(f_{1:t-1},x_{1:t}) = 3 \right\}
      2c_1 \delta^2\\
    &\leq 2c_1\delta^2 N^{p_2}_3,
  \end{align*}
Plugging this expression into the first inequality yields the claim.  
\end{proof}

Now we can assemble the general lower bound. Define $r = \argmin_{i\in\{1,2\}} N^{p_i}_3$.  Lemma~\ref{lem:N.lower.bound} implies that, for $i=1,2$,
$N^{p_{3-r}}_{r} \geq N^{p_{r}}_{r}  - Tc_1\delta\sqrt{ N^{p_{r}}_{3}}$, and we trivially have $N^{p_{r}}_{r} \geq N^{p_{r}}_{r}  - Tc_1\delta\sqrt{ N^{p_{r}}_{3}}$. Together, this implies that
\[
  N^{p_{i}}_{r} \geq N^{p_{r}}_{r}  - Tc_1\delta \sqrt{ N^{p_{r}}_{3}}.
\]
Now consider an adversary strategy that plays $p_1$ and $p_2$ with equal probability. The expected regret has the lower bound
\begin{align*}
    \Regret_T
  &=
    \frac{1}{2}\left(\Regret_T^{p_1} + \Regret_T^{p_2}\right)\\
  &\geq
    \frac{1}{2}\left(
    \epsilon\delta (N^{p_1}_2 + N^{p_2}_1) + \beta (N_3^{p_1} + N_3^{p_2})
    \right)\\
  &\geq
    \frac{1}{2}\epsilon\delta\left(    
    N^{p_r}_2 + N^{p_r}_1 - 2Tc_1\delta\sqrt{N^{p_r}_3}
    \right)
    +
    \frac{\beta}{2}(N_3^{p_r} + N_3^{p_r})\\
  &=
    \frac{1}{2}\epsilon\delta\left(    
    T-N^{p_r}_3 - 2Tc_1\delta\sqrt{N^{p_r}_3}
    \right)
    +
    \beta N_3^{p_r}.
\end{align*}
Now, we can choose $\delta = c_2 T^{-1/3}$, for some to be determined constant $c$, \todo{what about contexts or adversary strategies that don't need to know the time horizon?} to obtain
\begin{align*}
  \Regret_T
  &\geq
    \frac{1}{2}\epsilon\delta\left(    
    T-N^{p_r}_3 - 2Tc_1\delta\sqrt{N^{p_r}_3}
    \right)
    +
    \beta N_3^{p_r}\\
  &\geq
    T^{2/3}\left(    
    \frac{\epsilon c_2}{2}
    +
    \left(\beta -  \frac{\epsilon c_2}{2}\right)N^{p_r}_3T^{-2/3}
    - \epsilon c_1c_2^2\sqrt{N^{p_r}_3T^{-2/3}}
    \right).
\end{align*}
The proof will be complete once we can prove existence of a $c_2$, independent of $T$, such that the term on the right is positive for every $N^{p_r}_3$. First, we recognize that the coefficient is a quadratic in $x = \sqrt{N^{p_r}_3T^{-2/3}}$ and we complete the square:
\begin{align*}
    \frac{\epsilon c_2}{2}
    +
    \left(\beta -  \frac{\epsilon c_2}{2}\right)x^2
  - \epsilon c_1c_2^2 x
  &=
    \left(\beta -  \frac{\epsilon c_2}{2}\right)
    \left(x - \frac{\epsilon c_1 c_2^2}{2\left(\beta -  \frac{\epsilon c_2}{2}\right)}\right)^2
    -
   \frac{(\epsilon c_1 c_2^2)^2}{4\left(\beta -  \frac{\epsilon c_2}{2}\right)}
    + \frac{\epsilon c_2}{2},
\end{align*}
so positivity is equivalent to
\[
  \frac{1}{\epsilon}
  \geq
  \frac{c_1 c_2^3}{\beta -  \frac{\epsilon c_2}{2}}.
\]
Since the right hand side is a decreasing function of $c_2$, we can choose $c_2$ small enough to satisfy this inequality and $c_2 T^{-1/3}\leq\delta_{\max}$ condition required by Lemma~12.

% \textbf{problem!} I don't think we can actually argue for the $T^{2/3}$ rate here. We need to group $N_3^{p_r}/T^{2/3}$ then complete the square to get a sufficient condition on positivity. However, since the lemma below has a $\epsilon$ instead of $\epsilon^2$ dependence, the $\delta \sqrt{N}$ becomes a $\sqrt{\delta N}$ which does not yield the proper scaling.

% Two options:
% \begin{enumerate}
% \item Change the counter example so that the KL distance between $p_1$ and $p_2$ has the $\epsilon^2$ scaling. Will have to do something besides a uniform mixture.
% \item See if we can avoid Pinsker's inequality. The tricky part will be replacing the KL chain rule with some analog using TV. 
% \end{enumerate}

\end{proof}

\section{Previous settings}
We briefly recount the settings of previous online collaborative filtering algorithms.
\begin{itemize}
\item \emph{A latent Source Model for Online Collaborative Filtering}
  \begin{itemize}
  \item Cosine-similarity
  \item Assume $k$ user types with identical preferences
  \item Each round, every user is recommended an item and then rates the item in $\{-1,1\}$
  \item Cannot recommend an item to a user twice (key distinction from bandit literature)
  \item Algorithms either explores greedily using cosine similarity or does one of two exploration steps: either a random unrated item is given to each user, or a random permutation is selected and each user get recommended the first unrated item on the list. Otherwise, unrated items maximizing a score are recommended for each user
  \item The score of an item for a user is an average rating of that item from users in the neighborhood (defined as having large enough cosine similarity)    
  \item They model the feedback as follows
    \begin{itemize}
    \item Each user $u$ has a length $m$ vector $p_u\in[0,1]^m$ with entries that indicates the probability user $u$ will like that item (if $p_{ui}\geq 1/2$).
    \item Every $p_u$ is one of $k$ types, each which happen with equal probability      
    \item No ambiguous items: $|p_{ui}-\frac12|\geq \Delta$ 
    \item $\gamma$ incoherence: users $u$ and $v$ have preference vectors $p_u$ and $p_v$, then
      \[
        \langle p_u - \frac12, p_v - \frac12 \rangle \geq \gamma\Delta^2m
      \]
    \end{itemize}
  \item Proof:
    \begin{itemize}
    \item The lower bound the probability of ``good neighborhood,'' that is, a user has many neighbors from the same type and few neighbors of different types
    \item This is done by controlling the exploration steps
    \end{itemize}
  \end{itemize}
\item \cite{bresler2016collaborative}
  \begin{itemize}
  \item Empirical evidence: item-item performs better
  \item Setting
    \begin{itemize}
    \item Setting: time $t$, a uniformly random user requests a recommendation
    \item and must be presented with a new one
    \item $N$ users, infinity set of items
    \item Each item $i$ has a type $L \in \{-1,1\}^N$ of all users approvals
    \item For each item, user $u$ has a binary preference
    \item Population of items is drawn from a distribution $\mu$ of types. 
    \item The algorithm can recommend a new item which corresponds to drawing an element from $\mu$
    \item Regret: with respect to perfect recommendations
    \end{itemize}
  \end{itemize}  
\item \cite{deshpande2012linear}
  \begin{itemize}
  \item Modeled CF as an online linear bandit problem. They assumed that some matrix factorization methods was used to extract item features $x_1,\ldots, x_N$
  \item The user (each user is modeled separately) has a parameter vector $\theta$
  \item Reward is $\theta^\top x_i$; goal is to have low regret by recommending good $x_i$
  \end{itemize}
\item \cite{bresler2017regret}
  \begin{itemize}
  \item 
  \end{itemize}
\item \cite{song2016blind}
  \begin{itemize}
  \item Model
    \begin{itemize}
    \item Every user $u$ has a features $x_1(u)$ and every item $i$ has a feature $x_2(i)$.
    \item Reward is $y(u,i) = f(x_1(u), x_2(i))+ \eta_{ui}$
    \item The \emph{blind} part of the regression is that only the indices $(u,i)$, not the features, are observed
    \item Otherwise, we could just apply some non-parametric regression to find $f$. 
    \item This problem is offline: assuming a certain sampling density of the entries of the user-item matrix, they provide high probability recovery guarantees
    \end{itemize}
  \end{itemize}  
\item There have been several papers proposing online matrix factorization methods
  \begin{itemize}
  \item \cite{abernethy2007online} Used online gradient descent to describe an online matrix factorization algorithm 
  \end{itemize}
\end{itemize}

\end{document}

### 2018-09-07 pap.tex
bib now works.
fixed the bug, i think.

@@
 \newpage
-\bibliography{../../big_bib}
+\bibliography{big_bib}
 
@@
     +  \sum_{i=1}^N R_i\left(\hist^{t},\rho\right)
+    \right]
+    \\
+    &=
+      \argmin_{q \in \triangle_{N}}
+  \sup_{p_t \in \triangle_{\mathcal D}'} 
+    \sum_{i=1}^N
+    \exop_{\hat\Delta_t \sim p_t}\left[
+    q^\top\hat D_i\Delta_t
+    +   R_i\left(\hist^{t},\rho\right)
     \right].
@@
 By construction, no entry of $q_t(\rho)$ is below $\gamma$. It is then easy to check that, for $q_t^* = \ex_{\rho}[q_t^*(\rho)]$ and $q_t = \ex_{\rho}[q_t(\rho)]$,
-\[
-  \ex_{I_t \sim q_t}\left[ \Delta_t ^\top e_{I_t} \right] = \Delta_t^\top q_t  
-  \leq \Delta_t^\top q_t^* + N\gamma
+\begin{align*}
+  \ex_{I_t \sim q_t}\left[ e_{I_t}^\top \Delta_t \right] = q_t^\top \Delta_t^\top 
+  \leq (q_t^*)^\top \Delta_t + N\gamma
   =
-  \ex_{I_t\sim q_t} \left[ \hat\Delta_t(I_t,j_t,q_t)^\top q_t^* \right] + N\gamma,
-\]
+  \sum_{i=1}^N \ex_{I_t\sim q_t} \left[(q_t^*)^\top D_i \hat\Delta_t(I_t,j_t,q_t)^\top \right] + N\gamma,
+\end{align*}
 where we have been explicit that $\hat\Delta_t(I_t,j_t,q_t)$ is a random variable that depends on the player actions and the $q_t$ (since $\Z_t$'s distribution is a function of $q_t$).
@@
   &\leq
-    \sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[(q_t^*)^\top\hat\Delta_t(I_t,j_t,q_t)
-    +
-  \sum_{i=1}^N\exop_\rho\left[R_i\left(\hist^{t},\rho\right)\right]\right]\\
+    \sup_{j_t} \sum_{i=1}^N \mathop{\ex}_{I_t\sim q_t}\left[(q_t^*)^\top D_i \hat\Delta_t(I_t,j_t,q_t)
+    +
+  \exop_\rho\left[R_i\left(\hist^{t},\rho\right)\right]\right]\\
   &\quad+(T-t+1)N\gamma.
@@
 \begin{align*}
-\lefteqn{\sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[(q_t^*)^\top\hat\Delta_t(I_t,j_t,q_t)
-    +
-  \mathop{\ex}_{\rho}\left[\sum_{i=1}^N \sup_{\pi} - \pi(x_t)^\top \hat\Delta_t(I_t,j_t,q_t) + A_i(\pi)
+\lefteqn{\sup_{j_t} \sum_{i=1}^N \mathop{\ex}_{I_t\sim q_t}\left[(q_t^*)^\top D_i \hat\Delta_t(I_t,j_t,q_t)
+    +
+  \mathop{\ex}_{\rho}\left[\sup_{\pi} - \pi(x_t)^\top \hat\Delta_t(I_t,j_t,q_t) + A_i(\pi)
   \right]\right] + (T-t+1)N\gamma}\\
   &\quad =
-    \sup_{j_t} \sum_{i=1}^N\mathop{\ex}_{I_t\sim q_t}\left[\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top \hat \Delta_t (I_t,j_t,q_t)
+    \sup_{j_t} \sum_{i=1}^N\mathop{\ex}_{I_t\sim q_t}\left[\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top D_i \hat \Delta_t (I_t,j_t,q_t)
     +
@@
   &\lefteqn{
-        \sup_{j_t } \sum_{i=1}^N\mathop{\ex}_{I_t\sim q_t}\left[\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top \hat \Delta_t(I_t,j_t,q_t)
+        \sup_{j_t } \sum_{i=1}^N\mathop{\ex}_{I_t\sim q_t}\left[\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top D_i \hat \Delta_t(I_t,j_t,q_t)
     +
@@
     \sup_{p_t\in\triangle_{\mathcal D}'} \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[
-    \sum_{i=1}^N\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top \hat \Delta_t
+    \sum_{i=1}^N\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top D_i \hat \Delta_t
     +
@@
     &\leq
-\mathop{\ex}_{\rho}\left[ \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^*(\rho)^\top \hat \Delta_t
+\mathop{\ex}_{\rho}\left[ \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^*(\rho)^\top D_i \hat \Delta_t
     +
@@
   \begin{align*}
-    &\lefteqn{\sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^*(\rho)^\top \hat \Delta_t
+    &\lefteqn{\sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^*(\rho)^\top D_i \hat \Delta_t
     +
@@
       \inf_{q_t}
-      \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top \hat \Delta_t
+      \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top D_i \hat \Delta_t
     +
@@
           \inf_{q_t}
-      \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top \hat \Delta_t
+      \sup_{p_t\in\triangle_{\mathcal D}'}\sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top D_i \hat \Delta_t
     +
@@
     \inf_{q_t}    
-    \sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top \hat \Delta_t
+    \sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top D_i \hat \Delta_t
     +
@@
     \inf_{j}  
-    \sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j^\top \hat \Delta_t
+    \sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j^\top D_i \hat \Delta_t
     +
@@
     \sum_{i=1}^N
-    \inf_{j}  \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j^\top \hat \Delta_t\right]
+    \inf_{j}  \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j^\top D_i \hat \Delta_t\right]
     +
@@
 \end{align*}
-Because $\pi$ is deterministic, we must have
-\[
-  \min_i \ex_{\hat\Delta_t \sim p_t}
-  [e_i^\top\hat\Delta_t]
+Because $\pi$ is deterministic, we must have, for all $i$,
+\[
+  \sum_{i=1}^N \min_j \ex_{\hat\Delta_t \sim p_t}
+  [e_j^\top D_i \hat\Delta_t]
+  =
+  \min_j \ex_{\hat\Delta_t \sim p_t}
+  [e_j^\top\hat\Delta_t]
   \leq
-  \ex_{\hat\Delta_t \sim p_t }[\pi(x_t)^\top\hat\Delta_t]
+  \ex_{\hat\Delta_t \sim p_t }[\pi(x_t)^\top \hat\Delta_t]
   \leq
+  \sum_{i=1}^N
   \ex_{\hat\Delta_t \sim p_t }[\pi(x_t)^\top D_i\hat\Delta_t],
@@
      \sup_{\pi\in\Pi} A_i(\pi)
-    +\min_j \ex_{\hat\Delta_t'\sim p_t}\left[e_j^\top \hat\Delta_t'\right]
+    +\min_j \ex_{\hat\Delta_t'\sim p_t}\left[e_j^\top D_i \hat\Delta_t'\right]
     -\pi(x_t)^\top D_i \hat\Delta_t

### 2018-09-08 pap.tex
working on a different relaxation.

@@
     +
-    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]\\
-  &=
-    \sup_{p_t\in\triangle_{\mathcal D}'}
-    \sum_{i=1}^N
-    \inf_{j}  \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j^\top D_i \hat \Delta_t\right]
-    +
-\mathop{\ex}_{\hat\Delta_t\sim p_t}\left[  \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]. 
+    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right].
+%   &=
+%     \sup_{p_t\in\triangle_{\mathcal D}'}
+%     \sum_{i=1}^N
+%     \inf_{j}  \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j^\top D_i \hat \Delta_t\right]
+%     +
+% \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[  \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]. 
 \end{align*}
@@
 \[
-  \sum_{i=1}^N \min_j \ex_{\hat\Delta_t \sim p_t}
+  \min_j \sum_{i=1}^N \ex_{\hat\Delta_t \sim p_t}
   [e_j^\top D_i \hat\Delta_t]
@@
         \sup_{p_t\in\triangle_{\mathcal D}'}
+    \inf_{j}  
+    \sum_{i=1}^N \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j^\top D_i \hat \Delta_t
+    +
+    \sup_{\pi\in\Pi} -\pi(x_t)^\top D_i\hat\Delta_t + A_i(\pi)\right]
+    }\\
+    &\leq
+    \sup_{p_t\in\triangle_{\mathcal D}'}
       \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t}\left[
-     \sup_{\pi\in\Pi} A_i(\pi)
-    +\min_j \ex_{\hat\Delta_t'\sim p_t}\left[e_j^\top D_i \hat\Delta_t'\right]
+      \ex_{\hat\Delta_t'\sim p_t}\left[\pi(x_t)^\top D_i \hat\Delta_t'\right]
+     +\sup_{\pi\in\Pi} A_i(\pi)
     -\pi(x_t)^\top D_i \hat\Delta_t
-  \right]
-  }\\
+  \right]\\
   &\leq
@@
 
-\iffalse
-\paragraph{A detail}\todo{remove this paragraph}
-While the regret bound, proven w.r.t. the two policies $\pi_1$ and $\pi_2$ and all fixed actions, 
-In particular, there exist a policy that is better that $\pi_1$ and $\pi_2$: the one that switches exactly correctly at $\beta$. To fix this, we can instead define
-\[
-  P_1 = q_1 \mathds{1}\{x \leq \beta-\beta_2\} + q_2 \mathds{1}\{x \geq \beta+\beta_1\}
-  + \frac{q_1+q_2}{2}\mathds{1}\{\beta-\beta_2\leq x \leq \beta+\beta_1\}
-\]
-and analogously for $P_2$. Then the ``perfect policy'' from before does not do any better than $\pi_1$ or $\pi_2$ in expectation.
-
-Unfortunately, this set of alternatives does not have the nice property that the feedback distribution of playing $\pi_1$ under $P_1$ and $P_2$ is identical; there is some small change from $S_1 q_a$ to $S_1 q_b$.
-It might be possible to balance terms and control the regret.
-\fi
+\section{Structured relaxation}
+The partial monitoring problem allows us to write $\Delta_t = V(I_t) S_{I_t} e_{j_t}$ (for some appropriate matrices $V(1),\ldots, V(N)$) and therefore form the unbiased estimate
+\[
+  \hat\Delta_t = V(I_t) Y_t X_t
+\]
+where $X_t \sim\text{Bern}\left(\frac{\gamma}{V_\infty} \frac{\Vert V(I_t)^\top Y_t \Vert}{q_t(I_t)}\right)$.
+The proof for unbiasedness is essentially the same as for the previous relaxation.
+In particular, $\hat\Delta_t$ can only take on $NM+1$ different values enumerated by the set $U = \{ V(i) S_i e_j: i\in\underbar N, j \in\underbar M\}\cup 0$.
+Let $u_1,\ldots, u_k$ be an enumeration of this set.
 

### 2018-09-15 pap.tex
working through new relaxation.

@@
 \usepackage{tikz}
-%\usepackage[colorinlistoftodos, textwidth=26mm, shadow,color=blue!30!white]{todonotes}
-\usepackage[disable]{todonotes}
 
@@
 \usepackage{wrapfig}
+
+\usepackage[colorinlistoftodos, textwidth=26mm, shadow,color=blue!30!white]{todonotes}
+%\usepackage[disable]{todonotes}
 
@@
 \section{Structured relaxation}
-The partial monitoring problem allows us to write $\Delta_t = V(I_t) S_{I_t} e_{j_t}$ (for some appropriate matrices $V(1),\ldots, V(N)$) and therefore form the unbiased estimate
-\[
-  \hat\Delta_t = V(I_t) Y_t X_t
-\]
-where $X_t \sim\text{Bern}\left(\frac{\gamma}{V_\infty} \frac{\Vert V(I_t)^\top Y_t \Vert}{q_t(I_t)}\right)$.
-The proof for unbiasedness is essentially the same as for the previous relaxation.
-In particular, $\hat\Delta_t$ can only take on $NM+1$ different values enumerated by the set $U = \{ V(i) S_i e_j: i\in\underbar N, j \in\underbar M\}\cup 0$.
-Let $u_1,\ldots, u_k$ be an enumeration of this set.
+The partial monitoring problem allows us to write $\Delta_t = V(I_t) S_{I_t} e_{j_t}$ (for some appropriate matrices $V(1),\ldots, V(N)$) and therefore form the unbiased estimate \todo{either assume that $V(I_t)Y_t$ is never zero or resolve this case}
+\[
+  \hat\Delta_t = \frac{V(I_t) Y_t}{\Vert V(I_t) Y_t\Vert}  Z_t \frac{ V_\infty}{\gamma}
+\]
+where $Z_t \sim\text{Bern}\left(\frac{\gamma}{V_\infty} \frac{\Vert V(I_t)^\top Y_t \Vert}{q_t(I_t)}\right)$.
+
+Two important properties of the unbiased loss estimate must hold. First, it must be unbiased. Treating $j_t$ like a constant and recalling that $Y_t = S_{I_t} e_{j_t}$,
+\begin{align*}
+  \ex_{I_t, Z_t}\left[ \hat\Delta_t \right]
+  &=
+    \ex_{I_t}\left[
+    \frac{V(I_t) S_{I_t} e_{j_t}}{\Vert V(I_t) S_{I_t} e_{j_t} \Vert}  \ex[Z_t|I_t] \frac{ V_\infty}{\gamma}
+    \right]
+  \\
+  &=
+    \ex_{I_t}\left[
+    \frac{V(I_t) S_{I_t} e_{j_t}}{\Vert V(I_t) S_{I_t} e_{j_t}\Vert}  \frac{\gamma}{V_\infty} \frac{\Vert V(I_t) S_{I_t} e_{j_t} \Vert}{q_t(I_t)}
+    \frac{ V_\infty}{\gamma}
+    \right]
+  \\
+  &=
+    \ex_{I_t}\left[
+    \frac{V(I_t) S_{I_t} e_{j_t}}{q_t(I_t)}
+    \right]
+  \\
+  &=
+    \Delta_t.
+\end{align*}
+
+Second, it must have large probability of being zero. Calculating,
+\begin{align*}
+  P(Z_t = 0)
+  &=
+    \sum_{i=1}^N
+    q_t(i)
+    \left( 1 - \frac{\gamma}{V_\infty} \frac{\Vert V(i) S_i e_{j_t} \Vert}{q_t(i)} \right)
+  \\
+  &=
+    1-
+    \sum_{i=1}^N
+    \frac{\gamma}{V_\infty} \Vert V(i)S_i e_{j_t}  \Vert
+  \\
+  &\geq
+  1-N \gamma.
+\end{align*}
+
+Another property is that $\hat\Delta_t$ can only take on $NM+1$ different values enumerated by the set $U = \{ V(i) S_i e_j: i\in\underbar N, j \in\underbar M\}\cup 0$, and let $\overline U_j = \{ \frac{V_\infty V(i)S_i e_j}{\gamma \Vert V(i)S_i e_j \Vert}: i\in\underbar N \}$ be the enumeration of all directions $\hat\Delta_t$ could take if the adversary chooses action $j$.
+To summarize, $\hat\Delta_t$ is unbiased  and, for every $j_t$, takes values in the set
+$\overline U_j$ and has probability at least $1-N\gamma$ of being zero.
+
+Hence, when we relax the optimization around $\hat\Delta_t$, we can relax it to $p_t \in \triangle_{\mathcal D}'$, where $\triangle_{\mathcal D}'$ is equal to the distributions over $\bigcup_j \overline U_j$ with $p_t(0) \geq 1-N\gamma$.
+
+For the sequel, redefine the following notation (this, once this proof is verified, it will be easier to just slot in).
+Let $\hat\Delta_t$ have the form above, and let $\overline U = \bigcup_j \overline U_j$ with $K := |\overline U|$.
+We will enumerate $\overline U$ with $u_1,\ldots, u_K$, and use $\U$ to denote the matrix with these columns.
+Finally, let $\eepsilon_t$ and $\Z_t$ be random diagonal matries with $K$ independent Rademacher random variables on the diagonal and $K$ independent random variables with value $V_\infty \gamma^{-1}$ with probability $N\gamma$ and $0$ otherwise.
+
+We will also say that $\delta\in \{ -1,0,1\}$ is a \emph{Lazy Rademacher} random variable with parameter $a$ if $p(\delta=0) = a$ and $p(\delta = -1) = p(\delta = 1) = \frac{1-a}{2}$.
+
+Finally, define
+\[
+  R(\hist^{t-1},\rho_t) =
+        \sup_{\pi\in\Pi}
+      -
+      \sum_{s=1}^{t} \pi(x_s)^\top \hat\Delta_s
+      -
+      \sum_{s=t+1}^{T} 2 \pi(x_s)^\top \eepsilon_{s}\Z_s \U
+      + (T-t)\gamma.
+  \]
+We can then prove:
+\begin{theorem}\label{thm:admissibility_new}  
+  The relaxation
+  \begin{align}
+    \rel(\hist^{t-1})
+    &= \exop_{
+      \rho_t}
+      \left[
+      R(\hist_{t-1},\rho_t)
+      \right]
+      \label{eqn:relaxation.def}
+  \end{align}
+  and the strategy that samples $\rho_t = \bigcup_{s \in t+1:T} \{x_{s},\eepsilon_{s},\Z_{s}\}$ and plays  $q_t= (1-N\gamma)q_t^* + \gamma \ones$ for
+\begin{align}
+  q_t^*
+  &=
+  \argmin_{q \in \triangle_{N}}
+  \sup_{p_t \in \triangle_{\mathcal D}'} 
+    \exop_{\hat\Delta_t \sim p_t}\left[
+    q^\top\hat\Delta_t
+    +   R\left(\hist^{t-1},\rho_t\right)\right]
+  \label{eqn:strategy.def}
+\end{align}
+are admissible.
+\end{theorem}
+
+\begin{proof}[Proof of Theorem~\ref{thm:admissibility}]
+  The base case is a consequence of the unbiasness of $\hat\Delta_t$.Using the convexity of supremum and the unbiasedness of $\hat\Delta_t$, we have 
+  \begin{align*}
+  \ex_{I_{1:T},\hat Z_{1:T}}\left[
+  \rel(\hist^T)
+  \right]
+  &=
+    \ex_{I_{1:T},\hat Z_{1:T}}\left[
+    \sup_{\pi\in\Pi}
+    -\sum_{s=1}^T \pi(x_t)^\top \hat \Delta_t
+    \right]\\
+  &\geq
+    \sup_{\pi\in\Pi}
+    \sum_{s=1}^T
+    \ex_{I_{1:T},\hat Z_{1:T}}\left[
+    - \pi(x_t)^\top \hat \Delta_t
+    \right]\\
+  &=
+    \sup_{\pi\in\Pi}
+    -\sum_{s=1}^T \pi(x_t)^\top \Delta_t.
+  \end{align*}
+
+  We now check the inductive step.
+We define $\rho = (x_t,\eepsilon_t,\Z_t)_{t+1:T}$ to the collection of random variables in the relaxation.
+Recall that our aim is to prove admissibility of the strategy $q_t(\rho) = (1-N\gamma)q_t^*(\rho)+ \gamma \ones$ where $q_t^*(\rho)$ was defined by
+  \begin{align*}
+  q_t^*(\rho)
+  &=
+  \argmin_{q \in \triangle_{N}}
+  \sup_{p_t \in \triangle_{\mathcal D}'} 
+    \exop_{\hat\Delta_t \sim p_t}\left[
+    q^\top\hat\Delta_t
+    + R\left(\hist^{t},\rho\right)
+    \right].
+\end{align*}
+By construction, no entry of $q_t(\rho)$ is below $\gamma$. It is then easy to check that, for $q_t^* = \ex_{\rho}[q_t^*(\rho)]$ and $q_t = \ex_{\rho}[q_t(\rho)]$,
+\begin{align*}
+  \ex_{I_t \sim q_t}\left[ e_{I_t}^\top \Delta_t \right] = q_t^\top \Delta_t^\top 
+  \leq (q_t^*)^\top \Delta_t + N\gamma
+  =  
+  \ex_{I_t\sim q_t} \left[(q_t^*)^\top \hat\Delta_t(I_t,j_t,q_t)^\top \right] + N\gamma,
+\end{align*}
+where we have been explicit that $\hat\Delta_t(I_t,j_t,q_t)$ is a random variable that depends on the player actions and the $q_t$ (since $\Z_t$'s distribution is a function of $q_t$).
+
+For every fixed $x_t$, we can apply the above inequality and unpack the definition of the relaxation to find
+\begin{align*}
+\lefteqn{  \sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[e_{I_t}^\top\hat\Delta_t(I_t,j_t,q_t)
+  +  \rel(\hist^t) \right]}&\\
+  &=
+    \sup_{j_t } \mathop{\ex}_{I_t\sim q_t}\left[e_{I_t}^\top\hat\Delta_t(I_t,j_t,q_t)
+    +  \exop_\rho\left[R\left(\hist^{t},\rho\right)\right]\right]
+    + (T-t)N\gamma\\
+  &\leq
+    \sup_{j_t}  \mathop{\ex}_{I_t\sim q_t}\left[(q_t^*)^\top \hat\Delta_t(I_t,j_t,q_t)
+    +
+  \exop_\rho\left[R\left(\hist^{t},\rho\right)\right]\right]
+  +(T-t+1)N\gamma.
+\end{align*}
+Defining $A(\pi) = -\sum_{s=1}^{t-1} \pi(x_s)^\top \hat \Delta_s - \sum_{s=t+1}^T 2\pi(x_s)^\top  \eepsilon_{s} \Z_{s}$, the previous line is equal to
+\begin{align*}
+    \sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top \hat \Delta_t (I_t,j_t,q_t)
+    +
+    \sup_{\pi\in\Pi} -\pi(x_t)^\top \hat\Delta_t(I_t,j_t,q_t) + A(\pi)\right]\right] + (T-t+1)N\gamma.
+\end{align*}
+This optimization is intractable without strong assumptions on $\Pi$; the $j_t$ optimization needs to account for how the supremum over the policy class will be affected.
+Therefore, we reduce the constraints on the adversary to relax the problem by allowing play of distributions over $\hat\Delta_t$ instead of constraining the $\hat\Delta_t$ to correspond to a specific choice of $I_t,j_t$, and $q_t$.
+
+We carefully defined $\hat\Delta_t(I_t,j_t,q_t)$ so that it would take on finite support and have a large probability of being zero. Conditioning on $\rho$, we can calculate
+\begin{align*}
+  P(\hat\Delta_t = 0)
+  &=
+    \sum_{i=1}^N
+    q_t(i)
+    \left( 1 - \frac{\gamma}{V_\infty} \frac{\Vert V(i) S_i e_{j_t} \Vert}{q_t(i)} \right)
+  \\
+  &=
+    1-
+    \sum_{i=1}^N
+    \frac{\gamma}{V_\infty} \Vert V(i)S_i e_{j_t}  \Vert
+  \\
+  &\geq
+  1-N \gamma.
+\end{align*}
+Therefore, the distribution of $\hat\Delta_t(I_t,j_t,q_t)$ is always in $\triangle_{\mathcal D}'$, and we obtain an upper bound by allowing the adversary to play distributions over a random variable named $\hat\Delta_t$ with distribution in $\triangle_{\mathcal D}'$.
+With this substitution, $I_t$ and $j_t$ no longer appear in the expression.
+Suppressing the $(T-t+1)N\gamma$ term, we have
+\begin{align*}
+  &\lefteqn{
+        \sup_{j_t } \mathop{\ex}_{I_t\sim q_t}\left[\mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top \hat \Delta_t(I_t,j_t,q_t)
+    +
+    \sup_{\pi\in\Pi} -\pi(x_t)^\top\hat\Delta_t(I_t,j_t,q_t) + A(\pi)\right]\right]
+    }\\
+  &\leq
+    \sup_{p_t\in\triangle_{\mathcal D}'} \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[
+    \mathop{\ex}_{\rho}\left[ q_t^*(\rho)^\top  \hat \Delta_t
+    +
+    \sup_{\pi\in\Pi} -\pi(x_t)^\top \hat\Delta_t + A(\pi)\right]\right]\\
+    &\leq
+\mathop{\ex}_{\rho}\left[ \sup_{p_t\in\triangle_{\mathcal D}'} \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^*(\rho)^\top  \hat \Delta_t
+    +
+  \sup_{\pi\in\Pi} -\pi(x_t)^\top \hat\Delta_t + A(\pi)\right]\right].
+\end{align*}
+Our ability to move the expectation over $\rho$ to the outside allows us to obtain the same bound in expectation by sampling a single $\rho$ and playing $q_t(\rho)$ instead of calculating the infimum over $q$.
+
+For the remainder, fix some $\rho$. We defined
+\[
+  q_t^*(\rho)
+  =
+  \argmin_{q \in \triangle_{N}}
+  \sup_{p_t \in \triangle_{\mathcal D}'}
+  \ex_{\hat\Delta_t\sim p_t}\left[
+  q^\top \hat \Delta_t
+    +
+    \sup_{\pi\in\Pi} -\pi(x_t)^\top \hat\Delta_t + A(\pi)\right],
+  \]
+  and so
+  \begin{align*}
+    &\lefteqn{\sup_{p_t\in\triangle_{\mathcal D}'} \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^*(\rho)^\top  \hat \Delta_t
+    +
+  \sup_{\pi\in\Pi} -\pi(x_t)^\top \hat\Delta_t + A(\pi)\right]}\\
+    &=
+      \inf_{q_t}
+      \sup_{p_t\in\triangle_{\mathcal D}'} \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top  \hat \Delta_t
+    +
+      \sup_{\pi\in\Pi} -\pi(x_t)^\top \hat\Delta_t + A(\pi)\right]\\
+      &=
+    \sup_{p_t\in\triangle_{\mathcal D}'}
+    \inf_{q_t}    
+        \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ q_t^\top  \hat \Delta_t
+    +
+    \sup_{\pi\in\Pi} -\pi(x_t)^\top \hat\Delta_t + A(\pi)\right]\\
+  &\leq
+    \sup_{p_t\in\triangle_{\mathcal D}'}
+    \inf_{j}  
+    \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j^\top  \hat \Delta_t
+    +
+    \sup_{\pi\in\Pi} -\pi(x_t)^\top \hat\Delta_t + A(\pi)\right].
+  \end{align*}
+
+Because $\pi$ is deterministic, we must have, for all $i$,
+\[
+  \min_j \ex_{\hat\Delta_t \sim p_t}
+  [e_j^\top \hat\Delta_t]
+  =
+  \min_j \ex_{\hat\Delta_t \sim p_t}
+  [e_j^\top\hat\Delta_t]
+  \leq
+  \ex_{\hat\Delta_t \sim p_t }[\pi(x_t)^\top \hat\Delta_t],
+\]
+which allows us to upper bound the previous expression. Performing the usual symmeterization (using $\epsilon$ as a single Rademacher random variable) yields
+\begin{align*}
+  &\lefteqn{
+        \sup_{p_t\in\triangle_{\mathcal D}'}
+    \inf_{j}  
+    \mathop{\ex}_{\hat\Delta_t\sim p_t}\left[ e_j^\top  \hat \Delta_t
+    +
+    \sup_{\pi\in\Pi} -\pi(x_t)^\top \hat\Delta_t + A(\pi)\right]
+    }\\
+    &\leq
+    \sup_{p_t\in\triangle_{\mathcal D}'}
+     \ex_{\hat\Delta_t\sim p_t}\left[
+      \ex_{\hat\Delta_t'\sim p_t}\left[\pi(x_t)^\top  \hat\Delta_t'\right]
+     +\sup_{\pi\in\Pi} A(\pi)
+    -\pi(x_t)^\top  \hat\Delta_t
+  \right]\\
+  &\leq
+    \sup_{p_t\in\triangle_{\mathcal D}'}
+     \ex_{\hat\Delta_t\sim p_t}\left[
+     \sup_{\pi\in\Pi} A(\pi)
+    +\ex_{\hat\Delta_t'\sim p_t}\left[\pi(x_t)^\top  \hat\Delta_t'\right]
+    -\pi(x_t)^\top  \hat\Delta_t
+  \right]\\
+  &\leq
+    \sup_{p_t\in\triangle_{\mathcal D}'}
+     \ex_{\hat\Delta_t\sim p_t, \hat\Delta_t'\sim p_t}\left[
+        \sup_{\pi\in\Pi} A(\pi)+
+        \pi(x_t)^\top \left(\hat\Delta_t'-\hat\Delta_t\right)
+    \right]\\
+  &=
+    \sup_{p_t\in\triangle_{\mathcal D}'}
+     \ex_{\hat\Delta_t\sim p_t, \hat\Delta_t'\sim p_t,\epsilon_i}\left[
+        \sup_{\pi\in\Pi} A(\pi)+
+        \epsilon\pi(x_t)^\top \left(\hat\Delta_t'-\hat\Delta_t\right)
+    \right]\\
+    &\leq
+    \sup_{p_t\in\triangle_{\mathcal D}'}
+     \ex_{\hat\Delta_t\sim p_t,\epsilon_i}\left[
+        \sup_{\pi\in\Pi} A(\pi)+
+        2\epsilon\pi(x_t)^\top \hat\Delta_t
+        \right].
+\end{align*}
+
+Next, we will upper bound the previous expression for all $p_t$.
+Condition on $\hat \Delta_t= u_i$ and define the random variable $Y = \sum_{j\neq i} u_i + \delta_j u_j$, where $\delta_j$ is a Lazy Rademacher random variable with parameter $N\gamma$.
+We clearly have $\ex[Y|\hat \Delta_t] = u_i$ and $\epsilon Y \stackrel{\mathcal L}{=} \sum_{i=1}^K \delta_i u_i$.
+
+Since $\hat\Delta_t = e_i \hat\Delta_t(i)$, each term in the sum only involves one coordinate if $\hat\Delta_t$.
+Therefore, it is without loss of generality to assume that $p_t$ is a product distribution.
+Using $\triangle_{\{0,V_\infty \gamma^{-1}\}}$ to denote distributions over the singletons $0$ and $V_\infty\gamma^{-1}$, define
+\[
+  \triangle_1' := \left\{ p \in \triangle_{\{0,V_\infty \gamma^{-1}\}}: p(0)\geq 1-N\gamma\right\}
+\]
+and observe that if $\hat\Delta_t\sim p_t\in\triangle_{\mathcal D'}$, then $\epsilon_i X_i e_t \stackrel{\mathcal L}{=} \hat\Delta_t(i)$ for some $X_i\sim p_i\in\triangle_1'$. We then have 
+\begin{align*}
+  \sup_{p_t\in\triangle_{\mathcal D}'}
+    \sum_{i=1}^N \ex_{\hat\Delta_t\sim p_t,\epsilon}\left[
+        \sup_{\pi\in\Pi} A(\pi)+
+        2\epsilon\pi(x_t)^\top D_i \hat\Delta_t
+  \right]
+  &=
+      \sup_{p_1,\ldots,p_N \in\triangle_1'}
+    \sum_{i=1}^N \ex_{X_i\sim p_i,\epsilon_{1:N}}\left[
+        \sup_{\pi\in\Pi} A(\pi)+
+        2\pi(x_t)^\top e_i \epsilon_i X_i
+    \right]\\
+      &=
+        \sum_{i=1}^N
+        \sup_{p_i \in\triangle_1'}
+        \ex_{X_i\sim p_i,\epsilon_i}\left[
+        \sup_{\pi\in\Pi} A(\pi)+
+        2\pi(x_t)^\top e_i \epsilon_i X_i
+    \right].
+\end{align*}
+
+We will now argue that a witness to the supremum of 
+\begin{align*}
+    \sup_{p_i\in\triangle_1'}
+  \ex_{X_i\sim p_i,\epsilon_i}\left[
+        \sup_{\pi\in\Pi} A_i(\pi)+
+        2\pi(x_t)^\top e_i \epsilon_i X_i
+        \right]
+\end{align*}
+is the distribution that puts mass $N\gamma$ on $V_\infty\gamma^{-1}$ and the rest on $0$. Define the convex function 
+$g(x) :=  \sup_{\pi\in\Pi} A_i(\pi)+ 2\pi(x_t)^\top e_i x$.
+The expectation $\ex_{\epsilon}\left[ g(\epsilon x)\right]$ is increasing on $x\geq 0$.
+To see this, consider some $0\leq a<b$.
+We can write $a = \theta b + (1-\theta)(-b)$ for some $\theta$ and use the definition of convexity to conclude
+ \begin{align*}
+   \ex_{\epsilon}\left[ g(\epsilon a) \right]
+   = 
+   \frac{g(a)+g(-a)}{2}
+   \leq
+   \frac{\theta g(b)+(1-\theta)g(-b)
+   +(1-\theta) g(b)+\theta g(-b)
+   }{2}
+   =
+   \ex_{\epsilon}\left[ g(\epsilon b) \right].
+ \end{align*}
+ Since $\ex_{\epsilon}\left[ g(\epsilon x)\right]$ is increasing, the supremum of $p_i$ puts maximum mass on $V_\infty\gamma^{-1}$.
+Hence, defining the random vector $Z_{t}$ with elements $P(Z_{t,i} = 0) = 1-N\gamma$ and $P( Z_{t,i} = V_\infty\gamma^{-1}) = N\gamma$, we have
+ \begin{align*}
+    \sup_{p\in\triangle_1'}
+  \ex_{X_i\sim p,\epsilon_i}\left[
+        \sup_{\pi\in\Pi} A_i(\pi)+
+        2\pi(x_t)^\top e_i \epsilon_i X_i
+        \right]
+   &=
+  \ex_{\epsilon_i}\left[
+        \sup_{\pi\in\Pi} A_i(\pi)+
+        2\epsilon_i\pi(x_t)^\top D_i \Z_t
+   \right]\\
+      &=
+  \ex_{\eepsilon_t}\left[
+        \sup_{\pi\in\Pi} A_i(\pi)+
+        2\pi(x_t)^\top \eepsilon_{t,i} \Z_t
+        \right].
+\end{align*}
+
+In total, we have shown that, for every fixed $x_t$ and $\rho$, playing $q_t(\rho)$ allows the bound
+\begin{align*}
+    \sup_{j_t} \mathop{\ex}_{I_t\sim q_t}\left[ e_{I_t}^\top \Delta_t
+  + \rel(\hist^t)\right]
+  &\leq
+    \ex_\rho \left[
+  \ex_{\epsilon_t, \Z_t}\left[
+  \sum_{i=1}^N
+        \sup_{\pi\in\Pi} A_i(\pi)+
+        2\pi(x_t)^\top \eepsilon_{t,i} \Z_t
+  \right]
+  \right]
+  + (T-t+1)N\gamma\\
+  &=
+  \rel(\hist^{t-1}),
+\end{align*}
+as required.
+\end{proof}
+
+The proof of admissibility seems to be unchanged until the line
+\[
+      \sup_{p_t\in\triangle_{\mathcal D}'}
+      \ex_{\hat\Delta_t\sim p_t,\epsilon}\left[
+        \sup_{\pi\in\Pi} A_i(\pi)+
+        2\epsilon\pi(x_t)^\top
+        \hat\Delta_t
+      \right]
+\]
 
